### 2019-10-15 Geopyter.tex
init for jgs

\documentclass[letter, 11pt]{article}

% Need to do some more reading: http://plain-text.co/pull-it-together.html
% Convert from LaTeX to Markdown:
% pandoc --standalone --wrap=none --base-header-level=2 -f latex -t gfm+smart+citations --filter=pandoc-citeproc --bibliography=Geopyter.bib --csl=apalike.csl -M Title="GeoPyTeR: Mashable (Geography) Teaching Resources for Python" -B header.md Geopyter.tex -o README.md

\newcommand{\gp}{\textsc{g}eo\textsc{p}y\textsc{t}e\textsc{r}~\/}
\newcommand{\eg}{e.g.~\/}
\newcommand{\ie}{i.e.~\/}
\newcommand{\comment}[1]{\todo[inline, color=green!40]{#1}}

%% Language and font encodings
\usepackage[english]{babel}
\usepackage[utf8x]{inputenc}
\usepackage[T1]{fontenc}
\usepackage[hyphens]{url}
\usepackage{fancyvrb} % facny verbatim
\usepackage{textcomp}
\usepackage{upquote}  % Turns quotes in verbatim to straight
\usepackage{varioref} % for vpageref[above]
\usepackage{fullpage}
\usepackage{graphicx} 
\usepackage[font=small,labelfont=bf]{caption}
%% Sets page size and margins
%\usepackage[letter,top=3cm,bottom=2cm,left=3cm,right=3cm,marginparwidth=1.75cm]{geometry}

%% For citations
%\usepackage{natbib}
\usepackage[%
% numberedbib,
  natbibapa
]{apacite}

%% Useful packages
\usepackage{amsmath}
\usepackage[colorinlistoftodos]{todonotes}
\usepackage[colorlinks=false, allcolors=black]{hyperref}
\def\tightlist{}

\title{Geographical Python Teaching Resources: GeoPyTeR}
\author{Jonathan Reades \and Sergio Rey}

\begin{document}
\maketitle

\begin{abstract}
  Abstract goes here.
\end{abstract}

\section{Introduction}\label{introduction}

Although \citet{Donoho2017} traces the origins of data science back to \citeauthor{Tukey1962}'s \textit{The future of data analysis} (\citeyear{Tukey1962}), its emergence has led to a flourishing of tools and programmes as universities seek to engage with this nascent---and lucrative---interdisciplinary field. One of the most profound practical impacts of this development has been on the ways in which developers share and execute code: although systems for combining code, commentary, and results have been around for some time (\eg R-Markdown) these were intended primarily for replicating research outputs, not for the online teaching of a new technique or writing of an interactive introduction to a software library. The `open' ethos and rapid growth of data science, and the need for a lightweight interactive code-sharing platform mean that teachers in many disciplines are playing catch-up on both the methods taught and the formats used.

As a discipline, geography has been acutely aware of these developments: changes in the volume and extent of geographical data available (\eg \citealp{Graham2013,gonzalez2013big,reades2016}), and in the number and range of methods employed to identify patterns in those data (\eg \citealp{Fan2016,Naik2017,Santibanez2015,Stevens2015,ArribasBel2017}), are allowing us to tackle questions thought unanswerable just a few years ago. At the same time, however, there is growing evidence of a `hollowing out' of traditional domain-specific skills: most of what a trained geographer used to do with specialist software can now be done using free online resources (\eg Google's Fusion Tables, MapBox, Carto, ArcOnline), but there remains unmet demand for graduates who can \textit{code} in order to take advantage of the latest (geo)algorithms and techniques \citep{Singleton2014,Singleton2016}.

\section{Statement of Need}

Regardless of whether this is a revolution \citep{Wyly2014,Torrens2010} or an evolution \citep{Barnes2013,Barnes2014,ArribasBel2018}, there is no question that teaching students and researchers how to code and how to \textit{think} computationally presents new challenges \citep{Etherington2016,Muller2014}. Specifically, because of the way that these changes have unfurled, the majority of resources used in the classroom are being developed from scratch at each institution, and often by newly-appointed lecturers and assistant professors coming from the relatively small number of elite research facilities active in this domain (see discussion in \citealp{esrc2013}). Not only has this led to the duplication of effort at multiple sites, but it has also had an enormous impact on the productivity---as commonly measured by institutions---of early career researchers, many of whom are in their first teaching post.

The overarching purpose of this software is therefore to address two seemingly contradictory issues: the recognition that no two instructors teach in precisely the same way, and the fact that few instructors have the luxury of both time and tenure to develop compelling new course material in splendid isolation. This where the Geographical Python Teaching Resource (\textsc{g}eo\textsc{p}y\textsc{t}e\textsc{r}) comes into play: although developed with geographers in mind, it provides a generic means by which instructors in any discipline can selectively incorporate and re-mix programming and conceptual content from existing Jupyter notebooks while providing their own `gloss' for this content. Our approach therefore seeks to develop a richer, open community of developers and teachers who work \textit{together} on new teaching materials, and to support instructors in quickly adapting existing materials to new delivery formats.

\section{Origins and Inspirations}\label{origins-and-inspirations}

Like many open source projects, \gp grew out of a need to scratch an itch: in our roles as teachers of (geo)computational concepts and methods we grappled on an annual basis with the demands of developing and updating instructional material with complex interdependencies. And, while the fundamental computing and analytic concepts may remain fairly stable, the field is highly-dynamic in terms of both the content and the rate of evolution of new features in contributed code upon which we rely for functionality. Currently, every time a widely-used library package is updated a protracted sequence of revisions ensues: some teachers, pressed for time, may resist making use of---or even alerting students to---the new features; others may leap in enthusiastically, only to discover that they hadn't fully thought through the challenges of conflicting dependencies and evolving feature sets.

We initially explored with others---faculty with whom we regularly collaborate---the potential to `divide and conquer': we would each focus on different topics and then combine these contributions to create a full course. Though received with some enthusiasm, this plan did not survive contact with reality: first, when actually faced with the uncertain promise of receiving a full course in return for our efforts, many of us felt more secure developing our own material; second, we had underestimated the value that we each attach to the ability to personalise material to our own style of teaching. Since the shared materials were being developed individually according to the individual's own `style', any rationalisation or personalisation would have to be done by the instructor at a later date and without support from the rest of the project.

Going back to the drawing board, we looked again at our own experiences of successful programming and instructional efforts. In particular, we took note of two powerful streams of thought in software and education: the importance attached by the Free and Open Source Software (\textsc{foss}) movement to both peer-recognition \citep{raymond_cathedral_1999} and distributed version control; and the rise of Virtual Learning Environments (\textsc{vle}s) and of Massive Open Online Courses (\textsc{mooc}s) with their use of rich, online learning experiences (\eg \citealp{Trafford2011,Cabiria2012}). Reflecting on our initial failure, we decided against trying to supply a single set of polished course materials and instead looked for a way to provide teachers with building blocks from which they could assemble an \textit{individualised} course suited to their institutional and pedagogical needs.

We drew inspiration from the \textsc{foss} iPython project \citep{perez_2007_ipython} and its transformative impact on the sharing of integrated code, text, and multimedia resources. And we also looked to the Python Spatial Analysis Library (\textsc{p}y\textsc{sal}) project in which contributors take ownership of application areas through managed collaboration instead of a contributory `free-for-all'. The subsequent emergence of the Jupyter project \cite{kluyver16} from iPython has led to the rapid growth of a rich ecosystem of tools that support the kinds of interaction needed to make this an effective teaching resource: multiuser-notebooks (jupyterhub), automation of grading (nbgrader), interactive widgets for visualisation (ipywidgets), and packages to facilitate manipulation of the notebook itself (nbformat and notebook).

With these criteria and resources in mind, we began looking for ways to lower the `entry costs' for faculty to both take from, and share into, the corpus of instructional material, as well as for a way to address the customisation/localisation challenge. \gp was our solution: a single, open hub of teaching concepts and practical programming tasks that can be flexibly assembled into classes, or entire courses! In the \gp approach we recognise that teachers need to be able to develop their own classroom `story' at their own pace (and in conformance with their institution's teaching pattern), while supporting them with a library of up-to-date materials from which they can pick and choose. So in much the same way that a computer application is compiled from the contributions of many developers, \gp allows individual classes and entire courses to be `compiled' from the contributions of many skilled teachers and developers.

\section{System Architecture}\label{system-architecture}

At least in our experience, the pace of technological and methodological change is such that the gap between what instructors know and what students know is substantial, and in many cases might even be growing. The differences within individual student cohorts may be greater still, however, and while students need to be able to begin their learning `journey' at some arbitrary point in the progression from `learning to code' towards `computational social scientist' \citep{Lazer2009}, instructors require the ability to assemble a course that meets students where they \emph{are} rather than where the instructor or institution might \emph{wish} them to be.

The guiding insight behind \gp is that instruction in computation proceeds from fundamental units of learning typically built around \textit{computing} concepts (variables, lists/arrays, dictionaries/hashes, functions/subroutines, etc.) to fundamental units of learning built around \textit{analytic} concepts (cluster analysis, point patterns, spatial autocorrelation, etc.). These units must then be assembled in a way that speaks to the student cohort and its background; we mean this in two ways: first, that the examples used must speak to budding geographers, political scientists, historians, etc.; and second, that it should be possible to develop courses for different cohorts without having to start over from scratch. In other words, how can we enable teachers to reuse many of the building blocks employed in an advanced course for Masters students in an introductory undergraduate course?

\subsection{Components}\label{components}

From these constraints it was clear that our system needed to support a `bottom-up' approach to instructional design. So although the development of a course or class could---and probably should---start with a clear set of learning aims and outcomes, at a certain point the instructor will be searching for a examples with which to teach a particular concept: What \emph{is} a list or dictionary? What is \textit{k}-means clustering? We settled on the term `Atoms' to refer to these basic instructional units and, much like entries in the atomic table, we felt that they could be grouped together into sets of related concepts: the fundamentals of programming, point pattern analysis, machine learning, etc.

Each atom would employ \emph{domain-specific} illustrative examples and code so as to anchor learning in relevant problems and applications. An atom on arrays written for geographers could start by showing how these simple data structures can be used to hold data about cities (\eg name, country, population); but a subsequent unit could build on this with an an illustration of how nested data structures allow us to add city location using nested coordinate pairs (\eg latitude and longitude within an array-of-arrays). This is why managed collaboration around sets of related concepts is so important: quite aside from the fact that no \textsc{foss} project succeeds without some level of executive control, it's critical that \textit{someone} is responsible for ensuring a degree of consistency in how the units fit together within a set of closely-related topics.

The purpose of the bottom-up approach is that these units can then be flexibly assembled into Sessions: from the \textit{same ingredients} (\ie atoms) the instructor could create two quite different modules---one for delivery in person, one for delivery online---by organising the elements in different ways. Quite simply, we don't want to have to rewrite material for each format: lecture/practical, `flipped' classroom, or \textsc{mooc}. Furthermore, sessions designed for experts (\eg those pursuing continuing online education) might be able to `move' students through many more atoms of instruction in a single session than a similarly laid out (in terms of time) session designed for first-time programmers in an undergraduate programme. So sessions need to be able to \emph{incorporate} atoms in a way that minimises the level of effort involved in finding the `best' way to, for example, explain the concept of recursion while maximising the ability of the instructor to relate this concept to the students' practical experience (\eg by providing a `context' that is anchored in a locally relevant `story' or data).

Naturally, Sessions can then be grouped into Learning Modules (modules) that offer a coherent instructional programme over a period of weeks or months. Modules represent the highest level of abstraction in the proposed system, but they are also obviously the starting point from which instructors can organise their (remixed) atomic and sessional material into something incorporating a set of learning outcomes and a package of assessment appropriate to their students. However, our design reflects the expectation that contributions to \gp at each \emph{level} of instruction might be made by \emph{different} people: a domain expert in Spatial Bayes might be the right person to develop an Atom on the concept and its application, but not the right person to develop a module tackling advanced spatial analytic concepts where this is just one approach amongst many. Similarly, given the global diversity of delivery formats, a 10-week term in Britain enables students to cover a very different `volume' of content from a 15-week American semester. \gp recognises and seeks to respond to that diversity at all levels.

\subsection{Tools}\label{tools}

So we are trying to design a system in which sessions and modules are \emph{composed} out of atomic units of instruction that can be, optionally, surrounded by the instructor's own `narrative' or supported by more locally relevant (geo)data. We therefore want to produce a set of teaching materials that are highly portable, easily re-used or edited, and that enable the instructor to select only the elements from which they wish to compose their materials. As intimated above, \gp operationalises this through the Jupyter project and its ability to provide in-browser access to the Python interpreter\footnote{Other languages are also possible using different kernels; there is no reason that \gp couldn't be used to create instructional materials in these languages as well.}, and it also takes advantage of the dominance of Git and GitHub for version control as a means for tracking authorship across many versions and changes.

In theory, thanks to the combination of Jupyter and GitHub is not even necessary for the novice user to have Python installed on their own computer: since all interaction with Python is via the browser, the environment could be hosted on a server halfway round the world. In practice, however, there are few such services and most users simply download and install a free version of Python (\eg Anaconda) that will run on their system. In our field many people are already using this approach: notebooks can be found covering everything from introductory concepts \citep{millington_reades_2017_code} to advanced spatial analysis methods \citep{darribas_gds15}, and combined for both complete courses or workshops \citep{rey_2016_narsc}.

Jupyter notebooks are written as a mix of executable code cells and non-executable text formatted with the widely-used `markdown' syntax. Notebook structure is provided through headers in markdown cells: a `\#' pre-pended to a line of text is generally taken to be the title of the notebook; `\#\#' at the start of a line provides a second level of structure (\ie level 2 headers); `\#\#\#' indicates level 3 headers; etc. For our purposes, what's relevant is that these headers yield a natural hierarchy that corresponds closely to the h1\ldots{}h6 model used by the HTML markup language that lies at the heart of the World Wide Web. This hierarchy allows us to `abstract out' the problem of inferring the \emph{meaning} of different sections of the notebook since the instructor does it for us.

Notebooks are actually stored in a format called JavaScript Object Notation (\textsc{json}) that is not particularly easy for most humans to read, but as it is nonetheless highly-structured we can interact with it programmatically. The extensible nature of the \textsc{json} format also allows us to read and write both data and metadata not only to each notebook, but also to each and every sub-element of a notebook. Since metadata that is not understood by Jupyter is simply ignored, we can add our own fields to provide useful information related to instruction: the amount of time that the notebook's examples and content might take to cover in class; who should be given credit for contributing; any dependencies or requirements for installed libraries\ldots

In short, within the notebook we have two types of structure that can be used not only to provide information that may assist course instructors in developing content, but also to uniquely specify a section, subsection, or sub-subsection of content (code or text) at a fine-grained level. Taken together, this provides the foundation for a mechanism to select content in one notebook and import it into another. This therefore creates an opportunity to remix/mash up content while still being able to add an institutional or course-specific gloss wherever necessary. Each notebook might start with the instructor's contact information or by providing instructions for setting up the computing environment, but then make use of material developed by others for actual instruction. This process may seem quite abstract---and probably quite convoluted as well---but an illustration (Figure \ref{fig:Illustration} \vpageref[above]{fig:Illustration}) may help to clarify why this process is so useful.

\begin{figure}[hbtp]
  \centering
  \includegraphics[width=\textwidth, angle=0]{Illustration.pdf}
  \caption{Illustration of \gp Notebook `Compilation' Process}
  \label{fig:Illustration}
\end{figure}

Finally, in order to manage all of this output we make use of Git and the GitHub web platform to monitor, approve, and roll back alterations to any submitted revisions to \gp Atoms, Sessions, or Modules. This provides additional capabilities such as allowing an instructor to `fix' their course to a particular version of an Atom (a `release' or `tag' in Git) such that development of the Atom by others can continue without the instructor having to worry that the explanation or code upon which they rely will suddenly change! Git also gives us the contributor information that we can propagate into compiled notebooks so that credit is retained, and users and institutions can better-understand the extent to which \gp is a \emph{joint} project relying on the contributions of many teachers and developers.

\subsection{Syntax}\label{syntax}

The final element of \gp is necessarily a syntax for specifying how Sessions and Modules should be compiled from Atoms. We've noted the technical mechanism, as well as the mapping between markdown and \textsc{html} formatting, above but how do we select some mix of code and markdown material in one notebook to be incorporated into another? And how do we do this in a way that is both simple to express and able to resolve ambiguity? Fortunately, such a model already exists and was hinted at in Figure \ref{fig:Illustration}: Cascading Style Sheets (\textsc{css}) uses well-understood `selectors' to specify one or more elements on a web page to which a set of presentational styles should be applied: an `h1' in a style sheet indicates that all level 1 headers should observe the styling rules declared immediately afterwards; while `h1.important' specifies that only a level 1 header of the class `important' should be selected. More powerfully, `h1.important a.super' selects \emph{only} anchors of the class `super' that fall `within' (or `under', if you prefer') a level 1 header of the `important' class.

What is particularly elegant about this approach is that it provides a means for selecting and styling multiple pieces of content in the document in one declaration (where this is desirable) \textit{and} a means for disambiguating content with the same name but in different positions within a document's hierarchical structure (where it is not). We have essentially repurposed \textsc{css} as a means of selecting and importing content from one notebook into another! Unfortunately, the nature of Jupyter notebooks does not allow this to happen dynamically at run-time, but it does allow something similar to happen when an instructor is `compiling' new Sessional and Module content. In short, the instructor writes whatever content they wish but, using syntax similar to the examples below, wherever they want to incorporate contributed content from \gp (or elsewhere) they have only to write an `include' statement to select and import sections of another notebook. This approach also works recursively: a notebook can include content from a notebook that itself includes content from another notebook\ldots

Adopting the language used above: we envision an Atom as a short notebook focussing on a single concept or method (\eg lists, object-oriented design, or spatial autocorrelation); some or all of each Atom can then be selected and imported into a Session, which is itself a notebook; and the sessions can then be selected and managed through a Module, which can be a notebook or a set of notebooks. This process is initiated by the instructor creating a blank text cell in a Jupyter notebook and writing an `include' statement. The statement should be the \textit{only} content in the cell since \gp will be replacing the cell with an unknown number of whole text and code cells from the referenced notebook.

Crucially, `include' statements can be freely intermingled with the instructor's own content, allowing the instructor to `frame' the concepts in a way that suits their teaching style but which saves them having to reinvent the wheel for each class. A Session tackling standardisation could include elements of the relevant Atom from \gp while still allowing the instructor to interject comments, observations, questions, and additional tasks to ground the learning experience in the local context (individual, institutional, etc.). To illustrate this more clearly, an Atom on Python's approach to dealing with lists could be incorporated into a longer Session as follows:

\begin{Verbatim}[fontsize=\small]
@include {
    'nb'     = 'http://geopyter.org/atoms/fundamentals/lists.ipynb',
    'select' = 'h1.Understanding Lists'
}
\end{Verbatim}

The `nb' is the path---local or remote---to a valid Jupyter notebook from which the Session or Module developer wants to import content. The `select' parameter specifies a selector for which the \gp tool will search within the source notebook. All content from that point onwards \emph{up to the next selector at the same level} will the then be copied into the compiled notebook. In example above, if there were a following h1 covering, for example, `List Operators' then this would \emph{not} be included because, from a structural standpoint, it is of equal importance (at the same level in the hierarchy) to `Understanding Lists' but has \emph{not} been selected. Furthermore, any h2 or h3 subsections within the `Understanding Lists' section \textit{would} be included since they are presumed to be providing pedagogical and logical structure to the Understanding Lists section and so should be carried over.

Clearly, an instructor might want to import only part of of a section, or to suppress a subsection falling in the middle of a larger resource. In anticipation of this need, more complex `include' statements with no equivalent in \textsc{css} are also possible:

\begin{Verbatim}[fontsize=\small]
@include {
    'resource' = 'http://geopyter.org/atoms/fundamentals/lists.ipynb',
    'select' = 'h1.Understanding Lists -h3.Lists Example; h1.Using Lists -h3.Another Example'
}
\end{Verbatim}

In this second example, two level 1 section are imported at the same time (the selections are separated with semi-colons) and a level 3 subsection from \emph{within} each of those sections is suppressed using the `-' syntax to indicate that the section should be removed. An additional difference from true \textsc{css} is that we allow spaces in the `selector' because we felt that asking teachers to translate between a natural language header (``Understanding Lists'') and what \textsc{css} would consider a safe header (``Understanding\_Lists'') would detract from ease-of-use. A second departure from the \textsc{css} standard is that selectors are not separated with commas: we wanted to allow for this punctuation to be part of a section heading and felt that semi-colons are rather more rare in that context.

\section{Use Cases}\label{uses}

Two use-cases: an online module of eight sessions, and an in-person module lasting one semester might serve to further illustrate how this approach offers a substantively new way to think about teaching programming material more generally.

\subsection{An Online Module}\label{an-online-module}

There has been increasing interest at universities in Massive Open Online Courses, or \textsc{mooc}s, as vehicles for expanding access to higher education through online delivery. For HE managers, the \textsc{mooc} promises both enhanced revenue---depending on the model this implies either more people paying to take the modules or more paying for a certificate of completion at the end---and enhanced access as people who would be unable to attend the university in person are nonetheless able to pursue a degree. Our own experience suggests substantial student interest in `computational social science' \citep{Lazer2009}, with international students being the most keen on such modules as they are seen to provide marketable skills (programming) for graduates from a discipline (geography) with generally high employment rates (see \citeauthor{rgs2017} \citeyear{rgs2017}).

Whatever the motivation, while expanding access to, and enrolment by, students is probably a `good thing', the process for developing a new online module or `converting' an existing face-to-face module to an online format remains far from easy. Typically, this step requires significant effort and investment by both the instructor and their  institution: all course materials need to be online, recordings/videos of the lecturer presenting a key concept need to be made, the sessions need to be `chunked' so that a 3-hour practical becomes something more manageable, and assessments often need to be substantively redesigned.

\gp significantly reduces the overhead of several of these stages: rather than focussing on tasks such as how to explain or illustrate a particular concept, the instructor can focus the \emph{majority} of their effort on constructing the type of richer, fuller narrative that \textsc{mooc}s seem to require to engage students working remotely. If the instructor isn't there to ask questions and probe student learning on the spot, then the `frame' for each session (as well as for the module as a whole) becomes crucial as it helps to anchor student learning in concrete outcomes and encourages them to reflect upon, and integrate, their progress over the self-guided learning process.

\subsection{In-Class Delivery}\label{in-class-delivery}

In a more traditional in-person format, \gp could be employed differently: since the instructor is able to interactively provide the `frame' or `scaffold' (see relevant discussion of `hypermedia' in \citealp{Azevedo2008}) upon which the learning is built, there is less need for a narrative around each task or weekly session. Here, the instructor might simply import a group of atoms to cover a set of concepts and step through them interactively with the students. The session could wrap up with a targeted project or mini-assessment that requires the students to translate the concepts into a new problem domain or investigate the process in more detail: \emph{e.g.} ``we've seen how we can use pandas and bokeh to analyse demographic groups in London, here's a link to the open data store page for Phoenix with similar data\ldots{}'' or ``Do you think that socioeconomic class is a useful metric? Investigate the source of the data we've just used and see if it squares with your expectations\ldots{}''

In effect, courses built on top of \gp become mash-ups, with the author able to mix and match atoms as needed for the module objectives, time available, and level of student. Both sessions and modules are built via composition: atoms are composed into sessions, sessions into modules. The role of the instructor is then to provide an integrative narrative that, in a sense, explains their choice of components. In both the online and offline contexts we think that this has the potential to free up instructors to concentrate on where they can most effectively `add value', not in figuring out yet \emph{another} way to explain how a list or dictionary works, but in explaining why they matter to a geography, political science, or literature major.

\section{Engagement}\label{engagement}

Because it builds on \textsc{foss} approaches to software development, we expect \gp to benefit from network effects: the more people use it, the more useful it becomes, and the more people use it. In common with many such projects we \emph{do} expect that a relatively small subset of users will contribute the majority of the content; however, unlike a traditional software project there is an important role for \emph{teachers}, not just highly-skilled \emph{developers} and we think that this represents a really exciting opportunity for innovative new approaches to rise to the surface. There may be only a few who can write the Python code to conduct a Geographically-Weighted Regression analysis, but it will be interesting to see how many creative, insightful ways there are to explain it!

\gp also acknowledges the reality of the need for peer and professional recognition by incorporating attribution mechanisms through which faculty can be credited for their work. Contributors can be recognised as the author of content elements through a Digital Object Identifier (\textsc{doi}), and The \gp team would also include section editors responsible for the curation of atoms and for maintenance of official \gp \textsc{doi}s. Instructors who use \gp to compose sessions or courses can also contribute those modules back to the project and would also be recognised as the author of record in a similar fashion. Authorship of \gp educational components, be they atoms, sessions or modules, provides the contributors with peer-evaluated, impactful materials to add to their promotion applications.

Although the open source nature of the project means that we cannot strictly enforce attribution, \gp seeks to make this the easier to do `by default' through the use of Git to track and insert contributory metadata into the compiled notebook. We can use this metadata to append a list of contributors to the end of each notebook, along with any other relevant acknowledgments or copyright notices. To facilitate re-use while protecting contributions from unacknowledged exploitation textual content in \gp is covered by a Creative Commons license; however, to deal with the fact that \gp relies extensively on open source contributions which are incompatible with some \textsc{cc} licenses \citep[see discussion in][]{osswatch2013}, code blocks are licensed under the \textsc{mit} license. The manner in which contributions from authors at different institutions can be combined also `pollutes' the materials in ways that inhibit institutional assertions of ownership over \gp content.

\section{Limitations}\label{limitations}

Although we are profoundly excited both by the possibility of shifting to compositional approaches to content-development, and by the impact that shared materials could have on our teaching, it is highly unlikely that our first attempt to divide up the entire field into discrete units of instruction will be entirely successful. For this, it will be useful to reference documents such as the \textit{Body of Knowledge} \citep{bok2018} and \textit{Subject Benchmark Statement} \citep{QAA2014} for the `why' and `what' of instruction, leaving \gp to deal with the `how'. Moreover, the open source, contributory nature of the project positions us to build \gp on top of the shared understanding of many specialists with a range of ideas about how to break apart---and put back together---the constituent elements of our domain knowledge in ways that speak to different types of students.

For the time being we have also deliberately hobbled \gp in one important way: an \texttt{include} command must be in a cell that does not contain any other text or code. This was done primarily for simplicity: it's a lot easier to look for whole (text) cells that match a target pattern than to have to try to parse long blocks of text or code on the off-chance that something might match; it also avoids any ambiguity as to whether the \texttt{include} is a \gp command or `real' code as might happen, for instance, if \gp were used with a different programming language. Not coincidentally, it is also a good deal easier to know that you are replacing an entire block (the one with an include) with one or more entire blocks that are a mix of text and code, than trying to work out whether the block in which an include was found needs to be `closed out' before the import starts.

\section{Conclusion \& Future Directions}\label{future}

From practical experience, conference presentations, and code we tend to already know who is a good \emph{programmer} or \emph{theorist}, \gp provides a mechanism for discovering who is a good \emph{teacher}. Sometimes these abilities may reside in the same person, but more often we expect that they will not: the strongest developers tend to be people who have been practicing software development for many years and, consequently, may have difficulty communicating their ideas to beginner- or intermediate-level students (or teachers!). For this reason we see the broad-based community-of-practice aspect of \gp as integral at all stages of the project: system enhancement, content development, expanding coverage, and instructional design.

Consequently, \gp has a lot in common---both philosophically and practically---with the Software Carpentry movement \citep{SCF2016}, and although we seek to tackle a slightly narrower set of issues with a more re-usable set of resources, we can take both inspiration and warning from their experience. The benefit, we think, is that while it is possible to design \gp sessions and modules that follow the popular `bootcamp' approach to instruction (though see critique in \citealp{Feldon2017}), we want to enable the \textit{same} content to be employed in a flexible carpentry format as well as a `normal' classroom or \textsc{mooc} as required\ldots Or even to enable the instructor to mash all of those formats together such that they use a `bootcamp' format for the introduction to Unix and the command line, an online-formatted \gp resource for foundational concepts in computer science, and a traditional course format for the (geo)data analysis instruction. All pulled from the same source!

Ultimately, although \gp was developed with teaching needs in mind there is, of course, no reason why it couldn't be put to other uses: in combination with with \texttt{nteract} (\url{https://github.com/nteract/nteract}) it would allow developers or researchers to build fully-fledged applications as scripts assembled from a collection of notebooks; or as an addition to Netflix's notebook ecology to allow for enhanced resource-sharing and standardisation during the development phase before features and interfaces are `fixed' as libraries \citep{Ufford2018}. Nonetheless, our focus for the time being remains the cohort of teachers at all levels tasked with introducing programming material to their students and wondering where to begin. We hope that \gp makes a valuable contribution to this application domain and look forward to working with others to roll out a rich, reusable teaching framework.

\bibliographystyle{apacite}
\bibliography{Geopyter.bib}

\end{document}

### 2019-10-15 Geopyter.tex
Semantic line breaks for diffs

@@
 
-Although \citet{Donoho2017} traces the origins of data science back to \citeauthor{Tukey1962}'s \textit{The future of data analysis} (\citeyear{Tukey1962}), its emergence has led to a flourishing of tools and programmes as universities seek to engage with this nascent---and lucrative---interdisciplinary field.
-One of the most profound practical impacts of this development has been on the ways in which developers share and execute code: although systems for combining code, commentary, and results have been around for some time (\eg R-Markdown) these were intended primarily for replicating research outputs, not for the online teaching of a new technique or writing of an interactive introduction to a software library.
-The `open' ethos and rapid growth of data science, and the need for a lightweight interactive code-sharing platform mean that teachers in many disciplines are playing catch-up on both the methods taught and the formats used.
+Although \citet{Donoho2017} traces the origins of data science back to
+\citeauthor{Tukey1962}'s \textit{The future of data analysis}
+(\citeyear{Tukey1962}), its emergence has led to a flourishing of tools and
+programmes as universities seek to engage with this nascent---and
+lucrative---interdisciplinary field. One of the most profound practical impacts
+of this development has been on the ways in which developers share and execute
+code: although systems for combining code, commentary, and results have been
+around for some time (\eg R-Markdown) these were intended primarily for
+replicating research outputs, not for the online teaching of a new technique or
+writing of an interactive introduction to a software library. The `open' ethos
+and rapid growth of data science, and the need for a lightweight interactive
+code-sharing platform mean that teachers in many disciplines are playing
+catch-up on both the methods taught and the formats used.
 
-As a discipline, geography has been acutely aware of these developments: changes in the volume and extent of geographical data available (\eg \citealp{Graham2013,gonzalez2013big,reades2016}), and in the number and range of methods employed to identify patterns in those data (\eg \citealp{Fan2016,Naik2017,Santibanez2015,Stevens2015,ArribasBel2017}), are allowing us to tackle questions thought unanswerable just a few years ago.
-At the same time, however, there is growing evidence of a `hollowing out' of traditional domain-specific skills: most of what a trained geographer used to do with specialist software can now be done using free online resources (\eg Google's Fusion Tables, MapBox, Carto, ArcOnline), but there remains unmet demand for graduates who can \textit{code} in order to take advantage of the latest (geo)algorithms and techniques \citep{Singleton2014,Singleton2016}.
+As a discipline, geography has been acutely aware of these developments: changes
+in the volume and extent of geographical data available (\eg
+\citealp{Graham2013,gonzalez2013big,reades2016}), and in the number and range of
+methods employed to identify patterns in those data (\eg
+\citealp{Fan2016,Naik2017,Santibanez2015,Stevens2015,ArribasBel2017}), are
+allowing us to tackle questions thought unanswerable just a few years ago. At
+the same time, however, there is growing evidence of a `hollowing out' of
+traditional domain-specific skills: most of what a trained geographer used to do
+with specialist software can now be done using free online resources (\eg
+Google's Fusion Tables, MapBox, Carto, ArcOnline), but there remains unmet
+demand for graduates who can \textit{code} in order to take advantage of the
+latest (geo)algorithms and techniques \citep{Singleton2014,Singleton2016}.
 
@@
 
-Regardless of whether this is a revolution \citep{Wyly2014,Torrens2010} or an evolution \citep{Barnes2013,Barnes2014,ArribasBel2018}, there is no question that teaching students and researchers how to code and how to \textit{think} computationally presents new challenges \citep{Etherington2016,Muller2014}.
-Specifically, because of the way that these changes have unfurled, the majority of resources used in the classroom are being developed from scratch at each institution, and often by newly-appointed lecturers and assistant professors coming from the relatively small number of elite research facilities active in this domain (see discussion in \citealp{esrc2013}).
-Not only has this led to the duplication of effort at multiple sites, but it has also had an enormous impact on the productivity---as commonly measured by institutions---of early career researchers, many of whom are in their first teaching post.
+Regardless of whether this is a revolution \citep{Wyly2014,Torrens2010} or an
+evolution \citep{Barnes2013,Barnes2014,ArribasBel2018}, there is no question
+that teaching students and researchers how to code and how to \textit{think}
+computationally presents new challenges \citep{Etherington2016,Muller2014}.
+Specifically, because of the way that these changes have unfurled, the majority
+of resources used in the classroom are being developed from scratch at each
+institution, and often by newly-appointed lecturers and assistant professors
+coming from the relatively small number of elite research facilities active in
+this domain (see discussion in \citealp{esrc2013}). Not only has this led to the
+duplication of effort at multiple sites, but it has also had an enormous impact
+on the productivity---as commonly measured by institutions---of early career
+researchers, many of whom are in their first teaching post.
 
-The overarching purpose of this software is therefore to address two seemingly contradictory issues: the recognition that no two instructors teach in precisely the same way, and the fact that few instructors have the luxury of both time and tenure to develop compelling new course material in splendid isolation.
-This where the Geographical Python Teaching Resource (\textsc{g}eo\textsc{p}y\textsc{t}e\textsc{r}) comes into play: although developed with geographers in mind, it provides a generic means by which instructors in any discipline can selectively incorporate and re-mix programming and conceptual content from existing Jupyter notebooks while providing their own `gloss' for this content.
-Our approach therefore seeks to develop a richer, open community of developers and teachers who work \textit{together} on new teaching materials, and to support instructors in quickly adapting existing materials to new delivery formats.
+The overarching purpose of this software is therefore to address two seemingly
+contradictory issues: the recognition that no two instructors teach in precisely
+the same way, and the fact that few instructors have the luxury of both time and
+tenure to develop compelling new course material in splendid isolation. This
+where the Geographical Python Teaching Resource
+(\textsc{g}eo\textsc{p}y\textsc{t}e\textsc{r}) comes into play: although
+developed with geographers in mind, it provides a generic means by which
+instructors in any discipline can selectively incorporate and re-mix programming
+and conceptual content from existing Jupyter notebooks while providing their own
+`gloss' for this content. Our approach therefore seeks to develop a richer, open
+community of developers and teachers who work \textit{together} on new teaching
+materials, and to support instructors in quickly adapting existing materials to
+new delivery formats.
 
@@
 
-Like many open source projects, \gp grew out of a need to scratch an itch: in our roles as teachers of (geo)computational concepts and methods we grappled on an annual basis with the demands of developing and updating instructional material with complex interdependencies.
-And, while the fundamental computing and analytic concepts may remain fairly stable, the field is highly-dynamic in terms of both the content and the rate of evolution of new features in contributed code upon which we rely for functionality.
-Currently, every time a widely-used library package is updated a protracted sequence of revisions ensues: some teachers, pressed for time, may resist making use of---or even alerting students to---the new features; others may leap in enthusiastically, only to discover that they hadn't fully thought through the challenges of conflicting dependencies and evolving feature sets.
+Like many open source projects, \gp grew out of a need to scratch an itch: in
+our roles as teachers of (geo)computational concepts and methods we grappled on
+an annual basis with the demands of developing and updating instructional
+material with complex interdependencies. And, while the fundamental computing
+and analytic concepts may remain fairly stable, the field is highly-dynamic in
+terms of both the content and the rate of evolution of new features in
+contributed code upon which we rely for functionality. Currently, every time a
+widely-used library package is updated a protracted sequence of revisions
+ensues: some teachers, pressed for time, may resist making use of---or even
+alerting students to---the new features; others may leap in enthusiastically,
+only to discover that they hadn't fully thought through the challenges of
+conflicting dependencies and evolving feature sets.
 
-We initially explored with others---faculty with whom we regularly collaborate---the potential to `divide and conquer': we would each focus on different topics and then combine these contributions to create a full course.
-Though received with some enthusiasm, this plan did not survive contact with reality: first, when actually faced with the uncertain promise of receiving a full course in return for our efforts, many of us felt more secure developing our own material; second, we had underestimated the value that we each attach to the ability to personalise material to our own style of teaching.
-Since the shared materials were being developed individually according to the individual's own `style', any rationalisation or personalisation would have to be done by the instructor at a later date and without support from the rest of the project.
+We initially explored with others---faculty with whom we regularly
+collaborate---the potential to `divide and conquer': we would each focus on
+different topics and then combine these contributions to create a full course.
+Though received with some enthusiasm, this plan did not survive contact with
+reality: first, when actually faced with the uncertain promise of receiving a
+full course in return for our efforts, many of us felt more secure developing
+our own material; second, we had underestimated the value that we each attach to
+the ability to personalise material to our own style of teaching. Since the
+shared materials were being developed individually according to the individual's
+own `style', any rationalisation or personalisation would have to be done by the
+instructor at a later date and without support from the rest of the project.
 
-Going back to the drawing board, we looked again at our own experiences of successful programming and instructional efforts.
-In particular, we took note of two powerful streams of thought in software and education: the importance attached by the Free and Open Source Software (\textsc{foss}) movement to both peer-recognition \citep{raymond_cathedral_1999} and distributed version control; and the rise of Virtual Learning Environments (\textsc{vle}s) and of Massive Open Online Courses (\textsc{mooc}s) with their use of rich, online learning experiences (\eg \citealp{Trafford2011,Cabiria2012}).
-Reflecting on our initial failure, we decided against trying to supply a single set of polished course materials and instead looked for a way to provide teachers with building blocks from which they could assemble an \textit{individualised} course suited to their institutional and pedagogical needs.
+Going back to the drawing board, we looked again at our own experiences of
+successful programming and instructional efforts. In particular, we took note of
+two powerful streams of thought in software and education: the importance
+attached by the Free and Open Source Software (\textsc{foss}) movement to both
+peer-recognition \citep{raymond_cathedral_1999} and distributed version control;
+and the rise of Virtual Learning Environments (\textsc{vle}s) and of Massive
+Open Online Courses (\textsc{mooc}s) with their use of rich, online learning
+experiences (\eg \citealp{Trafford2011,Cabiria2012}). Reflecting on our initial
+failure, we decided against trying to supply a single set of polished course
+materials and instead looked for a way to provide teachers with building blocks
+from which they could assemble an \textit{individualised} course suited to their
+institutional and pedagogical needs.
 
-We drew inspiration from the \textsc{foss} iPython project \citep{perez_2007_ipython} and its transformative impact on the sharing of integrated code, text, and multimedia resources.
-And we also looked to the Python Spatial Analysis Library (\textsc{p}y\textsc{sal}) project in which contributors take ownership of application areas through managed collaboration instead of a contributory `free-for-all'.
-The subsequent emergence of the Jupyter project \cite{kluyver16} from iPython has led to the rapid growth of a rich ecosystem of tools that support the kinds of interaction needed to make this an effective teaching resource: multiuser-notebooks (jupyterhub), automation of grading (nbgrader), interactive widgets for visualisation (ipywidgets), and packages to facilitate manipulation of the notebook itself (nbformat and notebook).
+We drew inspiration from the \textsc{foss} iPython project
+\citep{perez_2007_ipython} and its transformative impact on the sharing of
+integrated code, text, and multimedia resources. And we also looked to the
+Python Spatial Analysis Library (\textsc{p}y\textsc{sal}) project in which
+contributors take ownership of application areas through managed collaboration
+instead of a contributory `free-for-all'. The subsequent emergence of the
+Jupyter project \cite{kluyver16} from iPython has led to the rapid growth of a
+rich ecosystem of tools that support the kinds of interaction needed to make
+this an effective teaching resource: multiuser-notebooks (jupyterhub),
+automation of grading (nbgrader), interactive widgets for visualisation
+(ipywidgets), and packages to facilitate manipulation of the notebook itself
+(nbformat and notebook).
 
-With these criteria and resources in mind, we began looking for ways to lower the `entry costs' for faculty to both take from, and share into, the corpus of instructional material, as well as for a way to address the customisation/localisation challenge.
-\gp was our solution: a single, open hub of teaching concepts and practical programming tasks that can be flexibly assembled into classes, or entire courses!
-In the \gp approach we recognise that teachers need to be able to develop their own classroom `story' at their own pace (and in conformance with their institution's teaching pattern), while supporting them with a library of up-to-date materials from which they can pick and choose.
-So in much the same way that a computer application is compiled from the contributions of many developers, \gp allows individual classes and entire courses to be `compiled' from the contributions of many skilled teachers and developers.
+With these criteria and resources in mind, we began looking for ways to lower
+the `entry costs' for faculty to both take from, and share into, the corpus of
+instructional material, as well as for a way to address the
+customisation/localisation challenge. \gp was our solution: a single, open hub
+of teaching concepts and practical programming tasks that can be flexibly
+assembled into classes, or entire courses! In the \gp approach we recognise that
+teachers need to be able to develop their own classroom `story' at their own
+pace (and in conformance with their institution's teaching pattern), while
+supporting them with a library of up-to-date materials from which they can pick
+and choose. So in much the same way that a computer application is compiled from
+the contributions of many developers, \gp allows individual classes and entire
+courses to be `compiled' from the contributions of many skilled teachers and
+developers.
 
@@
 
-At least in our experience, the pace of technological and methodological change is such that the gap between what instructors know and what students know is substantial, and in many cases might even be growing.
-The differences within individual student cohorts may be greater still, however, and while students need to be able to begin their learning `journey' at some arbitrary point in the progression from `learning to code' towards `computational social scientist' \citep{Lazer2009}, instructors require the ability to assemble a course that meets students where they \emph{are} rather than where the instructor or institution might \emph{wish} them to be.
+At least in our experience, the pace of technological and methodological change
+is such that the gap between what instructors know and what students know is
+substantial, and in many cases might even be growing. The differences within
+individual student cohorts may be greater still, however, and while students
+need to be able to begin their learning `journey' at some arbitrary point in the
+progression from `learning to code' towards `computational social scientist'
+\citep{Lazer2009}, instructors require the ability to assemble a course that
+meets students where they \emph{are} rather than where the instructor or
+institution might \emph{wish} them to be.
 
-The guiding insight behind \gp is that instruction in computation proceeds from fundamental units of learning typically built around \textit{computing} concepts (variables, lists/arrays, dictionaries/hashes, functions/subroutines, etc.) to fundamental units of learning built around \textit{analytic} concepts (cluster analysis, point patterns, spatial autocorrelation, etc.).
-These units must then be assembled in a way that speaks to the student cohort and its background; we mean this in two ways: first, that the examples used must speak to budding geographers, political scientists, historians, etc.; and second, that it should be possible to develop courses for different cohorts without having to start over from scratch.
-In other words, how can we enable teachers to reuse many of the building blocks employed in an advanced course for Masters students in an introductory undergraduate course?
+The guiding insight behind \gp is that instruction in computation proceeds from
+fundamental units of learning typically built around \textit{computing} concepts
+(variables, lists/arrays, dictionaries/hashes, functions/subroutines, etc.) to
+fundamental units of learning built around \textit{analytic} concepts (cluster
+analysis, point patterns, spatial autocorrelation, etc.). These units must then
+be assembled in a way that speaks to the student cohort and its background; we
+mean this in two ways: first, that the examples used must speak to budding
+geographers, political scientists, historians, etc.; and second, that it should
+be possible to develop courses for different cohorts without having to start
+over from scratch. In other words, how can we enable teachers to reuse many of
+the building blocks employed in an advanced course for Masters students in an
+introductory undergraduate course?
 
@@
 
-From these constraints it was clear that our system needed to support a `bottom-up' approach to instructional design.
-So although the development of a course or class could---and probably should---start with a clear set of learning aims and outcomes, at a certain point the instructor will be searching for a examples with which to teach a particular concept: What \emph{is} a list or dictionary?
-What is \textit{k}-means clustering?
-We settled on the term `Atoms' to refer to these basic instructional units and, much like entries in the atomic table, we felt that they could be grouped together into sets of related concepts: the fundamentals of programming, point pattern analysis, machine learning, etc.
+From these constraints it was clear that our system needed to support a
+`bottom-up' approach to instructional design. So although the development of a
+course or class could---and probably should---start with a clear set of learning
+aims and outcomes, at a certain point the instructor will be searching for a
+examples with which to teach a particular concept: What \emph{is} a list or
+dictionary? What is \textit{k}-means clustering? We settled on the term `Atoms'
+to refer to these basic instructional units and, much like entries in the atomic
+table, we felt that they could be grouped together into sets of related
+concepts: the fundamentals of programming, point pattern analysis, machine
+learning, etc.
 
-Each atom would employ \emph{domain-specific} illustrative examples and code so as to anchor learning in relevant problems and applications.
-An atom on arrays written for geographers could start by showing how these simple data structures can be used to hold data about cities (\eg name, country, population); but a subsequent unit could build on this with an an illustration of how nested data structures allow us to add city location using nested coordinate pairs (\eg latitude and longitude within an array-of-arrays).
-This is why managed collaboration around sets of related concepts is so important: quite aside from the fact that no \textsc{foss} project succeeds without some level of executive control, it's critical that \textit{someone} is responsible for ensuring a degree of consistency in how the units fit together within a set of closely-related topics.
+Each atom would employ \emph{domain-specific} illustrative examples and code so
+as to anchor learning in relevant problems and applications. An atom on arrays
+written for geographers could start by showing how these simple data structures
+can be used to hold data about cities (\eg name, country, population); but a
+subsequent unit could build on this with an an illustration of how nested data
+structures allow us to add city location using nested coordinate pairs (\eg
+latitude and longitude within an array-of-arrays). This is why managed
+collaboration around sets of related concepts is so important: quite aside from
+the fact that no \textsc{foss} project succeeds without some level of executive
+control, it's critical that \textit{someone} is responsible for ensuring a
+degree of consistency in how the units fit together within a set of
+closely-related topics.
 
-The purpose of the bottom-up approach is that these units can then be flexibly assembled into Sessions: from the \textit{same ingredients} (\ie atoms) the instructor could create two quite different modules---one for delivery in person, one for delivery online---by organising the elements in different ways.
-Quite simply, we don't want to have to rewrite material for each format: lecture/practical, `flipped' classroom, or \textsc{mooc}.
-Furthermore, sessions designed for experts (\eg those pursuing continuing online education) might be able to `move' students through many more atoms of instruction in a single session than a similarly laid out (in terms of time) session designed for first-time programmers in an undergraduate programme.
-So sessions need to be able to \emph{incorporate} atoms in a way that minimises the level of effort involved in finding the `best' way to, for example, explain the concept of recursion while maximising the ability of the instructor to relate this concept to the students' practical experience (\eg by providing a `context' that is anchored in a locally relevant `story' or data).
+The purpose of the bottom-up approach is that these units can then be flexibly
+assembled into Sessions: from the \textit{same ingredients} (\ie atoms) the
+instructor could create two quite different modules---one for delivery in
+person, one for delivery online---by organising the elements in different ways.
+Quite simply, we don't want to have to rewrite material for each format:
+lecture/practical, `flipped' classroom, or \textsc{mooc}. Furthermore, sessions
+designed for experts (\eg those pursuing continuing online education) might be
+able to `move' students through many more atoms of instruction in a single
+session than a similarly laid out (in terms of time) session designed for
+first-time programmers in an undergraduate programme. So sessions need to be
+able to \emph{incorporate} atoms in a way that minimises the level of effort
+involved in finding the `best' way to, for example, explain the concept of
+recursion while maximising the ability of the instructor to relate this concept
+to the students' practical experience (\eg by providing a `context' that is
+anchored in a locally relevant `story' or data).
 
-Naturally, Sessions can then be grouped into Learning Modules (modules) that offer a coherent instructional programme over a period of weeks or months.
-Modules represent the highest level of abstraction in the proposed system, but they are also obviously the starting point from which instructors can organise their (remixed) atomic and sessional material into something incorporating a set of learning outcomes and a package of assessment appropriate to their students.
-However, our design reflects the expectation that contributions to \gp at each \emph{level} of instruction might be made by \emph{different} people: a domain expert in Spatial Bayes might be the right person to develop an Atom on the concept and its application, but not the right person to develop a module tackling advanced spatial analytic concepts where this is just one approach amongst many.
-Similarly, given the global diversity of delivery formats, a 10-week term in Britain enables students to cover a very different `volume' of content from a 15-week American semester.
-\gp recognises and seeks to respond to that diversity at all levels.
+Naturally, Sessions can then be grouped into Learning Modules (modules) that
+offer a coherent instructional programme over a period of weeks or months.
+Modules represent the highest level of abstraction in the proposed system, but
+they are also obviously the starting point from which instructors can organise
+their (remixed) atomic and sessional material into something incorporating a set
+of learning outcomes and a package of assessment appropriate to their students.
+However, our design reflects the expectation that contributions to \gp at each
+\emph{level} of instruction might be made by \emph{different} people: a domain
+expert in Spatial Bayes might be the right person to develop an Atom on the
+concept and its application, but not the right person to develop a module
+tackling advanced spatial analytic concepts where this is just one approach
+amongst many. Similarly, given the global diversity of delivery formats, a
+10-week term in Britain enables students to cover a very different `volume' of
+content from a 15-week American semester. \gp recognises and seeks to respond to
+that diversity at all levels.
 
@@
 
-So we are trying to design a system in which sessions and modules are \emph{composed} out of atomic units of instruction that can be, optionally, surrounded by the instructor's own `narrative' or supported by more locally relevant (geo)data.
-We therefore want to produce a set of teaching materials that are highly portable, easily re-used or edited, and that enable the instructor to select only the elements from which they wish to compose their materials.
-As intimated above, \gp operationalises this through the Jupyter project and its ability to provide in-browser access to the Python interpreter\footnote{Other languages are also possible using different kernels; there is no reason that \gp couldn't be used to create instructional materials in these languages as well.}, and it also takes advantage of the dominance of Git and GitHub for version control as a means for tracking authorship across many versions and changes.
+So we are trying to design a system in which sessions and modules are
+\emph{composed} out of atomic units of instruction that can be, optionally,
+surrounded by the instructor's own `narrative' or supported by more locally
+relevant (geo)data. We therefore want to produce a set of teaching materials
+that are highly portable, easily re-used or edited, and that enable the
+instructor to select only the elements from which they wish to compose their
+materials. As intimated above, \gp operationalises this through the Jupyter
+project and its ability to provide in-browser access to the Python
+interpreter\footnote{Other languages are also possible using different kernels;
+there is no reason that \gp couldn't be used to create instructional materials
+in these languages as well.}, and it also takes advantage of the dominance of
+Git and GitHub for version control as a means for tracking authorship across
+many versions and changes.
 
-In theory, thanks to the combination of Jupyter and GitHub is not even necessary for the novice user to have Python installed on their own computer: since all interaction with Python is via the browser, the environment could be hosted on a server halfway round the world.
-In practice, however, there are few such services and most users simply download and install a free version of Python (\eg Anaconda) that will run on their system.
-In our field many people are already using this approach: notebooks can be found covering everything from introductory concepts \citep{millington_reades_2017_code} to advanced spatial analysis methods \citep{darribas_gds15}, and combined for both complete courses or workshops \citep{rey_2016_narsc}.
+In theory, thanks to the combination of Jupyter and GitHub is not even necessary
+for the novice user to have Python installed on their own computer: since all
+interaction with Python is via the browser, the environment could be hosted on a
+server halfway round the world. In practice, however, there are few such
+services and most users simply download and install a free version of Python
+(\eg Anaconda) that will run on their system. In our field many people are
+already using this approach: notebooks can be found covering everything from
+introductory concepts \citep{millington_reades_2017_code} to advanced spatial
+analysis methods \citep{darribas_gds15}, and combined for both complete courses
+or workshops \citep{rey_2016_narsc}.
 
-Jupyter notebooks are written as a mix of executable code cells and non-executable text formatted with the widely-used `markdown' syntax.
-Notebook structure is provided through headers in markdown cells: a `\#' pre-pended to a line of text is generally taken to be the title of the notebook; `\#\#' at the start of a line provides a second level of structure (\ie level 2 headers); `\#\#\#' indicates level 3 headers; etc.
-For our purposes, what's relevant is that these headers yield a natural hierarchy that corresponds closely to the h1\ldots{}h6 model used by the HTML markup language that lies at the heart of the World Wide Web.
-This hierarchy allows us to `abstract out' the problem of inferring the \emph{meaning} of different sections of the notebook since the instructor does it for us.
+Jupyter notebooks are written as a mix of executable code cells and
+non-executable text formatted with the widely-used `markdown' syntax. Notebook
+structure is provided through headers in markdown cells: a `\#' pre-pended to a
+line of text is generally taken to be the title of the notebook; `\#\#' at the
+start of a line provides a second level of structure (\ie level 2 headers);
+`\#\#\#' indicates level 3 headers; etc. For our purposes, what's relevant is
+that these headers yield a natural hierarchy that corresponds closely to the
+h1\ldots{}h6 model used by the HTML markup language that lies at the heart of
+the World Wide Web. This hierarchy allows us to `abstract out' the problem of
+inferring the \emph{meaning} of different sections of the notebook since the
+instructor does it for us.
 
-Notebooks are actually stored in a format called JavaScript Object Notation (\textsc{json}) that is not particularly easy for most humans to read, but as it is nonetheless highly-structured we can interact with it programmatically.
-The extensible nature of the \textsc{json} format also allows us to read and write both data and metadata not only to each notebook, but also to each and every sub-element of a notebook.
-Since metadata that is not understood by Jupyter is simply ignored, we can add our own fields to provide useful information related to instruction: the amount of time that the notebook's examples and content might take to cover in class; who should be given credit for contributing; any dependencies or requirements for installed libraries\ldots
+Notebooks are actually stored in a format called JavaScript Object Notation
+(\textsc{json}) that is not particularly easy for most humans to read, but as it
+is nonetheless highly-structured we can interact with it programmatically. The
+extensible nature of the \textsc{json} format also allows us to read and write
+both data and metadata not only to each notebook, but also to each and every
+sub-element of a notebook. Since metadata that is not understood by Jupyter is
+simply ignored, we can add our own fields to provide useful information related
+to instruction: the amount of time that the notebook's examples and content
+might take to cover in class; who should be given credit for contributing; any
+dependencies or requirements for installed libraries\ldots
 
-In short, within the notebook we have two types of structure that can be used not only to provide information that may assist course instructors in developing content, but also to uniquely specify a section, subsection, or sub-subsection of content (code or text) at a fine-grained level.
-Taken together, this provides the foundation for a mechanism to select content in one notebook and import it into another.
-This therefore creates an opportunity to remix/mash up content while still being able to add an institutional or course-specific gloss wherever necessary.
-Each notebook might start with the instructor's contact information or by providing instructions for setting up the computing environment, but then make use of material developed by others for actual instruction.
-This process may seem quite abstract---and probably quite convoluted as well---but an illustration (Figure \ref{fig:Illustration} \vpageref[above]{fig:Illustration}) may help to clarify why this process is so useful.
+In short, within the notebook we have two types of structure that can be used
+not only to provide information that may assist course instructors in developing
+content, but also to uniquely specify a section, subsection, or sub-subsection
+of content (code or text) at a fine-grained level. Taken together, this provides
+the foundation for a mechanism to select content in one notebook and import it
+into another. This therefore creates an opportunity to remix/mash up content
+while still being able to add an institutional or course-specific gloss wherever
+necessary. Each notebook might start with the instructor's contact information
+or by providing instructions for setting up the computing environment, but then
+make use of material developed by others for actual instruction. This process
+may seem quite abstract---and probably quite convoluted as well---but an
+illustration (Figure \ref{fig:Illustration} \vpageref[above]{fig:Illustration})
+may help to clarify why this process is so useful.
 
@@
 
-Finally, in order to manage all of this output we make use of Git and the GitHub web platform to monitor, approve, and roll back alterations to any submitted revisions to \gp Atoms, Sessions, or Modules.
-This provides additional capabilities such as allowing an instructor to `fix' their course to a particular version of an Atom (a `release' or `tag' in Git) such that development of the Atom by others can continue without the instructor having to worry that the explanation or code upon which they rely will suddenly change!
-Git also gives us the contributor information that we can propagate into compiled notebooks so that credit is retained, and users and institutions can better-understand the extent to which \gp is a \emph{joint} project relying on the contributions of many teachers and developers.
+Finally, in order to manage all of this output we make use of Git and the GitHub
+web platform to monitor, approve, and roll back alterations to any submitted
+revisions to \gp Atoms, Sessions, or Modules. This provides additional
+capabilities such as allowing an instructor to `fix' their course to a
+particular version of an Atom (a `release' or `tag' in Git) such that
+development of the Atom by others can continue without the instructor having to
+worry that the explanation or code upon which they rely will suddenly change!
+Git also gives us the contributor information that we can propagate into
+compiled notebooks so that credit is retained, and users and institutions can
+better-understand the extent to which \gp is a \emph{joint} project relying on
+the contributions of many teachers and developers.
 
@@
 
-The final element of \gp is necessarily a syntax for specifying how Sessions and Modules should be compiled from Atoms.
-We've noted the technical mechanism, as well as the mapping between markdown and \textsc{html} formatting, above but how do we select some mix of code and markdown material in one notebook to be incorporated into another?
-And how do we do this in a way that is both simple to express and able to resolve ambiguity?
-Fortunately, such a model already exists and was hinted at in Figure \ref{fig:Illustration}: Cascading Style Sheets (\textsc{css}) uses well-understood `selectors' to specify one or more elements on a web page to which a set of presentational styles should be applied: an `h1' in a style sheet indicates that all level 1 headers should observe the styling rules declared immediately afterwards; while `h1.important' specifies that only a level 1 header of the class `important' should be selected.
-More powerfully, `h1.important a.super' selects \emph{only} anchors of the class `super' that fall `within' (or `under', if you prefer') a level 1 header of the `important' class.
+The final element of \gp is necessarily a syntax for specifying how Sessions and
+Modules should be compiled from Atoms. We've noted the technical mechanism, as
+well as the mapping between markdown and \textsc{html} formatting, above but how
+do we select some mix of code and markdown material in one notebook to be
+incorporated into another? And how do we do this in a way that is both simple to
+express and able to resolve ambiguity? Fortunately, such a model already exists
+and was hinted at in Figure \ref{fig:Illustration}: Cascading Style Sheets
+(\textsc{css}) uses well-understood `selectors' to specify one or more elements
+on a web page to which a set of presentational styles should be applied: an `h1'
+in a style sheet indicates that all level 1 headers should observe the styling
+rules declared immediately afterwards; while `h1.important' specifies that only
+a level 1 header of the class `important' should be selected. More powerfully,
+`h1.important a.super' selects \emph{only} anchors of the class `super' that
+fall `within' (or `under', if you prefer') a level 1 header of the `important'
+class.
 
-What is particularly elegant about this approach is that it provides a means for selecting and styling multiple pieces of content in the document in one declaration (where this is desirable) \textit{and} a means for disambiguating content with the same name but in different positions within a document's hierarchical structure (where it is not).
-We have essentially repurposed \textsc{css} as a means of selecting and importing content from one notebook into another!
-Unfortunately, the nature of Jupyter notebooks does not allow this to happen dynamically at run-time, but it does allow something similar to happen when an instructor is `compiling' new Sessional and Module content.
-In short, the instructor writes whatever content they wish but, using syntax similar to the examples below, wherever they want to incorporate contributed content from \gp (or elsewhere) they have only to write an `include' statement to select and import sections of another notebook.
-This approach also works recursively: a notebook can include content from a notebook that itself includes content from another notebook\ldots
+What is particularly elegant about this approach is that it provides a means for
+selecting and styling multiple pieces of content in the document in one
+declaration (where this is desirable) \textit{and} a means for disambiguating
+content with the same name but in different positions within a document's
+hierarchical structure (where it is not). We have essentially repurposed
+\textsc{css} as a means of selecting and importing content from one notebook
+into another! Unfortunately, the nature of Jupyter notebooks does not allow this
+to happen dynamically at run-time, but it does allow something similar to happen
+when an instructor is `compiling' new Sessional and Module content. In short,
+the instructor writes whatever content they wish but, using syntax similar to
+the examples below, wherever they want to incorporate contributed content from
+\gp (or elsewhere) they have only to write an `include' statement to select and
+import sections of another notebook. This approach also works recursively: a
+notebook can include content from a notebook that itself includes content from
+another notebook\ldots
 
-Adopting the language used above: we envision an Atom as a short notebook focussing on a single concept or method (\eg lists, object-oriented design, or spatial autocorrelation); some or all of each Atom can then be selected and imported into a Session, which is itself a notebook; and the sessions can then be selected and managed through a Module, which can be a notebook or a set of notebooks.
-This process is initiated by the instructor creating a blank text cell in a Jupyter notebook and writing an `include' statement.
-The statement should be the \textit{only} content in the cell since \gp will be replacing the cell with an unknown number of whole text and code cells from the referenced notebook.
+Adopting the language used above: we envision an Atom as a short notebook
+focussing on a single concept or method (\eg lists, object-oriented design, or
+spatial autocorrelation); some or all of each Atom can then be selected and
+imported into a Session, which is itself a notebook; and the sessions can then
+be selected and managed through a Module, which can be a notebook or a set of
+notebooks. This process is initiated by the instructor creating a blank text
+cell in a Jupyter notebook and writing an `include' statement. The statement
+should be the \textit{only} content in the cell since \gp will be replacing the
+cell with an unknown number of whole text and code cells from the referenced
+notebook.
 
-Crucially, `include' statements can be freely intermingled with the instructor's own content, allowing the instructor to `frame' the concepts in a way that suits their teaching style but which saves them having to reinvent the wheel for each class.
-A Session tackling standardisation could include elements of the relevant Atom from \gp while still allowing the instructor to interject comments, observations, questions, and additional tasks to ground the learning experience in the local context (individual, institutional, etc.).
-To illustrate this more clearly, an Atom on Python's approach to dealing with lists could be incorporated into a longer Session as follows:
+Crucially, `include' statements can be freely intermingled with the instructor's
+own content, allowing the instructor to `frame' the concepts in a way that suits
+their teaching style but which saves them having to reinvent the wheel for each
+class. A Session tackling standardisation could include elements of the relevant
+Atom from \gp while still allowing the instructor to interject comments,
+observations, questions, and additional tasks to ground the learning experience
+in the local context (individual, institutional, etc.). To illustrate this more
+clearly, an Atom on Python's approach to dealing with lists could be
+incorporated into a longer Session as follows:
 
@@
 
-The `nb' is the path---local or remote---to a valid Jupyter notebook from which the Session or Module developer wants to import content.
-The `select' parameter specifies a selector for which the \gp tool will search within the source notebook.
-All content from that point onwards \emph{up to the next selector at the same level} will the then be copied into the compiled notebook.
-In example above, if there were a following h1 covering, for example, `List Operators' then this would \emph{not} be included because, from a structural standpoint, it is of equal importance (at the same level in the hierarchy) to `Understanding Lists' but has \emph{not} been selected.
-Furthermore, any h2 or h3 subsections within the `Understanding Lists' section \textit{would} be included since they are presumed to be providing pedagogical and logical structure to the Understanding Lists section and so should be carried over.
+The `nb' is the path---local or remote---to a valid Jupyter notebook from which
+the Session or Module developer wants to import content. The `select' parameter
+specifies a selector for which the \gp tool will search within the source
+notebook. All content from that point onwards \emph{up to the next selector at
+the same level} will the then be copied into the compiled notebook. In example
+above, if there were a following h1 covering, for example, `List Operators' then
+this would \emph{not} be included because, from a structural standpoint, it is
+of equal importance (at the same level in the hierarchy) to `Understanding
+Lists' but has \emph{not} been selected. Furthermore, any h2 or h3 subsections
+within the `Understanding Lists' section \textit{would} be included since they
+are presumed to be providing pedagogical and logical structure to the
+Understanding Lists section and so should be carried over.
 
-Clearly, an instructor might want to import only part of of a section, or to suppress a subsection falling in the middle of a larger resource.
-In anticipation of this need, more complex `include' statements with no equivalent in \textsc{css} are also possible:
+Clearly, an instructor might want to import only part of of a section, or to
+suppress a subsection falling in the middle of a larger resource. In
+anticipation of this need, more complex `include' statements with no equivalent
+in \textsc{css} are also possible:
 
@@
 
-In this second example, two level 1 section are imported at the same time (the selections are separated with semi-colons) and a level 3 subsection from \emph{within} each of those sections is suppressed using the `-' syntax to indicate that the section should be removed.
-An additional difference from true \textsc{css} is that we allow spaces in the `selector' because we felt that asking teachers to translate between a natural language header (``Understanding Lists'') and what \textsc{css} would consider a safe header (``Understanding\_Lists'') would detract from ease-of-use.
-A second departure from the \textsc{css} standard is that selectors are not separated with commas: we wanted to allow for this punctuation to be part of a section heading and felt that semi-colons are rather more rare in that context.
+In this second example, two level 1 section are imported at the same time (the
+selections are separated with semi-colons) and a level 3 subsection from
+\emph{within} each of those sections is suppressed using the `-' syntax to
+indicate that the section should be removed. An additional difference from true
+\textsc{css} is that we allow spaces in the `selector' because we felt that
+asking teachers to translate between a natural language header (``Understanding
+Lists'') and what \textsc{css} would consider a safe header
+(``Understanding\_Lists'') would detract from ease-of-use. A second departure
+from the \textsc{css} standard is that selectors are not separated with commas:
+we wanted to allow for this punctuation to be part of a section heading and felt
+that semi-colons are rather more rare in that context.
 
@@
 
-Two use-cases: an online module of eight sessions, and an in-person module lasting one semester might serve to further illustrate how this approach offers a substantively new way to think about teaching programming material more generally.
+Two use-cases: an online module of eight sessions, and an in-person module
+lasting one semester might serve to further illustrate how this approach offers
+a substantively new way to think about teaching programming material more
+generally.
 
@@
 
-There has been increasing interest at universities in Massive Open Online Courses, or \textsc{mooc}s, as vehicles for expanding access to higher education through online delivery.
-For HE managers, the \textsc{mooc} promises both enhanced revenue---depending on the model this implies either more people paying to take the modules or more paying for a certificate of completion at the end---and enhanced access as people who would be unable to attend the university in person are nonetheless able to pursue a degree.
-Our own experience suggests substantial student interest in `computational social science' \citep{Lazer2009}, with international students being the most keen on such modules as they are seen to provide marketable skills (programming) for graduates from a discipline (geography) with generally high employment rates (see \citeauthor{rgs2017} \citeyear{rgs2017}).
+There has been increasing interest at universities in Massive Open Online
+Courses, or \textsc{mooc}s, as vehicles for expanding access to higher education
+through online delivery. For HE managers, the \textsc{mooc} promises both
+enhanced revenue---depending on the model this implies either more people paying
+to take the modules or more paying for a certificate of completion at the
+end---and enhanced access as people who would be unable to attend the university
+in person are nonetheless able to pursue a degree. Our own experience suggests
+substantial student interest in `computational social science'
+\citep{Lazer2009}, with international students being the most keen on such
+modules as they are seen to provide marketable skills (programming) for
+graduates from a discipline (geography) with generally high employment rates
+(see \citeauthor{rgs2017} \citeyear{rgs2017}).
 
-Whatever the motivation, while expanding access to, and enrolment by, students is probably a `good thing', the process for developing a new online module or `converting' an existing face-to-face module to an online format remains far from easy.
-Typically, this step requires significant effort and investment by both the instructor and their  institution: all course materials need to be online, recordings/videos of the lecturer presenting a key concept need to be made, the sessions need to be `chunked' so that a 3-hour practical becomes something more manageable, and assessments often need to be substantively redesigned.
+Whatever the motivation, while expanding access to, and enrolment by, students
+is probably a `good thing', the process for developing a new online module or
+`converting' an existing face-to-face module to an online format remains far
+from easy. Typically, this step requires significant effort and investment by
+both the instructor and their institution: all course materials need to be
+online, recordings/videos of the lecturer presenting a key concept need to be
+made, the sessions need to be `chunked' so that a 3-hour practical becomes
+something more manageable, and assessments often need to be substantively
+redesigned.
 
-\gp significantly reduces the overhead of several of these stages: rather than focussing on tasks such as how to explain or illustrate a particular concept, the instructor can focus the \emph{majority} of their effort on constructing the type of richer, fuller narrative that \textsc{mooc}s seem to require to engage students working remotely.
-If the instructor isn't there to ask questions and probe student learning on the spot, then the `frame' for each session (as well as for the module as a whole) becomes crucial as it helps to anchor student learning in concrete outcomes and encourages them to reflect upon, and integrate, their progress over the self-guided learning process.
+\gp significantly reduces the overhead of several of these stages: rather than
+focussing on tasks such as how to explain or illustrate a particular concept,
+the instructor can focus the \emph{majority} of their effort on constructing the
+type of richer, fuller narrative that \textsc{mooc}s seem to require to engage
+students working remotely. If the instructor isn't there to ask questions and
+probe student learning on the spot, then the `frame' for each session (as well
+as for the module as a whole) becomes crucial as it helps to anchor student
+learning in concrete outcomes and encourages them to reflect upon, and
+integrate, their progress over the self-guided learning process.
 
@@
 
-In a more traditional in-person format, \gp could be employed differently: since the instructor is able to interactively provide the `frame' or `scaffold' (see relevant discussion of `hypermedia' in \citealp{Azevedo2008}) upon which the learning is built, there is less need for a narrative around each task or weekly session.
-Here, the instructor might simply import a group of atoms to cover a set of concepts and step through them interactively with the students.
-The session could wrap up with a targeted project or mini-assessment that requires the students to translate the concepts into a new problem domain or investigate the process in more detail: \emph{e.g.} ``we've seen how we can use pandas and bokeh to analyse demographic groups in London, here's a link to the open data store page for Phoenix with similar data\ldots{}'' or ``Do you think that socioeconomic class is a useful metric?
-Investigate the source of the data we've just used and see if it squares with your expectations\ldots{}''
+In a more traditional in-person format, \gp could be employed differently: since
+the instructor is able to interactively provide the `frame' or `scaffold' (see
+relevant discussion of `hypermedia' in \citealp{Azevedo2008}) upon which the
+learning is built, there is less need for a narrative around each task or weekly
+session. Here, the instructor might simply import a group of atoms to cover a
+set of concepts and step through them interactively with the students. The
+session could wrap up with a targeted project or mini-assessment that requires
+the students to translate the concepts into a new problem domain or investigate
+the process in more detail: \emph{e.g.} ``we've seen how we can use pandas and
+bokeh to analyse demographic groups in London, here's a link to the open data
+store page for Phoenix with similar data\ldots{}'' or ``Do you think that
+socioeconomic class is a useful metric? Investigate the source of the data we've
+just used and see if it squares with your expectations\ldots{}''
 
-In effect, courses built on top of \gp become mash-ups, with the author able to mix and match atoms as needed for the module objectives, time available, and level of student.
-Both sessions and modules are built via composition: atoms are composed into sessions, sessions into modules.
-The role of the instructor is then to provide an integrative narrative that, in a sense, explains their choice of components.
-In both the online and offline contexts we think that this has the potential to free up instructors to concentrate on where they can most effectively `add value', not in figuring out yet \emph{another} way to explain how a list or dictionary works, but in explaining why they matter to a geography, political science, or literature major.
+In effect, courses built on top of \gp become mash-ups, with the author able to
+mix and match atoms as needed for the module objectives, time available, and
+level of student. Both sessions and modules are built via composition: atoms are
+composed into sessions, sessions into modules. The role of the instructor is
+then to provide an integrative narrative that, in a sense, explains their choice
+of components. In both the online and offline contexts we think that this has
+the potential to free up instructors to concentrate on where they can most
+effectively `add value', not in figuring out yet \emph{another} way to explain
+how a list or dictionary works, but in explaining why they matter to a
+geography, political science, or literature major.
 
@@
 
-Because it builds on \textsc{foss} approaches to software development, we expect \gp to benefit from network effects: the more people use it, the more useful it becomes, and the more people use it.
-In common with many such projects we \emph{do} expect that a relatively small subset of users will contribute the majority of the content; however, unlike a traditional software project there is an important role for \emph{teachers}, not just highly-skilled \emph{developers} and we think that this represents a really exciting opportunity for innovative new approaches to rise to the surface.
-There may be only a few who can write the Python code to conduct a Geographically-Weighted Regression analysis, but it will be interesting to see how many creative, insightful ways there are to explain it!
+Because it builds on \textsc{foss} approaches to software development, we expect
+\gp to benefit from network effects: the more people use it, the more useful it
+becomes, and the more people use it. In common with many such projects we
+\emph{do} expect that a relatively small subset of users will contribute the
+majority of the content; however, unlike a traditional software project there is
+an important role for \emph{teachers}, not just highly-skilled \emph{developers}
+and we think that this represents a really exciting opportunity for innovative
+new approaches to rise to the surface. There may be only a few who can write the
+Python code to conduct a Geographically-Weighted Regression analysis, but it
+will be interesting to see how many creative, insightful ways there are to
+explain it!
 
-\gp also acknowledges the reality of the need for peer and professional recognition by incorporating attribution mechanisms through which faculty can be credited for their work.
-Contributors can be recognised as the author of content elements through a Digital Object Identifier (\textsc{doi}), and The \gp team would also include section editors responsible for the curation of atoms and for maintenance of official \gp \textsc{doi}s.
-Instructors who use \gp to compose sessions or courses can also contribute those modules back to the project and would also be recognised as the author of record in a similar fashion.
-Authorship of \gp educational components, be they atoms, sessions or modules, provides the contributors with peer-evaluated, impactful materials to add to their promotion applications.
+\gp also acknowledges the reality of the need for peer and professional
+recognition by incorporating attribution mechanisms through which faculty can be
+credited for their work. Contributors can be recognised as the author of content
+elements through a Digital Object Identifier (\textsc{doi}), and The \gp team
+would also include section editors responsible for the curation of atoms and for
+maintenance of official \gp \textsc{doi}s. Instructors who use \gp to compose
+sessions or courses can also contribute those modules back to the project and
+would also be recognised as the author of record in a similar fashion.
+Authorship of \gp educational components, be they atoms, sessions or modules,
+provides the contributors with peer-evaluated, impactful materials to add to
+their promotion applications.
 
-Although the open source nature of the project means that we cannot strictly enforce attribution, \gp seeks to make this the easier to do `by default' through the use of Git to track and insert contributory metadata into the compiled notebook.
-We can use this metadata to append a list of contributors to the end of each notebook, along with any other relevant acknowledgments or copyright notices.
-To facilitate re-use while protecting contributions from unacknowledged exploitation textual content in \gp is covered by a Creative Commons license; however, to deal with the fact that \gp relies extensively on open source contributions which are incompatible with some \textsc{cc} licenses \citep[see discussion in][]{osswatch2013}, code blocks are licensed under the \textsc{mit} license.
-The manner in which contributions from authors at different institutions can be combined also `pollutes' the materials in ways that inhibit institutional assertions of ownership over \gp content.
+Although the open source nature of the project means that we cannot strictly
+enforce attribution, \gp seeks to make this the easier to do `by default'
+through the use of Git to track and insert contributory metadata into the
+compiled notebook. We can use this metadata to append a list of contributors to
+the end of each notebook, along with any other relevant acknowledgments or
+copyright notices. To facilitate re-use while protecting contributions from
+unacknowledged exploitation textual content in \gp is covered by a Creative
+Commons license; however, to deal with the fact that \gp relies extensively on
+open source contributions which are incompatible with some \textsc{cc} licenses
+\citep[see discussion in][]{osswatch2013}, code blocks are licensed under the
+\textsc{mit} license. The manner in which contributions from authors at
+different institutions can be combined also `pollutes' the materials in ways
+that inhibit institutional assertions of ownership over \gp content.
 
@@
 
-Although we are profoundly excited both by the possibility of shifting to compositional approaches to content-development, and by the impact that shared materials could have on our teaching, it is highly unlikely that our first attempt to divide up the entire field into discrete units of instruction will be entirely successful.
-For this, it will be useful to reference documents such as the \textit{Body of Knowledge} \citep{bok2018} and \textit{Subject Benchmark Statement} \citep{QAA2014} for the `why' and `what' of instruction, leaving \gp to deal with the `how'.
-Moreover, the open source, contributory nature of the project positions us to build \gp on top of the shared understanding of many specialists with a range of ideas about how to break apart---and put back together---the constituent elements of our domain knowledge in ways that speak to different types of students.
+Although we are profoundly excited both by the possibility of shifting to
+compositional approaches to content-development, and by the impact that shared
+materials could have on our teaching, it is highly unlikely that our first
+attempt to divide up the entire field into discrete units of instruction will be
+entirely successful. For this, it will be useful to reference documents such as
+the \textit{Body of Knowledge} \citep{bok2018} and \textit{Subject Benchmark
+Statement} \citep{QAA2014} for the `why' and `what' of instruction, leaving \gp
+to deal with the `how'. Moreover, the open source, contributory nature of the
+project positions us to build \gp on top of the shared understanding of many
+specialists with a range of ideas about how to break apart---and put back
+together---the constituent elements of our domain knowledge in ways that speak
+to different types of students.
 
-For the time being we have also deliberately hobbled \gp in one important way: an \texttt{include} command must be in a cell that does not contain any other text or code.
-This was done primarily for simplicity: it's a lot easier to look for whole (text) cells that match a target pattern than to have to try to parse long blocks of text or code on the off-chance that something might match; it also avoids any ambiguity as to whether the \texttt{include} is a \gp command or `real' code as might happen, for instance, if \gp were used with a different programming language.
-Not coincidentally, it is also a good deal easier to know that you are replacing an entire block (the one with an include) with one or more entire blocks that are a mix of text and code, than trying to work out whether the block in which an include was found needs to be `closed out' before the import starts.
+For the time being we have also deliberately hobbled \gp in one important way:
+an \texttt{include} command must be in a cell that does not contain any other
+text or code. This was done primarily for simplicity: it's a lot easier to look
+for whole (text) cells that match a target pattern than to have to try to parse
+long blocks of text or code on the off-chance that something might match; it
+also avoids any ambiguity as to whether the \texttt{include} is a \gp command or
+`real' code as might happen, for instance, if \gp were used with a different
+programming language. Not coincidentally, it is also a good deal easier to know
+that you are replacing an entire block (the one with an include) with one or
+more entire blocks that are a mix of text and code, than trying to work out
+whether the block in which an include was found needs to be `closed out' before
+the import starts.
 
@@
 
-From practical experience, conference presentations, and code we tend to already know who is a good \emph{programmer} or \emph{theorist}, \gp provides a mechanism for discovering who is a good \emph{teacher}.
-Sometimes these abilities may reside in the same person, but more often we expect that they will not: the strongest developers tend to be people who have been practicing software development for many years and, consequently, may have difficulty communicating their ideas to beginner- or intermediate-level students (or teachers!).
-For this reason we see the broad-based community-of-practice aspect of \gp as integral at all stages of the project: system enhancement, content development, expanding coverage, and instructional design.
+From practical experience, conference presentations, and code we tend to already
+know who is a good \emph{programmer} or \emph{theorist}, \gp provides a
+mechanism for discovering who is a good \emph{teacher}. Sometimes these
+abilities may reside in the same person, but more often we expect that they will
+not: the strongest developers tend to be people who have been practicing
+software development for many years and, consequently, may have difficulty
+communicating their ideas to beginner- or intermediate-level students (or
+teachers!). For this reason we see the broad-based community-of-practice aspect
+of \gp as integral at all stages of the project: system enhancement, content
+development, expanding coverage, and instructional design.
 
-Consequently, \gp has a lot in common---both philosophically and practically---with the Software Carpentry movement \citep{SCF2016}, and although we seek to tackle a slightly narrower set of issues with a more re-usable set of resources, we can take both inspiration and warning from their experience.
-The benefit, we think, is that while it is possible to design \gp sessions and modules that follow the popular `bootcamp' approach to instruction (though see critique in \citealp{Feldon2017}), we want to enable the \textit{same} content to be employed in a flexible carpentry format as well as a `normal' classroom or \textsc{mooc} as required\ldots Or even to enable the instructor to mash all of those formats together such that they use a `bootcamp' format for the introduction to Unix and the command line, an online-formatted \gp resource for foundational concepts in computer science, and a traditional course format for the (geo)data analysis instruction.
-All pulled from the same source!
+Consequently, \gp has a lot in common---both philosophically and
+practically---with the Software Carpentry movement \citep{SCF2016}, and although
+we seek to tackle a slightly narrower set of issues with a more re-usable set of
+resources, we can take both inspiration and warning from their experience. The
+benefit, we think, is that while it is possible to design \gp sessions and
+modules that follow the popular `bootcamp' approach to instruction (though see
+critique in \citealp{Feldon2017}), we want to enable the \textit{same} content
+to be employed in a flexible carpentry format as well as a `normal' classroom or
+\textsc{mooc} as required\ldots Or even to enable the instructor to mash all of
+those formats together such that they use a `bootcamp' format for the
+introduction to Unix and the command line, an online-formatted \gp resource for
+foundational concepts in computer science, and a traditional course format for
+the (geo)data analysis instruction. All pulled from the same source!
 
-Ultimately, although \gp was developed with teaching needs in mind there is, of course, no reason why it couldn't be put to other uses: in combination with with \texttt{nteract} (\url{https://github.com/nteract/nteract}) it would allow developers or researchers to build fully-fledged applications as scripts assembled from a collection of notebooks; or as an addition to Netflix's notebook ecology to allow for enhanced resource-sharing and standardisation during the development phase before features and interfaces are `fixed' as libraries \citep{Ufford2018}.
-Nonetheless, our focus for the time being remains the cohort of teachers at all levels tasked with introducing programming material to their students and wondering where to begin.
-We hope that \gp makes a valuable contribution to this application domain and look forward to working with others to roll out a rich, reusable teaching framework.
+Ultimately, although \gp was developed with teaching needs in mind there is, of
+course, no reason why it couldn't be put to other uses: in combination with with
+\texttt{nteract} (\url{https://github.com/nteract/nteract}) it would allow
+developers or researchers to build fully-fledged applications as scripts
+assembled from a collection of notebooks; or as an addition to Netflix's
+notebook ecology to allow for enhanced resource-sharing and standardisation
+during the development phase before features and interfaces are `fixed' as
+libraries \citep{Ufford2018}. Nonetheless, our focus for the time being remains
+the cohort of teachers at all levels tasked with introducing programming
+material to their students and wondering where to begin. We hope that \gp makes
+a valuable contribution to this application domain and look forward to working
+with others to roll out a rich, reusable teaching framework.
 

### 2019-10-15 Geopyter.tex
Refs

@@
 that teaching students and researchers how to code and how to \textit{think}
-computationally presents new challenges \citep{Etherington2016,Muller2014}.
-Specifically, because of the way that these changes have unfurled, the majority
-of resources used in the classroom are being developed from scratch at each
-institution, and often by newly-appointed lecturers and assistant professors
-coming from the relatively small number of elite research facilities active in
-this domain (see discussion in \citealp{esrc2013}). Not only has this led to the
-duplication of effort at multiple sites, but it has also had an enormous impact
-on the productivity---as commonly measured by institutions---of early career
-researchers, many of whom are in their first teaching post.
+computationally presents new challenges
+\citep{Etherington2016,Muller2014,rey_09jgs}. Specifically, because of the way
+that these changes have unfurled, the majority of resources used in the
+classroom are being developed from scratch at each institution, and often by
+newly-appointed lecturers and assistant professors coming from the relatively
+small number of elite research facilities active in this domain (see discussion
+in \citealp{esrc2013}). Not only has this led to the duplication of effort at
+multiple sites, but it has also had an enormous impact on the productivity---as
+commonly measured by institutions---of early career researchers, many of whom
+are in their first teaching post.
 
@@
 integrated code, text, and multimedia resources. And we also looked to the
-Python Spatial Analysis Library (\textsc{p}y\textsc{sal}) project in which
-contributors take ownership of application areas through managed collaboration
-instead of a contributory `free-for-all'. The subsequent emergence of the
-Jupyter project \cite{kluyver16} from iPython has led to the rapid growth of a
-rich ecosystem of tools that support the kinds of interaction needed to make
-this an effective teaching resource: multiuser-notebooks (jupyterhub),
-automation of grading (nbgrader), interactive widgets for visualisation
-(ipywidgets), and packages to facilitate manipulation of the notebook itself
-(nbformat and notebook).
+Python Spatial Analysis Library (\textsc{p}y\textsc{sal}) project
+\cite{rey_19sea}in which contributors take ownership of application areas
+through managed collaboration instead of a contributory `free-for-all'. The
+subsequent emergence of the Jupyter project \cite{kluyver16} from iPython has
+led to the rapid growth of a rich ecosystem of tools that support the kinds of
+interaction needed to make this an effective teaching resource:
+multiuser-notebooks (jupyterhub), automation of grading (nbgrader), interactive
+widgets for visualisation (ipywidgets), and packages to facilitate manipulation
+of the notebook itself (nbformat and notebook).
 

### 2019-10-19 Geopyter.tex
Minor edits to include some new refs. Otherwise good to go as far as I'm concerned.

@@
 computationally presents new challenges
-\citep{Etherington2016,Muller2014,rey_09jgs}. Specifically, because of the way
-that these changes have unfurled, the majority of resources used in the
+\citep{Etherington2016,Muller2014,rey_09jgs} and that there is little in the way of a tradition of doing so to a high level within Geography (see surveys in \citealp{Bowlick2017,Bowlick2018}).
+Consequently, the majority of resources used in the
 classroom are being developed from scratch at each institution, and often by
@@
 developers.
+As such, this positions our work within contextualised programming initiatives \citep{Guzdial2010,Lukkarinen2016} of which the CS+X approach is perhaps best-known \citep[][\eg]{Mir2017}; however, \gp represents a significant advance in this area since it not only builds in context at the level of individual modules, but at the level of a shared set of teaching resources as well!
 
@@
 institution might \emph{wish} them to be.
+At the same time, there is also a strong need to support more open-ended exploration during `practicals' \citep{Unwin1980}, particularly by more advanced students who may `tune out' if the means of progression is overly structured and rigid.
 

### 2019-10-19 Geopyter.tex
submission to jgs

@@
-\documentclass[letter, 11pt]{article}
+\documentclass[letter, 11pt,titlepage]{article}
 
@@
 \title{Geographical Python Teaching Resources: GeoPyTeR}
-\author{Jonathan Reades \and Sergio Rey}
+\author{Jonathan Reades \and Sergio J. Rey}
+%\author{}
 
@@
 \begin{abstract}
-  Abstract goes here.
+Geopyter, an acronym of Geographical Python Teaching Resources, provides
+a hub for the distribution of `best practice' in computational and
+spatial analytic instruction; enabling instructors to quickly and
+flexibly remix contributed content to suit their needs and delivery
+framework; and encouraging contributors from around the world to `give
+back' whether in terms of how to teach individual concepts or deliver
+whole courses. As such, Geopyter is positioned at the confluence of two
+powerful streams of thought in software and education: the Free and Open
+Source Software (FOSS) movement in which contributors help to build
+better software, usually on an unpaid basis, in return for having access
+to better tools and the recognition of their peers); and the rise of
+Massive Open Online Courses (MOOCs), which seek to radically expand
+access to education by moving course content online and providing access
+to students anywhere in the world at little or no cost. This paper sets
+out in greater detail the origins and inspiration for Geopyter, the
+design of the system and, through examples, the types of innovative
+workflows that it enables for teachers. We believe that tools like
+Geopyter, which build on open teaching practices and promote the
+development of a shared understanding of what it is to \emph{be} a
+computational geographer represent an opportunity to expand the impact
+of this second wave of innovation in instruction while reducing the
+demands placed on those actively teaching in this area.
 \end{abstract}

### 2020-02-24 Geopyter.tex
Major rewrite incorporating Reviewer feedback. Still more work to do.

@@
 \newcommand{\ie}{i.e.~\/}
+\newcommand{\pysal}{\textsc{p}y\textsc{sal}~\/}
 \newcommand{\comment}[1]{\todo[inline, color=green!40]{#1}}
@@
 \usepackage{graphicx} 
+\usepackage{tabularx}
+\newcolumntype{s}{>{\hsize=.25\hsize \raggedright\arraybackslash}X}
+\newcolumntype{b}{>{\hsize=.75\hsize}X} 	
+\usepackage{booktabs}
+\usepackage{wrapfig}
 \usepackage[font=small,labelfont=bf]{caption}
@@
 \citeauthor{Tukey1962}'s \textit{The future of data analysis}
-(\citeyear{Tukey1962}), its emergence has led to a flourishing of tools and
-programmes as universities seek to engage with this nascent---and
-lucrative---interdisciplinary field. One of the most profound practical impacts
-of this development has been on the ways in which developers share and execute
-code: although systems for combining code, commentary, and results have been
-around for some time (\eg R-Markdown) these were intended primarily for
-replicating research outputs, not for the online teaching of a new technique or
-writing of an interactive introduction to a software library. The `open' ethos
-and rapid growth of data science, and the need for a lightweight interactive
-code-sharing platform mean that teachers in many disciplines are playing
-catch-up on both the methods taught and the formats used.
+(\citeyear{Tukey1962}), it is only recently that this set of previously disparate practices for working with large data sets has begun to be formalised in a way that might allow it to be taught in a university context instead of acquired on the job through `learning-by-doing'.
+One of the most profound practical impacts of this development has been on the ways in which developers share and execute code: although systems for combining code, commentary, and results have been around for some time (\eg R-Markdown) these were intended primarily for replicating research outputs and not for the teaching of a new technique or writing of an interactive introduction to a software library.
+The emergence of data science---and the overall pace of change in the tools and programming libraries that it employs---gave new impetus to the search for lightweight ways of sharing code and documentation, and an interactive browser-based technology called Jupyter was the result.
 
-As a discipline, geography has been acutely aware of these developments: changes
-in the volume and extent of geographical data available (\eg
-\citealp{Graham2013,gonzalez2013big,reades2016}), and in the number and range of
-methods employed to identify patterns in those data (\eg
-\citealp{Fan2016,Naik2017,Santibanez2015,Stevens2015,ArribasBel2017}), are
-allowing us to tackle questions thought unanswerable just a few years ago. At
-the same time, however, there is growing evidence of a `hollowing out' of
-traditional domain-specific skills: most of what a trained geographer used to do
-with specialist software can now be done using free online resources (\eg
-Google's Fusion Tables, MapBox, Carto, ArcOnline), but there remains unmet
-demand for graduates who can \textit{code} in order to take advantage of the
-latest (geo)algorithms and techniques \citep{Singleton2014,Singleton2016}.
+As a result, university educators in many disciplines are playing catch-up with both the methods and the tools.
+Geography has been acutely aware of these developments: changes in the volume and extent of spatial data available (\eg \citealp{Graham2013,gonzalez2013big,reades2016}), and in the number and range of methods employed to identify patterns in those data (\eg \citealp{Fan2016,Naik2017,Santibanez2015,Stevens2015,ArribasBel2017}), are allowing us to tackle questions thought unanswerable just a few years ago.
+At the same time, there is growing evidence of a `hollowing out' of once domain-specific skills \citep{Singleton2014,Singleton2016}: most of what a trained geographer used to do with specialist Geographic Information Systems (\textsc{gis}) software can now be done using free online resources (\eg Google's Fusion Tables, MapBox, Carto, ArcOnline).
+However, there remains unmet demand for graduates who can \emph{code} using the latest (geo)algorithms and techniques, so in this article we present a means of bringing the benefits of data science-approaches to instruction using Jupyter `notebooks' to the teaching of Python programming and geospatial analysis: Geographical Python Teaching Resources (GeoPyTer).
 
@@
 
-Regardless of whether this is a revolution \citep{Wyly2014,Torrens2010} or an
-evolution \citep{Barnes2013,Barnes2014,ArribasBel2018}, there is no question
-that teaching students and researchers how to code and how to \textit{think}
-computationally presents new challenges
-\citep{Etherington2016,Muller2014,rey_09jgs} and that there is little in the way of a tradition of doing so to a high level within Geography (see surveys in \citealp{Bowlick2017,Bowlick2018}).
-Consequently, the majority of resources used in the
-classroom are being developed from scratch at each institution, and often by
-newly-appointed lecturers and assistant professors coming from the relatively
-small number of elite research facilities active in this domain (see discussion
-in \citealp{esrc2013}). Not only has this led to the duplication of effort at
-multiple sites, but it has also had an enormous impact on the productivity---as
-commonly measured by institutions---of early career researchers, many of whom
-are in their first teaching post.
+Regardless of whether this is a revolution \citep{Wyly2014,Torrens2010} or an evolution \citep{Barnes2013,Barnes2014,ArribasBel2018}, there is no question that teaching students and researchers how to code and how to \textit{think} computationally presents new challenges \citep{Etherington2016,Muller2014,rey_09jgs} and that there is little in the way of a tradition of doing so to a high level within Geography (see surveys in \citealp{Bowlick2017,Bowlick2018}).
+Consequently, the majority of resources used in the classroom are being developed from scratch at each institution, often by newly-appointed lecturers and assistant professors coming from the relatively small number of elite research facilities active in this domain (see discussion in \citealp{esrc2013}).
+Not only has this led to the duplication of effort at multiple sites, but it has also had an enormous impact on the productivity of early career researchers, many of whom are in their first teaching post, as commonly measured by universities.
 
@@
 the same way, and the fact that few instructors have the luxury of both time and
-tenure to develop compelling new course material in splendid isolation. This
+tenure to develop compelling new course material in splendid isolation. This is 
 where the Geographical Python Teaching Resource
@@
 and conceptual content from existing Jupyter notebooks while providing their own
-`gloss' for this content. Our approach therefore seeks to develop a richer, open
-community of developers and teachers who work \textit{together} on new teaching
-materials, and to support instructors in quickly adapting existing materials to
-new delivery formats.
+`gloss' for this content.
+Our approach therefore seeks to develop a rich community of developers and teachers who work \textit{together} on new teaching materials, who make their output entirely \emph{open} to the wider community, and whoe support instructors in quickly adapting existing materials to new delivery formats.
 
@@
 
-Like many open source projects, \gp grew out of a need to scratch an itch: in
-our roles as teachers of (geo)computational concepts and methods we grappled on
-an annual basis with the demands of developing and updating instructional
-material with complex interdependencies. And, while the fundamental computing
-and analytic concepts may remain fairly stable, the field is highly-dynamic in
-terms of both the content and the rate of evolution of new features in
-contributed code upon which we rely for functionality. Currently, every time a
-widely-used library package is updated a protracted sequence of revisions
-ensues: some teachers, pressed for time, may resist making use of---or even
-alerting students to---the new features; others may leap in enthusiastically,
-only to discover that they hadn't fully thought through the challenges of
-conflicting dependencies and evolving feature sets.
+Like many open source projects, \gp grew out of a need to scratch an itch: in our roles as teachers of (geo)computational concepts and methods we grappled on an annual basis with the demands of developing and updating instructional material with complex interdependencies.
+And, while the fundamental computing and analytic concepts may remain fairly stable, the field is highly-dynamic in terms of both the content (\ie what should be in the curriculum) and the code (\ie how applications should be taught).
+Currently, every time a widely-used library package is updated a protracted sequence of revisions ensues: some teachers, pressed for time, may resist making use of---or even alerting students to---the new features; others may leap in enthusiastically, only to discover that they hadn't fully thought through the challenges of conflicting dependencies and evolving feature sets.
 
@@
 Python Spatial Analysis Library (\textsc{p}y\textsc{sal}) project
-\cite{rey_19sea}in which contributors take ownership of application areas
+\cite{rey_19sea} in which contributors take ownership of application areas
 through managed collaboration instead of a contributory `free-for-all'. The
-subsequent emergence of the Jupyter project \cite{kluyver16} from iPython has
+subsequent emergence of the Jupyter project \citep{kluyver16} from iPython has
 led to the rapid growth of a rich ecosystem of tools that support the kinds of
@@
 developers.
-As such, this positions our work within contextualised programming initiatives \citep{Guzdial2010,Lukkarinen2016} of which the CS+X approach is perhaps best-known \citep[][\eg]{Mir2017}; however, \gp represents a significant advance in this area since it not only builds in context at the level of individual modules, but at the level of a shared set of teaching resources as well!
+As such, this positions our work within contextualised programming initiatives \citep{Guzdial2010,Lukkarinen2016} of which the CS+X approach is perhaps best-known \citep[\eg][]{Mir2017}; however, \gp represents a significant advance in this area since it not only builds in context at the level of individual modules, but at the level of a shared set of teaching resources as well!
 
@@
 need to be able to begin their learning `journey' at some arbitrary point in the
-progression from `learning to code' towards `computational social scientist'
-\citep{Lazer2009}, instructors require the ability to assemble a course that
-meets students where they \emph{are} rather than where the instructor or
-institution might \emph{wish} them to be.
-At the same time, there is also a strong need to support more open-ended exploration during `practicals' \citep{Unwin1980}, particularly by more advanced students who may `tune out' if the means of progression is overly structured and rigid.
+progression from `learning to code' towards environmental, geospatial, or computational social scientist \citep{Lazer2009}.
+Instructors require the ability to assemble a course that meets students where they \emph{are} rather than where the instructor or institution might \emph{wish} them to be.
+At the same time, there is also a strong need to support more open-ended exploration during `practicals' \citep{Unwin1980}, particularly by more advanced students who may `tune out' if the means of progression is overly structured and rigid\ldots or just too slow to keep their attention.
 
@@
 
-Each atom would employ \emph{domain-specific} illustrative examples and code so
-as to anchor learning in relevant problems and applications. An atom on arrays
-written for geographers could start by showing how these simple data structures
-can be used to hold data about cities (\eg name, country, population); but a
-subsequent unit could build on this with an an illustration of how nested data
-structures allow us to add city location using nested coordinate pairs (\eg
-latitude and longitude within an array-of-arrays). This is why managed
-collaboration around sets of related concepts is so important: quite aside from
-the fact that no \textsc{foss} project succeeds without some level of executive
-control, it's critical that \textit{someone} is responsible for ensuring a
-degree of consistency in how the units fit together within a set of
-closely-related topics.
+\begin{wraptable}{r}{0.6\textwidth}
+\centering
+\caption{Overview of System Components}
+\label{table:1}
+\begin{tabularx}{0.6\textwidth} {sb}
+ \toprule
+ \textbf{Concept} & \textbf{Usage} \\
+ \midrule
+ Atom  & A `teachable idea' combining explanation and code for a core, typically singular, concept (\eg Variables, Lists, Projections, Robust Rescaling\ldots). \\
+ Sessions  & A `teachable unit' combining \textbf{multiple atoms} for delivery in a particular learning context (\eg Beginner Lab-based Course, Intermediate Flipped Module, Online Course\ldots) \\
+ Modules  & A \textbf{sequence of sessions} designed to achieve one or more pedagogical objective and/or required for credit in a particular education context. \\
+\bottomrule 
+\end{tabularx}
+\end{wraptable}
 
-The purpose of the bottom-up approach is that these units can then be flexibly
-assembled into Sessions: from the \textit{same ingredients} (\ie atoms) the
-instructor could create two quite different modules---one for delivery in
-person, one for delivery online---by organising the elements in different ways.
-Quite simply, we don't want to have to rewrite material for each format:
-lecture/practical, `flipped' classroom, or \textsc{mooc}. Furthermore, sessions
-designed for experts (\eg those pursuing continuing online education) might be
-able to `move' students through many more atoms of instruction in a single
-session than a similarly laid out (in terms of time) session designed for
-first-time programmers in an undergraduate programme. So sessions need to be
-able to \emph{incorporate} atoms in a way that minimises the level of effort
-involved in finding the `best' way to, for example, explain the concept of
-recursion while maximising the ability of the instructor to relate this concept
-to the students' practical experience (\eg by providing a `context' that is
-anchored in a locally relevant `story' or data).
+Each atom would employ \emph{domain-specific} illustrative examples and code so as to anchor learning in problems and applications relevant to the learner.
+An Atom on arrays written for geographers could start by showing how these simple data structures can be used to hold data about cities (\eg name, country, population); but a subsequent unit could build on this with an an illustration of how nested data structures allow us to add city location using nested coordinate pairs (\eg latitude and longitude within an array-of-arrays).
+This is why managed collaboration around sets of related concepts is so important: quite aside from the fact that no \textsc{foss} project succeeds without some level of executive control, it's critical that \textit{someone} is responsible for ensuring a degree of consistency in how the units fit together within a set of closely-related topics.
 
-Naturally, Sessions can then be grouped into Learning Modules (modules) that
-offer a coherent instructional programme over a period of weeks or months.
-Modules represent the highest level of abstraction in the proposed system, but
-they are also obviously the starting point from which instructors can organise
-their (remixed) atomic and sessional material into something incorporating a set
-of learning outcomes and a package of assessment appropriate to their students.
-However, our design reflects the expectation that contributions to \gp at each
-\emph{level} of instruction might be made by \emph{different} people: a domain
-expert in Spatial Bayes might be the right person to develop an Atom on the
-concept and its application, but not the right person to develop a module
-tackling advanced spatial analytic concepts where this is just one approach
-amongst many. Similarly, given the global diversity of delivery formats, a
-10-week term in Britain enables students to cover a very different `volume' of
+The purpose of the bottom-up approach is that these units can then be flexibly assembled into Sessions: from the \textit{same ingredients} (\ie Atoms) the instructor could create two quite different modules---one for delivery in person, one for delivery online---by organising the elements in different ways.
+Quite simply, we don't want to have to rewrite material for each format: lecture/practical, `flipped' classroom, or online course.
+Furthermore, sessions designed for experts (\eg those pursuing continuing online education) might be able to `move' students through many more atoms of instruction in a single session than a similarly laid out (in terms of time) session designed for first-time programmers in an undergraduate programme.
+So sessions need to be able to \emph{incorporate} atoms in a way that minimises the level of effort involved in finding the `best' way to, for example, explain the concept of recursion while maximising the ability of the instructor to relate this concept to the students' practical experience (\eg by providing a `context' that is anchored in a locally relevant `story' or data).
+
+Naturally, Sessions can then be grouped into Learning Modules (modules) that offer a coherent instructional programme over a period of weeks or months.
+Modules represent the highest level of abstraction in the proposed system, but they are also obviously the starting point from which instructors can organise their (remixed) atomic and sessional material into something incorporating a set of learning outcomes and a package of assessment appropriate to their students.
+However, our design reflects the expectation that contributions to \gp at each \emph{level} of instruction might be made by \emph{different} people: a domain expert in Spatial Bayes might be the right person to develop an Atom on the concept and its application, but not the right person to develop a module tackling advanced spatial analytic concepts where this is just one approach amongst many.
+Similarly, given the global diversity of delivery formats, a 10-week term in Britain enables students to cover a very different `volume' of
 content from a 15-week American semester. \gp recognises and seeks to respond to
@@
 in these languages as well.}, and it also takes advantage of the dominance of
-Git and GitHub for version control as a means for tracking authorship across
-many versions and changes.
+the Git tool---and the GitHub service/web site---for version control as a means for tracking authorship across many versions and changes.
 
@@
 `\#\#\#' indicates level 3 headers; etc. For our purposes, what's relevant is
-that these headers yield a natural hierarchy that corresponds closely to the
-h1\ldots{}h6 model used by the HTML markup language that lies at the heart of
-the World Wide Web. This hierarchy allows us to `abstract out' the problem of
-inferring the \emph{meaning} of different sections of the notebook since the
-instructor does it for us.
+that these headers naturally yield a semantic hierarchy that corresponds closely to the h1\ldots{}h6 model used by the HTML markup language that lies at the heart of the World Wide Web.
+This hierarchy allows us to `abstract out' the problem of inferring the \emph{meaning} of different sections of the notebook since the instructor does it for us through their use of headers.
+
+\subsection{Approach}
+
+In order to compile Atoms into Sessions and Modules, \gp necessarily requires a compositional syntax.
+We've noted the conceptual mapping between markdown and \textsc{html} formatting, above but how do we select some mix of code and markdown material in one notebook to be incorporated into another?
+And how do we do this in a way that is both simple to express and able to resolve ambiguity?
+Fortunately, such a model already exists and was hinted at in Figure \ref{fig:Illustration}: Cascading Style Sheets (\textsc{css}) uses well-understood `selectors' to specify one or more elements on a web page to which a set of presentational styles should be applied.
+
+With \textsc{css} an `h1' in a style sheet indicates that all \textsc{html} Level 1 Headers should observe the styling rules declared immediately afterwards; while `h1.important' specifies that only a Level 1 Header of the class `important' should be selected.
+In fact, \textsc{css} also allows for nested selectors: for some element(s) of a `larger' selection to be selected in turn.
+This is normally used to do things like specify mouseover behaviours for a menu: that all anchors (\ie links) that are within a division of class \texttt{menu} should act in \textit{this} way when the mouse passes over them.
+What is particularly elegant about this approach is that it provides a means for selecting and styling multiple pieces of content in the document in one declaration (where this is desirable) \textit{and} a means for disambiguating content with the same name but in different locations within a document's hierarchical structure (where it is not).
+
+Conceptually, \gp reframes this in terms of the meaning imparted by the instructor to the notebook: all cells coming after a Level 1 Header are considered to be part of that element's semantic field until another Level 1 Header is encountered or the end of the document is encountered, whichever comes first.
+So a Level 2 Header coming `after' (or `below', if you prefer) a Level 1 Header is still part of that higher-level element's semantic field, \textit{but} we can select it uniquely within the notebook using the standard \textsc{css} form of \texttt{h1.content h2.subcontent}.
+And with this we have essentially repurposed \textsc{css} as a means of selecting and importing content from one notebook into another!
+This is illustrated in schematic form in Figure \ref{fig:structure} where the \texttt{>} imply a hierarchical relationship.
+
+\begin{wrapfigure}{r}{0.51\textwidth}
+  \centering
+  \caption{Illustration of Atom Structure}
+  \includegraphics[width=0.5\textwidth]{Single_Document.pdf}
+  \label{fig:structure}
+\end{wrapfigure}
+
+Unfortunately, the nature of Jupyter notebooks does not allow this to happen dynamically at run-time, but it does allow something similar to happen when an instructor is `compiling' new Sessional and Module content.
+In short, the instructor writes whatever content they wish but, using syntax similar to the examples below, wherever they want to incorporate contributed content from \gp (or elsewhere) they have only to `include' it by specifying both a source and a selection.
+And this approach also works recursively: a notebook can include content from a notebook that itself includes content from another notebook\ldots
+
+\subsection{Syntax}\label{syntax}
+
+Adopting the approach outlined above: we typically envision an Atom as a short notebook focussing on a core concept or method (\eg lists, object-oriented design, or spatial autocorrelation); some or all of each Atom can then be selected and imported into a Session, which is itself a notebook; and the sessions can then be selected and managed through a Module, which can be a notebook or a set of notebooks.
+This process is initiated by the instructor creating a blank text cell in a Jupyter notebook and writing an `include' statement.
+The statement should be the \textit{only} content in the cell since \gp will be replacing the cell with an unknown number of whole text and code cells from the referenced notebook.
+
+Crucially, \texttt{include} statements can be freely intermingled with the instructor's own content, allowing the instructor to `frame' the concepts in a way that suits their teaching style but which saves them having to reinvent the wheel for each class.
+A Session tackling standardisation could include elements of the relevant Atom from \gp while still allowing the instructor to interject comments, observations, questions, and additional tasks to ground the learning experience in the local context (individual, institutional, etc.).
+To illustrate this more clearly, an Atom on Python's approach to dealing with lists could be incorporated into a longer Session as follows:
+
+\begin{Verbatim}[fontsize=\small]
+@include {
+    'nb'     = 'http://geopyter.org/atoms/fundamentals/lists.ipynb',
+    'select' = 'h1.Lists'
+}
+\end{Verbatim}
+
+\texttt{nb} is a path---local or remote---to a valid Jupyter notebook from which the Session or Module developer wants to import content.
+The \texttt{select} parameter specifies a selector for which the \gp tool will search within the source notebook.
+All content from that point onwards \emph{up to the next selector at he same level} will the then be copied into the compiled notebook.
+In example above, if there were a following \texttt{h1} covering, for example, `List Operators' then this would \emph{not} be included because, from a structural standpoint, it is of equal importance (at the same level in the hierarchy) to `Lists' but has \emph{not} been selected.
+Furthermore, any \texttt{h2} or \texttt{h3} subsections within the `Lists' section \textit{would} be included since they are presumed to be providing pedagogical and logical structure to the Understanding Lists section and so should be carried over.
+
+Clearly, an instructor might want to import only part of of a section, or to
+suppress a subsection falling in the middle of a larger resource. In
+anticipation of this need, more complex `include' statements with no equivalent
+in \textsc{css} are also possible:
+
+\begin{Verbatim}[fontsize=\small]
+@include {
+    'resource' = 'http://geopyter.org/atoms/fundamentals/lists.ipynb',
+    'select' = 'h1.Lists -h3.Lists Example; h1.List Operators -h2.Concatenation'
+}
+\end{Verbatim}
+
+In this second example, two level 1 section are imported at the same time (the
+selections are separated with semi-colons) and a level 3 subsection from
+\emph{within} each of those sections is suppressed using the `-' syntax to
+indicate that the section should be removed. An additional difference from true
+\textsc{css} is that we allow spaces in the `selector' because we felt that
+asking teachers to translate between a natural language header (``Understanding
+Lists'') and what \textsc{css} would consider a safe header
+(``Understanding\_Lists'') would detract from ease-of-use. A second departure
+from the \textsc{css} standard is that selectors are not separated with commas:
+we wanted to allow for this punctuation to be part of a section heading and felt
+that semi-colons are rather more rare in that context.
+
+\subsection{Putting it All Together}
 
@@
   \centering
-  \includegraphics[width=\textwidth, angle=0]{Illustration.pdf}
+  \includegraphics[width=\textwidth, angle=0]{Multiple_Documents.pdf}
   \caption{Illustration of \gp Notebook `Compilation' Process}
@@
 
-\subsection{Syntax}\label{syntax}
-
-The final element of \gp is necessarily a syntax for specifying how Sessions and
-Modules should be compiled from Atoms. We've noted the technical mechanism, as
-well as the mapping between markdown and \textsc{html} formatting, above but how
-do we select some mix of code and markdown material in one notebook to be
-incorporated into another? And how do we do this in a way that is both simple to
-express and able to resolve ambiguity? Fortunately, such a model already exists
-and was hinted at in Figure \ref{fig:Illustration}: Cascading Style Sheets
-(\textsc{css}) uses well-understood `selectors' to specify one or more elements
-on a web page to which a set of presentational styles should be applied: an `h1'
-in a style sheet indicates that all level 1 headers should observe the styling
-rules declared immediately afterwards; while `h1.important' specifies that only
-a level 1 header of the class `important' should be selected. More powerfully,
-`h1.important a.super' selects \emph{only} anchors of the class `super' that
-fall `within' (or `under', if you prefer') a level 1 header of the `important'
-class.
-
-What is particularly elegant about this approach is that it provides a means for
-selecting and styling multiple pieces of content in the document in one
-declaration (where this is desirable) \textit{and} a means for disambiguating
-content with the same name but in different positions within a document's
-hierarchical structure (where it is not). We have essentially repurposed
-\textsc{css} as a means of selecting and importing content from one notebook
-into another! Unfortunately, the nature of Jupyter notebooks does not allow this
-to happen dynamically at run-time, but it does allow something similar to happen
-when an instructor is `compiling' new Sessional and Module content. In short,
-the instructor writes whatever content they wish but, using syntax similar to
-the examples below, wherever they want to incorporate contributed content from
-\gp (or elsewhere) they have only to write an `include' statement to select and
-import sections of another notebook. This approach also works recursively: a
-notebook can include content from a notebook that itself includes content from
-another notebook\ldots
-
-Adopting the language used above: we envision an Atom as a short notebook
-focussing on a single concept or method (\eg lists, object-oriented design, or
-spatial autocorrelation); some or all of each Atom can then be selected and
-imported into a Session, which is itself a notebook; and the sessions can then
-be selected and managed through a Module, which can be a notebook or a set of
-notebooks. This process is initiated by the instructor creating a blank text
-cell in a Jupyter notebook and writing an `include' statement. The statement
-should be the \textit{only} content in the cell since \gp will be replacing the
-cell with an unknown number of whole text and code cells from the referenced
-notebook.
-
-Crucially, `include' statements can be freely intermingled with the instructor's
-own content, allowing the instructor to `frame' the concepts in a way that suits
-their teaching style but which saves them having to reinvent the wheel for each
-class. A Session tackling standardisation could include elements of the relevant
-Atom from \gp while still allowing the instructor to interject comments,
-observations, questions, and additional tasks to ground the learning experience
-in the local context (individual, institutional, etc.). To illustrate this more
-clearly, an Atom on Python's approach to dealing with lists could be
-incorporated into a longer Session as follows:
-
-\begin{Verbatim}[fontsize=\small]
-@include {
-    'nb'     = 'http://geopyter.org/atoms/fundamentals/lists.ipynb',
-    'select' = 'h1.Understanding Lists'
-}
-\end{Verbatim}
-
-The `nb' is the path---local or remote---to a valid Jupyter notebook from which
-the Session or Module developer wants to import content. The `select' parameter
-specifies a selector for which the \gp tool will search within the source
-notebook. All content from that point onwards \emph{up to the next selector at
-the same level} will the then be copied into the compiled notebook. In example
-above, if there were a following h1 covering, for example, `List Operators' then
-this would \emph{not} be included because, from a structural standpoint, it is
-of equal importance (at the same level in the hierarchy) to `Understanding
-Lists' but has \emph{not} been selected. Furthermore, any h2 or h3 subsections
-within the `Understanding Lists' section \textit{would} be included since they
-are presumed to be providing pedagogical and logical structure to the
-Understanding Lists section and so should be carried over.
-
-Clearly, an instructor might want to import only part of of a section, or to
-suppress a subsection falling in the middle of a larger resource. In
-anticipation of this need, more complex `include' statements with no equivalent
-in \textsc{css} are also possible:
-
-\begin{Verbatim}[fontsize=\small]
-@include {
-    'resource' = 'http://geopyter.org/atoms/fundamentals/lists.ipynb',
-    'select' = 'h1.Understanding Lists -h3.Lists Example; h1.Using Lists -h3.Another Example'
-}
-\end{Verbatim}
-
-In this second example, two level 1 section are imported at the same time (the
-selections are separated with semi-colons) and a level 3 subsection from
-\emph{within} each of those sections is suppressed using the `-' syntax to
-indicate that the section should be removed. An additional difference from true
-\textsc{css} is that we allow spaces in the `selector' because we felt that
-asking teachers to translate between a natural language header (``Understanding
-Lists'') and what \textsc{css} would consider a safe header
-(``Understanding\_Lists'') would detract from ease-of-use. A second departure
-from the \textsc{css} standard is that selectors are not separated with commas:
-we wanted to allow for this punctuation to be part of a section heading and felt
-that semi-colons are rather more rare in that context.
-
 \section{Use Cases}\label{uses}
@@
 Courses, or \textsc{mooc}s, as vehicles for expanding access to higher education
-through online delivery. For HE managers, the \textsc{mooc} promises both
+through online delivery. For managers in Higher Education (\textsc{he}), the \textsc{mooc} promises both
 enhanced revenue---depending on the model this implies either more people paying
@@
 
-Because it builds on \textsc{foss} approaches to software development, we expect
-\gp to benefit from network effects: the more people use it, the more useful it
-becomes, and the more people use it. In common with many such projects we
-\emph{do} expect that a relatively small subset of users will contribute the
-majority of the content; however, unlike a traditional software project there is
-an important role for \emph{teachers}, not just highly-skilled \emph{developers}
-and we think that this represents a really exciting opportunity for innovative
-new approaches to rise to the surface. There may be only a few who can write the
-Python code to conduct a Geographically-Weighted Regression analysis, but it
-will be interesting to see how many creative, insightful ways there are to
-explain it!
+Because it builds on \textsc{foss} approaches to software development, we expect \gp to benefit from network effects: the more people use it, the more useful it becomes, and the more people use it.
+The open source, peer-generated approach would, however, also be likely to present something of a challenge over time: as more people seek both to use and to contribute to \gp we would be likely to see the emergence of divergent norms, examples, and data across Atoms, Sessions, and Modules.
+In other words, the coherence of \gp as a teaching platform would become a victim of any real success in terms of widespread adoption.
 
-\gp also acknowledges the reality of the need for peer and professional
-recognition by incorporating attribution mechanisms through which faculty can be
-credited for their work. Contributors can be recognised as the author of content
-elements through a Digital Object Identifier (\textsc{doi}), and The \gp team
-would also include section editors responsible for the curation of atoms and for
-maintenance of official \gp \textsc{doi}s. Instructors who use \gp to compose
-sessions or courses can also contribute those modules back to the project and
-would also be recognised as the author of record in a similar fashion.
-Authorship of \gp educational components, be they atoms, sessions or modules,
-provides the contributors with peer-evaluated, impactful materials to add to
-their promotion applications.
+\subsection{Coordination}
+
+Here we believe that the \pysal project offers a useful template.
+Although \pysal is a free open source project, it is \emph{not} a free-for-all: domain specialists tend to gravitate towards those parts of the project to which they have the most to contribute and, over time, those who coordinate and enable the most substantive contributions to the codebase in terms of features and performance tend to become the `managers' of individual components of the overall tool (\eg lib, model, explore, viz).
+Overall project coherence is maintained via regular calls and online discussion boards, and a synchronised release schedule so that changes can be coordinated and knock-on effects identified.
+
+We expect \gp to follow a similar path in that a relatively small subset of users will contribute the majority of the content; however, unlike a traditional software project there is an important role for \emph{teachers}, not just highly-skilled \emph{developers}.
+We think that this represents a really exciting opportunity for innovative new approaches to rise to the surface.
+There may be only a few who can write the Python code to conduct a Geographically-Weighted Regression analysis, but it will be interesting to see how many creative, insightful ways there are to explain it!
+
+The educational focus implies that it may well be the most committed teachers who end up coordinating and reviewing contributions to \gp from other instructors and developers.
+We would also argue for the relevance of grouping sets of Atoms together in basic libraries such as `foundations' (the basics of variables and data structures); `description' (describing data); etc.
+Within these sets, contributions would be managed by a coordinator to ensure that examples, style, and other features are consistent for use in a set of Sessions introducing Python programming, or Local Indicators of Spatial Autocorrelation.
+
+The key to \gp is that it seeks to balance the benefits of code-sharing with those of local expertise: the instructor is free to write their own exegesis, if you will, of the code and its relevance to a particular Session or Module, but the burden of developing a cogent, domain-specific demonstration can be shared with others and compelling examples more widely adopted without the effort of reinventing the wheel.
+Where irreconcilable differences arise between pedagogical approaches or models, then we might expect to see small groups of collaborators `fork' the codebase and offer their own models; this is, of course, a valid approach with open source code and one from which all instructors can ultimately benefit!
+
+\subsection{Credit}
+
+\gp also acknowledges the reality of the need for peer and professional recognition by incorporating attribution mechanisms through which faculty can be credited for their work.
+Contributors can be recognised as the author of content elements through a Digital Object Identifier (\textsc{doi}), and The \gp team would also include section editors responsible for the curation of atoms and for maintenance of official \gp \textsc{doi}s.
+Instructors who use \gp to compose sessions or courses can also contribute those modules back to the project and would also be recognised as the author of record in a similar fashion.
+Authorship of \gp educational components, be they atoms, sessions or modules, provides the contributors with peer-evaluated, impactful materials to add to their promotion applications.
 

### 2020-02-25 Geopyter.tex
Finished edits. Need to work in OER literature in 1 main paragraph and probably a few refs here and there.

@@
 %\usepackage[letter,top=3cm,bottom=2cm,left=3cm,right=3cm,marginparwidth=1.75cm]{geometry}
+
+%% For writing 'code' in LaTeX
+\usepackage{textcomp}
+\usepackage{listings}
+\lstloadlanguages{Python,Perl,HTML}
+\usepackage[dvipsnames]{xcolor}
+\definecolor{LightGray}{gray}{0.95}
+\lstset{% general command to set parameter(s)
+	basicstyle=\small\ttfamily, % print whole listing small
+	keywordstyle=\color{BrickRed}\bfseries,
+	% underlined bold black keywords
+	identifierstyle=\color{Fuchsia}, % nothing happens
+	commentstyle=\color{Orange}\bfseries, % white comments
+	stringstyle=\color{OliveGreen}, % typewriter type for strings
+	showstringspaces=false} % no special string spaces
+\lstset{frameround=tttt,backgroundcolor=\color{LightGray},framesep=8pt,rulecolor=\color{Gray}}
 
@@
 Although \citet{Donoho2017} traces the origins of data science back to
-\citeauthor{Tukey1962}'s \textit{The future of data analysis}
+\citeauthor{Tukey1962}'s \emph{The future of data analysis}
 (\citeyear{Tukey1962}), it is only recently that this set of previously disparate practices for working with large data sets has begun to be formalised in a way that might allow it to be taught in a university context instead of acquired on the job through `learning-by-doing'.
-One of the most profound practical impacts of this development has been on the ways in which developers share and execute code: although systems for combining code, commentary, and results have been around for some time (\eg R-Markdown) these were intended primarily for replicating research outputs and not for the teaching of a new technique or writing of an interactive introduction to a software library.
-The emergence of data science---and the overall pace of change in the tools and programming libraries that it employs---gave new impetus to the search for lightweight ways of sharing code and documentation, and an interactive browser-based technology called Jupyter was the result.
+One of the most profound practical impacts of this development has been on the ways in which developers share and execute code: although systems for combining code, commentary, and results have been around for some time (\eg R-Markdown) these were intended primarily for replicating research outputs, with teaching and interaction as secondary considerations.
+The emergence of data science---and the overall pace of change in the tools and programming libraries that it employs---gave new impetus to the search for lightweight ways of sharing code and documentation, and an interactive browser-based platform called Jupyter \citep{kluyver16} was the result.
 
-As a result, university educators in many disciplines are playing catch-up with both the methods and the tools.
-Geography has been acutely aware of these developments: changes in the volume and extent of spatial data available (\eg \citealp{Graham2013,gonzalez2013big,reades2016}), and in the number and range of methods employed to identify patterns in those data (\eg \citealp{Fan2016,Naik2017,Santibanez2015,Stevens2015,ArribasBel2017}), are allowing us to tackle questions thought unanswerable just a few years ago.
-At the same time, there is growing evidence of a `hollowing out' of once domain-specific skills \citep{Singleton2014,Singleton2016}: most of what a trained geographer used to do with specialist Geographic Information Systems (\textsc{gis}) software can now be done using free online resources (\eg Google's Fusion Tables, MapBox, Carto, ArcOnline).
-However, there remains unmet demand for graduates who can \emph{code} using the latest (geo)algorithms and techniques, so in this article we present a means of bringing the benefits of data science-approaches to instruction using Jupyter `notebooks' to the teaching of Python programming and geospatial analysis: Geographical Python Teaching Resources (GeoPyTer).
+University educators in many disciplines are playing catch-up, but Geography---thanks to its long relationship with computation \citep{ArribasBel2018}---has been acutely aware of these developments: changes in the volume and extent of spatial data available (\eg \citealp{Graham2013,gonzalez2013big,reades2016}), and in the number and range of methods employed to identify patterns in those data (\eg \citealp{Fan2016,Naik2017,Santibanez2015,Stevens2015,ArribasBel2017}), are allowing us to tackle questions thought unanswerable just a few years ago.
+At the same time, there is evidence of a `hollowing out' of once domain-specific skills \citep{Singleton2014,Singleton2016}: most of what used to be done with specialist Geographic Information Systems (\textsc{gis}) software can now be done using free online resources (\eg Google's Fusion Tables, MapBox, Carto, ArcOnline).
+However, there remains unmet demand for graduates who can \emph{code} using the latest (geo)algorithms and techniques, so in this article we present one means of bringing the benefits of (spatial) data science-approaches to instruction using Jupyter `notebooks' for the teaching of the Python programming language and geospatial analysis.
 
@@
 
-Regardless of whether this is a revolution \citep{Wyly2014,Torrens2010} or an evolution \citep{Barnes2013,Barnes2014,ArribasBel2018}, there is no question that teaching students and researchers how to code and how to \textit{think} computationally presents new challenges \citep{Etherington2016,Muller2014,rey_09jgs} and that there is little in the way of a tradition of doing so to a high level within Geography (see surveys in \citealp{Bowlick2017,Bowlick2018}).
-Consequently, the majority of resources used in the classroom are being developed from scratch at each institution, often by newly-appointed lecturers and assistant professors coming from the relatively small number of elite research facilities active in this domain (see discussion in \citealp{esrc2013}).
-Not only has this led to the duplication of effort at multiple sites, but it has also had an enormous impact on the productivity of early career researchers, many of whom are in their first teaching post, as commonly measured by universities.
+Regardless of whether this is a revolution \citep{Wyly2014,Torrens2010} or an evolution \citep{Barnes2013,Barnes2014}, there is no question that teaching students and researchers how to code and how to \emph{think} computationally presents new challenges \citep{Etherington2016,Muller2014,rey_09jgs} and that there is little in the way of a tradition of doing so to a high level within Geography (see surveys in \citealp{Bowlick2017,Bowlick2018}).
+Consequently, the majority of spatial data science and analytics resources used in the classroom are being developed from scratch at each institution, often by newly-appointed lecturers and assistant professors coming from the relatively small number of elite research facilities active in this domain (see discussion in \citealp{esrc2013}).
+Not only has this led to the duplication of effort at multiple sites, but it also has an enormous impact on the productivity of early career researchers, many of whom are in their first teaching post.
 
-The overarching purpose of this software is therefore to address two seemingly
-contradictory issues: the recognition that no two instructors teach in precisely
-the same way, and the fact that few instructors have the luxury of both time and
-tenure to develop compelling new course material in splendid isolation. This is 
-where the Geographical Python Teaching Resource
-(\textsc{g}eo\textsc{p}y\textsc{t}e\textsc{r}) comes into play: although
-developed with geographers in mind, it provides a generic means by which
-instructors in any discipline can selectively incorporate and re-mix programming
-and conceptual content from existing Jupyter notebooks while providing their own
-`gloss' for this content.
-Our approach therefore seeks to develop a rich community of developers and teachers who work \textit{together} on new teaching materials, who make their output entirely \emph{open} to the wider community, and whoe support instructors in quickly adapting existing materials to new delivery formats.
+The overarching purpose of this software is therefore to address two seemingly contradictory issues: the recognition that no two instructors teach in precisely the same way, and the fact that few instructors have the luxury of both time and tenure to develop compelling new course material in splendid isolation.
+This is  where the Geographical Python Teaching Resource (\textsc{g}eo\textsc{p}y\textsc{t}e\textsc{r}) comes into play: although developed with geographers in mind, it provides a generic means by which instructors in any discipline can selectively incorporate and re-mix programming and conceptual content from existing Jupyter notebooks while providing their own `gloss' for this content.
+Our approach therefore seeks to develop a rich community of developers and teachers who work \emph{together} on new teaching materials, who make their output entirely \emph{open} to the wider community, and who \emph{support} instructors in quickly adapting existing materials to new delivery formats.
 
@@
 And, while the fundamental computing and analytic concepts may remain fairly stable, the field is highly-dynamic in terms of both the content (\ie what should be in the curriculum) and the code (\ie how applications should be taught).
-Currently, every time a widely-used library package is updated a protracted sequence of revisions ensues: some teachers, pressed for time, may resist making use of---or even alerting students to---the new features; others may leap in enthusiastically, only to discover that they hadn't fully thought through the challenges of conflicting dependencies and evolving feature sets.
+Although, in principle, there should be little interaction between curriculum and code, in practice changes to software libraries can make certain connections much harder, or much easier, to establish in the mind of the learner.
+For instance, until the widespread adoption of \texttt{geopandas} in Python (which itself depended on the widespread adoption of \texttt{pandas}) it was common to find that each spatial analysis library had its own data structures\footnote{A similar effect could be seen in \texttt{R} with the \texttt{tidyverse} and \texttt{sftools}.} and, consequently, students needed to be taught to work with spatial data in different ways depending on the library.
+Consolidation around a single approach freed up a lot of `space' in the curriculum for teaching content, not format.
 
-We initially explored with others---faculty with whom we regularly
-collaborate---the potential to `divide and conquer': we would each focus on
-different topics and then combine these contributions to create a full course.
-Though received with some enthusiasm, this plan did not survive contact with
-reality: first, when actually faced with the uncertain promise of receiving a
-full course in return for our efforts, many of us felt more secure developing
-our own material; second, we had underestimated the value that we each attach to
-the ability to personalise material to our own style of teaching. Since the
-shared materials were being developed individually according to the individual's
-own `style', any rationalisation or personalisation would have to be done by the
-instructor at a later date and without support from the rest of the project.
+In spite of this, every time a widely-used library is updated a protracted sequence of revisions ensues as teachers need to decide whether to resist making use of, or even alerting students to, the new features in order to save updating their materials, or whether to leap in enthusiastically only to struggle with conflicting dependencies and unstable feature sets.
+But on the basis that `many hands makes light work', we initially explored with other faculty the potential to `divide and conquer': we would each focus on different topics using the latest codebase and then combine these contributions to create a full course.
+Though received with some enthusiasm, this plan did not survive contact with reality: first, when actually faced with the uncertain promise of receiving a full course in return for our individual efforts, many of us felt more secure developing our own material; second, we had underestimated the value that we each attach to the ability to personalise material to our own style of teaching.
+Since the shared materials were being developed according to individual `style', any rationalisation or personalisation would have to be done by the instructor at a later date.
 
-Going back to the drawing board, we looked again at our own experiences of
-successful programming and instructional efforts. In particular, we took note of
-two powerful streams of thought in software and education: the importance
-attached by the Free and Open Source Software (\textsc{foss}) movement to both
-peer-recognition \citep{raymond_cathedral_1999} and distributed version control;
-and the rise of Virtual Learning Environments (\textsc{vle}s) and of Massive
-Open Online Courses (\textsc{mooc}s) with their use of rich, online learning
-experiences (\eg \citealp{Trafford2011,Cabiria2012}). Reflecting on our initial
-failure, we decided against trying to supply a single set of polished course
-materials and instead looked for a way to provide teachers with building blocks
-from which they could assemble an \textit{individualised} course suited to their
-institutional and pedagogical needs.
+\comment{Work in OER material below}
 
-We drew inspiration from the \textsc{foss} iPython project
-\citep{perez_2007_ipython} and its transformative impact on the sharing of
-integrated code, text, and multimedia resources. And we also looked to the
-Python Spatial Analysis Library (\textsc{p}y\textsc{sal}) project
-\cite{rey_19sea} in which contributors take ownership of application areas
-through managed collaboration instead of a contributory `free-for-all'. The
-subsequent emergence of the Jupyter project \citep{kluyver16} from iPython has
-led to the rapid growth of a rich ecosystem of tools that support the kinds of
-interaction needed to make this an effective teaching resource:
-multiuser-notebooks (jupyterhub), automation of grading (nbgrader), interactive
-widgets for visualisation (ipywidgets), and packages to facilitate manipulation
-of the notebook itself (nbformat and notebook).
+Going back to the drawing board, we looked again at our own experiences of successful programming and instructional efforts.
+In particular, we took note of two powerful streams of thought in software and education: the importance attached by the Free and Open Source Software (\textsc{foss}) movement to both peer-recognition \citep{raymond_cathedral_1999} and distributed version control; and the rise of Virtual Learning Environments (\textsc{vle}s) and of Massive Open Online Courses (\textsc{mooc}s) with their use of rich, online learning experiences (\eg \citealp{Trafford2011,Cabiria2012}).
+Reflecting on our initial failure, we decided against trying to supply a single set of polished course materials and instead looked for a way to provide teachers with building blocks from which they could assemble an \emph{individualised} course suited to their institutional and pedagogical needs.
 
-With these criteria and resources in mind, we began looking for ways to lower
-the `entry costs' for faculty to both take from, and share into, the corpus of
-instructional material, as well as for a way to address the
-customisation/localisation challenge. \gp was our solution: a single, open hub
-of teaching concepts and practical programming tasks that can be flexibly
-assembled into classes, or entire courses! In the \gp approach we recognise that
-teachers need to be able to develop their own classroom `story' at their own
-pace (and in conformance with their institution's teaching pattern), while
-supporting them with a library of up-to-date materials from which they can pick
-and choose. So in much the same way that a computer application is compiled from
-the contributions of many developers, \gp allows individual classes and entire
-courses to be `compiled' from the contributions of many skilled teachers and
-developers.
-As such, this positions our work within contextualised programming initiatives \citep{Guzdial2010,Lukkarinen2016} of which the CS+X approach is perhaps best-known \citep[\eg][]{Mir2017}; however, \gp represents a significant advance in this area since it not only builds in context at the level of individual modules, but at the level of a shared set of teaching resources as well!
+We drew inspiration from the original \textsc{foss} iPython project \citep{perez_2007_ipython}---expanded into the Jupyter project in 2016 \citep{kluyver16}---and its transformative impact on the sharing of integrated code, text, and multimedia resources.
+And we also looked to the Python Spatial Analysis Library (\textsc{p}y\textsc{sal}) project \citep{rey_19sea} in which contributors take ownership of particular application areas through managed collaboration instead of a contributory `free-for-all'.
+The subsequent growth of the Jupyter project has produced a rich ecosystem of tools to support the kinds of interaction needed to make this an effective teaching resource: multiuser-notebooks (\texttt{jupyterhub}), automation of grading (\texttt{nbgrader}), interactive widgets for visualisation (\texttt{ipywidgets}), and packages to facilitate manipulation of the notebook itself (\texttt{nbformat} and \texttt{notebook}).
+
+With these criteria and resources in mind, we began looking for ways to lower the `entry costs' for faculty to both take from, and share into, a corpus of instructional material, as well as for a way to address the customisation/localisation challenge.
+\gp is our solution: a single, open hub of geographically-focussed teaching concepts and practical programming tasks that can be flexibly assembled into classes, or entire courses!
+\gp recognises that teachers need to be able to develop their own classroom `story' at their own pace (and in conformance with their institution's teaching patterns), while supporting them with a library of up-to-date materials from which they can pick and choose.
+So in much the same way that a computer application is compiled from the contributions of many developers, \gp allows individual classes and entire courses to be `compiled' from the contributions of many skilled teachers and developers.
+As such, this positions our work within contextualised programming initiatives \citep{Guzdial2010,Lukkarinen2016}, of which the CS+X approach is perhaps best-known \citep[\eg][]{Mir2017}; however, \gp represents an important advance in this area since it not only builds in context at the level of individual modules, but at the level of teaching resources as well!
 
@@
 
-At least in our experience, the pace of technological and methodological change
-is such that the gap between what instructors know and what students know is
-substantial, and in many cases might even be growing. The differences within
-individual student cohorts may be greater still, however, and while students
-need to be able to begin their learning `journey' at some arbitrary point in the
-progression from `learning to code' towards environmental, geospatial, or computational social scientist \citep{Lazer2009}.
-Instructors require the ability to assemble a course that meets students where they \emph{are} rather than where the instructor or institution might \emph{wish} them to be.
-At the same time, there is also a strong need to support more open-ended exploration during `practicals' \citep{Unwin1980}, particularly by more advanced students who may `tune out' if the means of progression is overly structured and rigid\ldots or just too slow to keep their attention.
+At least in our experience, the pace of technological and methodological change in spatial data science is such that the gap between what instructors know and what students know is substantial, and in many cases might even be growing.
+But the differences within individual student cohorts may be greater still, and instructors need to be able to quickly assemble and update a course that meets students where they \emph{are} rather than where the instructor or institution might \emph{wish} them to be.
+At the same time, there is also a strong need to support more open-ended exploration during `practicals' \citep{Unwin1980}, particularly by more advanced students who may `tune out' if progression is overly structured and rigid\ldots or just too slow to keep their attention.
 
-The guiding insight behind \gp is that instruction in computation proceeds from
-fundamental units of learning typically built around \textit{computing} concepts
-(variables, lists/arrays, dictionaries/hashes, functions/subroutines, etc.) to
-fundamental units of learning built around \textit{analytic} concepts (cluster
-analysis, point patterns, spatial autocorrelation, etc.). These units must then
-be assembled in a way that speaks to the student cohort and its background; we
-mean this in two ways: first, that the examples used must speak to budding
-geographers, political scientists, historians, etc.; and second, that it should
-be possible to develop courses for different cohorts without having to start
-over from scratch. In other words, how can we enable teachers to reuse many of
-the building blocks employed in an advanced course for Masters students in an
-introductory undergraduate course?
+The guiding insight behind \gp is that instruction in programming outside of Computer Science proceeds from fundamental units of learning typically built around \emph{computing} concepts (variables, lists/arrays, dictionaries/hashes, functions/subroutines, etc.) to fundamental units of learning built around \emph{analytic} concepts (cluster analysis, point patterns, spatial autocorrelation, etc.).
+These units must then be assembled in a way that speaks to the student cohort and its background; we mean this in two ways: first, that the examples used must be domain-specific in order to speak to budding geographers, political scientists, historians, etc.; and second, that it should be possible to develop courses for different types of cohorts without having to start over from scratch.
+In other words, how can we enable teachers to reuse many of the building blocks employed in an advanced course for Masters students in an introductory course for Undergraduates?
 
@@
 
-From these constraints it was clear that our system needed to support a
-`bottom-up' approach to instructional design. So although the development of a
-course or class could---and probably should---start with a clear set of learning
-aims and outcomes, at a certain point the instructor will be searching for a
-examples with which to teach a particular concept: What \emph{is} a list or
-dictionary? What is \textit{k}-means clustering? We settled on the term `Atoms'
-to refer to these basic instructional units and, much like entries in the atomic
-table, we felt that they could be grouped together into sets of related
-concepts: the fundamentals of programming, point pattern analysis, machine
-learning, etc.
+From these constraints it was clear that our system needed to support a compositional, `bottom-up' approach to instructional design.
+So although the development of a course or class should obviously start out with a clear set of learning aims and outcomes, at a certain point the instructor will be searching for examples and code with which to teach a particular concept: What \emph{is} a list or dictionary?
+What \emph{is} \emph{k}-means clustering?
+We settled on the term `Atoms' to refer to these basic instructional units and, much like entries in the atomic table, we felt that they could be grouped together into sets of related concepts: the fundamentals of programming, point pattern analysis, machine learning, etc.
 
-\begin{wraptable}{r}{0.6\textwidth}
+\begin{table}[htbp]
 \centering
@@
 \label{table:1}
-\begin{tabularx}{0.6\textwidth} {sb}
+\begin{tabularx}{0.6\textwidth}{sb}
  \toprule
@@
 \end{tabularx}
-\end{wraptable}
+\end{table}
 
-Each atom would employ \emph{domain-specific} illustrative examples and code so as to anchor learning in problems and applications relevant to the learner.
-An Atom on arrays written for geographers could start by showing how these simple data structures can be used to hold data about cities (\eg name, country, population); but a subsequent unit could build on this with an an illustration of how nested data structures allow us to add city location using nested coordinate pairs (\eg latitude and longitude within an array-of-arrays).
-This is why managed collaboration around sets of related concepts is so important: quite aside from the fact that no \textsc{foss} project succeeds without some level of executive control, it's critical that \textit{someone} is responsible for ensuring a degree of consistency in how the units fit together within a set of closely-related topics.
+Each Atom would employ \emph{domain-specific} illustrative examples and code so as to anchor learning in problems and applications relevant to the learner\footnote{\gp is intended for Geography and Planning students, but could quite easily be `forked' to provide a similar set of domain-specific resources for economics, sociology, or literature students.}.
+An Atom could start by showing how a list can be used to hold data about cities (\eg name, country, population), and a subsequent set of `cells' (the basic `unit' of Jupyter a notebook) could build on this with an an illustration of how a list-of-lists allows us to add location as a latitude and longitude coordinate pair.
+We will return to some of the issues that this approach raises in \nameref{sec:engagement}, but it points to the importance of ensuring a degree of consistency in how the Atoms for a set of closely-related topics fit together.
 
-The purpose of the bottom-up approach is that these units can then be flexibly assembled into Sessions: from the \textit{same ingredients} (\ie Atoms) the instructor could create two quite different modules---one for delivery in person, one for delivery online---by organising the elements in different ways.
-Quite simply, we don't want to have to rewrite material for each format: lecture/practical, `flipped' classroom, or online course.
-Furthermore, sessions designed for experts (\eg those pursuing continuing online education) might be able to `move' students through many more atoms of instruction in a single session than a similarly laid out (in terms of time) session designed for first-time programmers in an undergraduate programme.
-So sessions need to be able to \emph{incorporate} atoms in a way that minimises the level of effort involved in finding the `best' way to, for example, explain the concept of recursion while maximising the ability of the instructor to relate this concept to the students' practical experience (\eg by providing a `context' that is anchored in a locally relevant `story' or data).
+The purpose of the bottom-up approach is that these units can then be flexibly assembled into Sessions: from the \emph{same ingredients} (\ie Atoms) the instructor could create quite different modules by organising and presenting the elements in different ways.
+Quite simply, we don't want to have to rewrite material for each format (\eg lecture/practical, `flipped' classroom, or distance learning), but we also need to deal with the fact that different types of scaffolding are required and that the amount of content suitable to a `class' in each of these formats might differ substantially.
 
-Naturally, Sessions can then be grouped into Learning Modules (modules) that offer a coherent instructional programme over a period of weeks or months.
+Furthermore, sessions designed for experts (\eg those pursuing continuing online education) might be able to `move' students through many more Atoms of instruction in a single Session than a similarly laid out on designed for first-time programmers in an undergraduate programme.
+So Sessions need to be able to \emph{incorporate} Atoms in a way that minimises the level of effort involved in finding the `best' way to, for example, explain the concept of recursion while maximising the ability of the instructor to relate this concept to the students' practical experience (\eg by providing a `context' that is anchored in a locally relevant `story' or data).
+
+Naturally, Sessions can then be grouped into Learning Modules that offer a coherent instructional programme over a period of weeks or months.
 Modules represent the highest level of abstraction in the proposed system, but they are also obviously the starting point from which instructors can organise their (remixed) atomic and sessional material into something incorporating a set of learning outcomes and a package of assessment appropriate to their students.
 However, our design reflects the expectation that contributions to \gp at each \emph{level} of instruction might be made by \emph{different} people: a domain expert in Spatial Bayes might be the right person to develop an Atom on the concept and its application, but not the right person to develop a module tackling advanced spatial analytic concepts where this is just one approach amongst many.
-Similarly, given the global diversity of delivery formats, a 10-week term in Britain enables students to cover a very different `volume' of
-content from a 15-week American semester. \gp recognises and seeks to respond to
-that diversity at all levels.
+Similarly, given the global diversity of delivery formats, a 10-week term in Britain enables students to cover a very different `volume' of content from a 15-week American semester.
+\gp recognises and seeks to respond to that diversity.
 
@@
 
-So we are trying to design a system in which sessions and modules are
-\emph{composed} out of atomic units of instruction that can be, optionally,
-surrounded by the instructor's own `narrative' or supported by more locally
-relevant (geo)data. We therefore want to produce a set of teaching materials
-that are highly portable, easily re-used or edited, and that enable the
-instructor to select only the elements from which they wish to compose their
-materials. As intimated above, \gp operationalises this through the Jupyter
-project and its ability to provide in-browser access to the Python
-interpreter\footnote{Other languages are also possible using different kernels;
-there is no reason that \gp couldn't be used to create instructional materials
-in these languages as well.}, and it also takes advantage of the dominance of
-the Git tool---and the GitHub service/web site---for version control as a means for tracking authorship across many versions and changes.
+So we are trying to design a system in which Sessions and Modules are \emph{composed} out of Atoms that can be, optionally, surrounded by the instructor's own `narrative'.
+We therefore want to produce a set of teaching materials that are highly portable, easily re-used or edited, and that enable the instructor to select only the elements from which they wish to compose their materials.
+As intimated above, \gp operationalises this through the Jupyter project and its ability to provide in-browser access to the Python interpreter\footnote{Other languages are also possible using different kernels; there is no reason that \gp couldn't be used to create instructional materials in these languages as well.}, and it also takes advantage of the dominance of the Git version control tool---and the GitHub service/web site---as a means of tracking authorship across edits.
 
@@
 
-Jupyter notebooks are written as a mix of executable code cells and
-non-executable text formatted with the widely-used `markdown' syntax. Notebook
-structure is provided through headers in markdown cells: a `\#' pre-pended to a
-line of text is generally taken to be the title of the notebook; `\#\#' at the
-start of a line provides a second level of structure (\ie level 2 headers);
-`\#\#\#' indicates level 3 headers; etc. For our purposes, what's relevant is
-that these headers naturally yield a semantic hierarchy that corresponds closely to the h1\ldots{}h6 model used by the HTML markup language that lies at the heart of the World Wide Web.
-This hierarchy allows us to `abstract out' the problem of inferring the \emph{meaning} of different sections of the notebook since the instructor does it for us through their use of headers.
+Jupyter notebooks are written as a mix of executable code cells and non-executable text formatted with the widely-used `markdown' syntax.
+Notebook structure is provided through headers in markdown cells: a `\#' pre-pended to a line of text is generally taken to be the title of the notebook; `\#\#' at the start of a line provides a second level of structure (\ie Level 2 headers); `\#\#\#' indicates Level 3 headers; etc.
+For our purposes, what's relevant is that these headers naturally yield a semantic hierarchy that corresponds closely to the h1\ldots{}h6 model used by the HTML markup language that lies at the heart of the World Wide Web.
+This hierarchy allows us to `abstract out' the problem of inferring the \emph{meaning} of cells in different sections of the notebook since the instructor does it for us through their use of headers.
 
@@
 
-In order to compile Atoms into Sessions and Modules, \gp necessarily requires a compositional syntax.
-We've noted the conceptual mapping between markdown and \textsc{html} formatting, above but how do we select some mix of code and markdown material in one notebook to be incorporated into another?
+In order to assemble Atoms into Sessions and Modules, \gp necessarily requires a compositional syntax.
+We've noted the conceptual mapping between markdown and \textsc{html} formatting above, but how do we select some mix of code and markdown material in one notebook to be incorporated into another?
 And how do we do this in a way that is both simple to express and able to resolve ambiguity?
-Fortunately, such a model already exists and was hinted at in Figure \ref{fig:Illustration}: Cascading Style Sheets (\textsc{css}) uses well-understood `selectors' to specify one or more elements on a web page to which a set of presentational styles should be applied.
+Fortunately, such a model already exists and was hinted at in Figure \ref{fig:structure}: Cascading Style Sheets (\textsc{css}) uses well-understood `selectors' to specify one or more elements on a web page to which a set of presentational styles should be applied.
 
-With \textsc{css} an `h1' in a style sheet indicates that all \textsc{html} Level 1 Headers should observe the styling rules declared immediately afterwards; while `h1.important' specifies that only a Level 1 Header of the class `important' should be selected.
-In fact, \textsc{css} also allows for nested selectors: for some element(s) of a `larger' selection to be selected in turn.
-This is normally used to do things like specify mouseover behaviours for a menu: that all anchors (\ie links) that are within a division of class \texttt{menu} should act in \textit{this} way when the mouse passes over them.
-What is particularly elegant about this approach is that it provides a means for selecting and styling multiple pieces of content in the document in one declaration (where this is desirable) \textit{and} a means for disambiguating content with the same name but in different locations within a document's hierarchical structure (where it is not).
+In \textsc{css} an `h1' in a style sheet indicates that all \textsc{html} Level 1 Headers (\eg \texttt{<h1>A Title</h1>}) should observe the styling rules declared immediately afterwards; while `h1.important' specifies that only a Level 1 Header of the class `important' should be selected (\eg \texttt{<h1 class="important">A Title</h1>}) and all other Level 1 headers ignored (\eg \texttt{<h1 class="unimportant">A Title</h1>}).
+In fact, \textsc{css} also allows for nested selectors in which `child' element(s) of a `parent' can be selected in turn.
+This is normally used to do things like specify mouseover behaviours for a menu: that all anchors (\ie links) that are within a division of class \texttt{menu} should act in \emph{this} way when the mouse passes over them (\eg \texttt{div.menu a.hover}).
+What is particularly elegant about \textsc{css} is that it provides a means for selecting multiple pieces of content in the document in one declaration (where this is desirable) \emph{and} a means for disambiguating content with the same name but in different locations within a document hierarchy (where it is not).
 
-Conceptually, \gp reframes this in terms of the meaning imparted by the instructor to the notebook: all cells coming after a Level 1 Header are considered to be part of that element's semantic field until another Level 1 Header is encountered or the end of the document is encountered, whichever comes first.
-So a Level 2 Header coming `after' (or `below', if you prefer) a Level 1 Header is still part of that higher-level element's semantic field, \textit{but} we can select it uniquely within the notebook using the standard \textsc{css} form of \texttt{h1.content h2.subcontent}.
-And with this we have essentially repurposed \textsc{css} as a means of selecting and importing content from one notebook into another!
-This is illustrated in schematic form in Figure \ref{fig:structure} where the \texttt{>} imply a hierarchical relationship.
+Conceptually, \gp adapts this syntax to allow us to select some or all of a Jupyter notebook using the structure imparted by the instructor: all cells coming after a Level 1 Header are considered to be part of that element's semantic field until another Level 1 Header is encountered or the end of the document is encountered, whichever comes first.
+And a Level 2 Header coming `after' (a `child', if you prefer) a Level 1 Header is considered part of that `parent' element's semantic field, \emph{but} we can select it uniquely within the notebook using the standard \textsc{css} form of \texttt{h1.content h2.subcontent}.
+This is illustrated in schematic form in Figure \ref{fig:structure}, but note that the \texttt{>} is simply make clear the hierarchical relationship.
+With this, we have essentially repurposed \textsc{css} as a means of selecting and importing content from one notebook into another!
 
-\begin{wrapfigure}{r}{0.51\textwidth}
+\begin{figure}[htbp]
   \centering
-  \caption{Illustration of Atom Structure}
+  \caption{Illustration of a single Atom's Structure}
   \includegraphics[width=0.5\textwidth]{Single_Document.pdf}
   \label{fig:structure}
-\end{wrapfigure}
+\end{figure}
 
@@
 In short, the instructor writes whatever content they wish but, using syntax similar to the examples below, wherever they want to incorporate contributed content from \gp (or elsewhere) they have only to `include' it by specifying both a source and a selection.
-And this approach also works recursively: a notebook can include content from a notebook that itself includes content from another notebook\ldots
+And this approach works recursively: a notebook can include content from a notebook that itself includes content from another notebook\ldots
 
@@
 
-Adopting the approach outlined above: we typically envision an Atom as a short notebook focussing on a core concept or method (\eg lists, object-oriented design, or spatial autocorrelation); some or all of each Atom can then be selected and imported into a Session, which is itself a notebook; and the sessions can then be selected and managed through a Module, which can be a notebook or a set of notebooks.
+To recap, we typically envision an Atom as a short notebook focussing on a core concept or method (\eg lists, object-oriented design, or spatial autocorrelation); some or all of each Atom can then be selected and imported into a Session, which is itself a notebook; and the sessions can then be selected and managed through a Module, which can \emph{also} be a notebook or a set of notebooks.
 This process is initiated by the instructor creating a blank text cell in a Jupyter notebook and writing an `include' statement.
-The statement should be the \textit{only} content in the cell since \gp will be replacing the cell with an unknown number of whole text and code cells from the referenced notebook.
+The statement should be the \emph{only} content in the cell since \gp will be replacing the cell with an unknown number of whole text and code cells from the referenced notebook.
 
-Crucially, \texttt{include} statements can be freely intermingled with the instructor's own content, allowing the instructor to `frame' the concepts in a way that suits their teaching style but which saves them having to reinvent the wheel for each class.
+Crucially, \texttt{include} statements can be freely intermingled with the instructor's own content (as shown in Figure \ref{fig:include}, allowing the instructor to `frame' the concepts in a way that suits their teaching style but which saves them having to reinvent the wheel for each class.
 A Session tackling standardisation could include elements of the relevant Atom from \gp while still allowing the instructor to interject comments, observations, questions, and additional tasks to ground the learning experience in the local context (individual, institutional, etc.).
@@
 
-\begin{Verbatim}[fontsize=\small]
+\comment{Convert into a Jupyter-like example.}
+
+\begin{figure}
+	\centering 
+	\caption{Illustrative Jupyter notebook content}
+	\label{fig:include}
+\begin{lstlisting}[language=HTML,frame=single]
+# Session 2: Lists
+
+Module leader: Associate Prof. X
+Contact information: prof.x@foo.bar.uk
+\end{lstlisting}	
+\begin{lstlisting}[language=Python,frame=single]
 @include {
@@
 }
-\end{Verbatim}
+\end{lstlisting}
 
-\texttt{nb} is a path---local or remote---to a valid Jupyter notebook from which the Session or Module developer wants to import content.
+\begin{lstlisting}[language=HTML,frame=single]
+## All Done?
+
+For next week please read the following: ...
+\end{lstlisting}
+
+\end{figure}
+
+Here, \texttt{nb} is a path---local or remote---to a valid Jupyter notebook from which the instructor wants to import content.
 The \texttt{select} parameter specifies a selector for which the \gp tool will search within the source notebook.
-All content from that point onwards \emph{up to the next selector at he same level} will the then be copied into the compiled notebook.
-In example above, if there were a following \texttt{h1} covering, for example, `List Operators' then this would \emph{not} be included because, from a structural standpoint, it is of equal importance (at the same level in the hierarchy) to `Lists' but has \emph{not} been selected.
-Furthermore, any \texttt{h2} or \texttt{h3} subsections within the `Lists' section \textit{would} be included since they are presumed to be providing pedagogical and logical structure to the Understanding Lists section and so should be carried over.
+All content from that point onwards \emph{up to the next selector at the same level} will the then be copied into the compiled notebook.
+In the example above, if there were a following \texttt{h1} covering, for example, `List Operators' then this would \emph{not} be included because, from a structural standpoint, it is at the same level in the hierarchy as `Lists' but has \emph{not} been selected.
+Furthermore, any \texttt{h2} or \texttt{h3} subsections within the `Lists' section \emph{would} be included since they are presumed to be providing pedagogical and logical structure to the Lists section and so should be carried over.
 
-Clearly, an instructor might want to import only part of of a section, or to
-suppress a subsection falling in the middle of a larger resource. In
-anticipation of this need, more complex `include' statements with no equivalent
-in \textsc{css} are also possible:
-
+Clearly, an instructor might want to import only part of of a section, or to suppress a subsection falling in the middle of a larger resource.
+In anticipation of this need more complex `include' statements with no equivalent in \textsc{css} are also possible:
 \begin{Verbatim}[fontsize=\small]
@@
 \end{Verbatim}
-
-In this second example, two level 1 section are imported at the same time (the
-selections are separated with semi-colons) and a level 3 subsection from
-\emph{within} each of those sections is suppressed using the `-' syntax to
-indicate that the section should be removed. An additional difference from true
-\textsc{css} is that we allow spaces in the `selector' because we felt that
-asking teachers to translate between a natural language header (``Understanding
-Lists'') and what \textsc{css} would consider a safe header
-(``Understanding\_Lists'') would detract from ease-of-use. A second departure
-from the \textsc{css} standard is that selectors are not separated with commas:
-we wanted to allow for this punctuation to be part of a section heading and felt
-that semi-colons are rather more rare in that context.
+In this second example, two Level 1 section are imported at the same time and a Level 3 subsection from \emph{within} each of those sections is suppressed using the `-' syntax to indicate that the section should be removed.
+We diverged from the \textsc{css} standard since that selectors are not separated with commas: we wanted to allow for this punctuation to be part of a section heading and felt that semi-colons are rather more rare in that context.
+An additional point of difference from true \textsc{css} is that we allow spaces in the `selector' because we felt that asking teachers to translate between a natural language header (``List Operators'') and what \textsc{css} would consider a safe header (``List\_Operators'') would detract from ease-of-use.
 
@@
 
-Notebooks are actually stored in a format called JavaScript Object Notation
-(\textsc{json}) that is not particularly easy for most humans to read, but as it
-is nonetheless highly-structured we can interact with it programmatically. The
-extensible nature of the \textsc{json} format also allows us to read and write
-both data and metadata not only to each notebook, but also to each and every
-sub-element of a notebook. Since metadata that is not understood by Jupyter is
-simply ignored, we can add our own fields to provide useful information related
-to instruction: the amount of time that the notebook's examples and content
-might take to cover in class; who should be given credit for contributing; any
-dependencies or requirements for installed libraries\ldots
+Jupyter notebooks use a format called JavaScript Object Notation (\textsc{json}) that is not particularly easy for most humans to read, but as it is nonetheless highly-structured we can interact with it programmatically.
+The extensible nature of the \textsc{json} format also allows us to read and write both data and metadata not only to each notebook, but also to each and every cell in a notebook.
+Since metadata that is not understood by Jupyter is simply ignored, we can add our own fields to provide useful information related to instruction such as who should be given credit for contributing and any dependencies or requirements for installed libraries\ldots
 
-In short, within the notebook we have two types of structure that can be used
-not only to provide information that may assist course instructors in developing
-content, but also to uniquely specify a section, subsection, or sub-subsection
-of content (code or text) at a fine-grained level. Taken together, this provides
-the foundation for a mechanism to select content in one notebook and import it
-into another. This therefore creates an opportunity to remix/mash up content
-while still being able to add an institutional or course-specific gloss wherever
-necessary. Each notebook might start with the instructor's contact information
-or by providing instructions for setting up the computing environment, but then
-make use of material developed by others for actual instruction. This process
-may seem quite abstract---and probably quite convoluted as well---but an
-illustration (Figure \ref{fig:Illustration} \vpageref[above]{fig:Illustration})
-may help to clarify why this process is so useful.
+Taken together, this provides the foundation for remixing/mashing up content while still enabling to add an institutional or course-specific gloss wherever necessary.
+Each notebook might start with the instructor's contact information or by providing instructions for setting up the computing environment, but then make use of material developed by others for actual instruction.
+This process may seem quite abstract---and probably quite convoluted as well---but an illustration (Figure \ref{fig:compiling} \vpageref[below]{fig:compiling}) may help to clarify why this process is so useful.
 
@@
   \centering
+  \caption{Illustration of \gp Notebook `Compilation' Process}
+  \label{fig:compiling}
   \includegraphics[width=\textwidth, angle=0]{Multiple_Documents.pdf}
-  \caption{Illustration of \gp Notebook `Compilation' Process}
-  \label{fig:Illustration}
 \end{figure}
-
-Finally, in order to manage all of this output we make use of Git and the GitHub
-web platform to monitor, approve, and roll back alterations to any submitted
-revisions to \gp Atoms, Sessions, or Modules. This provides additional
-capabilities such as allowing an instructor to `fix' their course to a
-particular version of an Atom (a `release' or `tag' in Git) such that
-development of the Atom by others can continue without the instructor having to
-worry that the explanation or code upon which they rely will suddenly change!
-Git also gives us the contributor information that we can propagate into
-compiled notebooks so that credit is retained, and users and institutions can
-better-understand the extent to which \gp is a \emph{joint} project relying on
-the contributions of many teachers and developers.
 
@@
 
-Two use-cases: an online module of eight sessions, and an in-person module
-lasting one semester might serve to further illustrate how this approach offers
-a substantively new way to think about teaching programming material more
-generally.
+To illustrate how this approach offers a substantively new way to think about teaching programming material more generally, we here present two schematic use-cases: an online module of eight sessions, and an in-person module lasting one semester.
+It's important to stress that we envision \emph{both} of these modules being built out of the same Atoms and that, because they are drawing from version-controlled source code, the content can be fixed on a particular release (\ie version).
 
-\subsection{An Online Module}\label{an-online-module}
+\subsection{Distance Learning Module}\label{an-online-module}
 
-There has been increasing interest at universities in Massive Open Online
-Courses, or \textsc{mooc}s, as vehicles for expanding access to higher education
-through online delivery. For managers in Higher Education (\textsc{he}), the \textsc{mooc} promises both
-enhanced revenue---depending on the model this implies either more people paying
-to take the modules or more paying for a certificate of completion at the
-end---and enhanced access as people who would be unable to attend the university
-in person are nonetheless able to pursue a degree. Our own experience suggests
-substantial student interest in `computational social science'
-\citep{Lazer2009}, with international students being the most keen on such
-modules as they are seen to provide marketable skills (programming) for
-graduates from a discipline (geography) with generally high employment rates
-(see \citeauthor{rgs2017} \citeyear{rgs2017}).
+There has been increasing interest at universities in Massive Open Online Courses, or \textsc{mooc}s, as vehicles for expanding access to higher education through online delivery.
+For managers in Higher Education (\textsc{he}), the \textsc{mooc} promises both enhanced revenue and enhanced access to underserved groups.
+Our own experience suggests substantial student interest in `computational social science' \citep{Lazer2009}, with international students being the most keen on such modules as they are seen to provide marketable skills (programming) for graduates from a discipline (geography) with generally high employment rates (see: \citeauthor{rgs2017} \citeyear{rgs2017}).
 
-Whatever the motivation, while expanding access to, and enrolment by, students
-is probably a `good thing', the process for developing a new online module or
-`converting' an existing face-to-face module to an online format remains far
-from easy. Typically, this step requires significant effort and investment by
-both the instructor and their institution: all course materials need to be
-online, recordings/videos of the lecturer presenting a key concept need to be
-made, the sessions need to be `chunked' so that a 3-hour practical becomes
-something more manageable, and assessments often need to be substantively
-redesigned.
+Associate Professor X wishes to offer a \emph{Foundations of Spatial Data Science} module to second year undergraduates, but anticipates high demand exceeding her capacity to teach in a a computer cluster and high attrition since many students will decide that programming is `not for them'.
+Consequently, the decision is made to offer this module on a distance-learning basis with pre-recorded video content and other rich media to support student learning.
+However, the `remote desktop' environment that will support this module is only updated every 18 months, often in the middle of term.
 
-\gp significantly reduces the overhead of several of these stages: rather than
-focussing on tasks such as how to explain or illustrate a particular concept,
-the instructor can focus the \emph{majority} of their effort on constructing the
-type of richer, fuller narrative that \textsc{mooc}s seem to require to engage
-students working remotely. If the instructor isn't there to ask questions and
-probe student learning on the spot, then the `frame' for each session (as well
-as for the module as a whole) becomes crucial as it helps to anchor student
-learning in concrete outcomes and encourages them to reflect upon, and
-integrate, their progress over the self-guided learning process.
+\gp significantly reduces the overhead of several of these stages: rather than focussing on tasks such as how to explain or illustrate a particular concept, the instructor can focus on developing the multimedia content.
+The similarity between learning to code and learning a language is noted, and each `lesson' is designed to be completed in under an hour so that students can proceed at their own pace without being overwhelmed.
+As well, using Git tags Prof.
+X can `fix' the version of the code and explanations used to one that is appropriate to the computing environment while continuing to update her teaching materials.
 
@@
 
-In a more traditional in-person format, \gp could be employed differently: since
-the instructor is able to interactively provide the `frame' or `scaffold' (see
-relevant discussion of `hypermedia' in \citealp{Azevedo2008}) upon which the
-learning is built, there is less need for a narrative around each task or weekly
-session. Here, the instructor might simply import a group of atoms to cover a
-set of concepts and step through them interactively with the students. The
-session could wrap up with a targeted project or mini-assessment that requires
-the students to translate the concepts into a new problem domain or investigate
-the process in more detail: \emph{e.g.} ``we've seen how we can use pandas and
-bokeh to analyse demographic groups in London, here's a link to the open data
-store page for Phoenix with similar data\ldots{}'' or ``Do you think that
-socioeconomic class is a useful metric? Investigate the source of the data we've
-just used and see if it squares with your expectations\ldots{}''
+Associate Professor X \emph{also} teaches a module for Masters students that follows the more traditional lecture+practical format.
+Although these students are also new to programming, they have already completed a required \textsc{gis} module.
+Here, \gp could be employed differently: since the instructor is able to interactively provide the `frame' or `scaffold' (see relevant discussion of `hypermedia' in \citealp{Azevedo2008}) upon which the learning is built, there is less need for a narrative around each task or weekly session.
 
-In effect, courses built on top of \gp become mash-ups, with the author able to
-mix and match atoms as needed for the module objectives, time available, and
-level of student. Both sessions and modules are built via composition: atoms are
-composed into sessions, sessions into modules. The role of the instructor is
-then to provide an integrative narrative that, in a sense, explains their choice
-of components. In both the online and offline contexts we think that this has
-the potential to free up instructors to concentrate on where they can most
-effectively `add value', not in figuring out yet \emph{another} way to explain
-how a list or dictionary works, but in explaining why they matter to a
-geography, political science, or literature major.
+The instructor might therefore simply import the same group of Atoms as above, but use a Session and Module template that leaves out the rich media and more detailed explanations since these will be discussed in-class with the students.
+The Sessions then wrap up with an additional mini-project or mini-assessment that requires the students to translate the concepts into a new problem domain or investigate the process in more detail: ``we've seen how we can use \texttt{pandas} and \texttt{bokeh} to explore and compare the distribution of demographic groups in London, here's a link to equivalent open data for Phoenix, Arizona\ldots{}''
 
-\section{Engagement}\label{engagement}
+In addition, since students are working on their own machines, the instructor can update to the latest-and-greatest much more rapidly.
+To enable Professor X to manage these competing requirements, we make use of Git and the GitHub web platform not only to monitor, approve, and roll back alterations to any submitted revisions, but also to provide release `tags' to which an instructor can bind a particular instance of a Session or Module.
+This `fixes' their course to a particular version of an Atom such that development of the Atom by others can continue without the instructor having to worry that the explanation or code upon which they rely will suddenly change!
 
-Because it builds on \textsc{foss} approaches to software development, we expect \gp to benefit from network effects: the more people use it, the more useful it becomes, and the more people use it.
-The open source, peer-generated approach would, however, also be likely to present something of a challenge over time: as more people seek both to use and to contribute to \gp we would be likely to see the emergence of divergent norms, examples, and data across Atoms, Sessions, and Modules.
-In other words, the coherence of \gp as a teaching platform would become a victim of any real success in terms of widespread adoption.
+\section{Engagement}\label{sec:engagement}
+
+In effect, in both of the example use cases the role of the instructor is to provide an integrative narrative that guides the students and contextualises their choice of components.
+We think that this has the potential to free up instructors to focus on where they can most effectively `add value'; not in developing yet \emph{another} way to show how a list or dictionary works, but in explaining why they matter to a geography, political science, or literature major (see \citealp{Bort2015} for an application in literature).
+
+And because it builds on \textsc{foss} approaches to software development, we expect \gp to benefit from network effects: the more people use it, the more useful it becomes, and the more people use it.
+The open source, peer-generated approach is also, however, likely to present something of a challenge over time: as more people seek both to use and to contribute to \gp we would expect to see the emergence and use of divergent norms, examples, and data across Atoms, Sessions, and Modules.
+Although the emergence of difference styles of coding is actually quite natural in programming and could be seen as a benefit to students in terms of teaching them about this aspect of programming, it also the case that \gp could become a victim its own success if it ceases to be coherent.
 
@@
 
+\comment{Is this characterisation accurate?}
+
 Here we believe that the \pysal project offers a useful template.
-Although \pysal is a free open source project, it is \emph{not} a free-for-all: domain specialists tend to gravitate towards those parts of the project to which they have the most to contribute and, over time, those who coordinate and enable the most substantive contributions to the codebase in terms of features and performance tend to become the `managers' of individual components of the overall tool (\eg lib, model, explore, viz).
-Overall project coherence is maintained via regular calls and online discussion boards, and a synchronised release schedule so that changes can be coordinated and knock-on effects identified.
+Although \pysal is a \textsc{foss} project, it is \emph{not} a free-for-all: domain specialists tend to gravitate towards those parts of the project to which they have the most to contribute and, over time, those who coordinate and enable the most substantive contributions to the codebase in terms of features and performance are `invited' to help manage individual components of the tool (\eg \texttt{lib}, \texttt{model}, \texttt{explore}, \texttt{viz}).
+Overall coherence is maintained via regular calls and online discussion boards, as well as a synchronised release schedule so that changes can be coordinated, tested, and knock-on effects resolved.
 
-We expect \gp to follow a similar path in that a relatively small subset of users will contribute the majority of the content; however, unlike a traditional software project there is an important role for \emph{teachers}, not just highly-skilled \emph{developers}.
+\begin{table}[hbtp]
+\centering
+\caption{Indicative Groups of Atoms}
+\label{table:2}
+\begin{tabularx}{0.6\textwidth}{sb}
+ \toprule
+ \textbf{Atomic Group} & \textbf{Indicative Content} \\
+ \midrule
+	\texttt{auto} & Approaches to spatial autocorrelation analysis. \\
+	\texttt{context} & Why learn to program? Example applications. Interviews. \\
+	\texttt{networks} & Working with network data and costs. \\ 
+	\texttt{describe} & Simple descriptive statistics.  \\
+	\texttt{statistics} & More advanced methods of comparing data. \\
+	\texttt{clustering} & Non-spatial and spatial clustering analysis. \\
+	\texttt{foundations} & Foundations of programming and Computer Science. \\
+	\texttt{os} & Basic aspects of interacting with Operating Systems programmatically. \\
+	\texttt{viz} & Visualising data and making maps. \\
+	\texttt{ml} & Approaches to Machine Learning in a spatial context. \\ 
+	\texttt{points} & Point pattern generation and analysis. \\ 
+	\texttt{zones} & Zonal statistics and relationship. \\
+	\texttt{models} & From regression to \textsc{gwr}. \\
+\bottomrule 
+\end{tabularx}
+\end{table}
+
+In general, we would expect to see a relatively small number of committed educators and developers creating and maintaining groups of Atoms that align with their areas of expertise, interests, and teaching responsibilities.
+However, unlike a traditional software project there is an important role here for \emph{teachers}, not just developers.
 We think that this represents a really exciting opportunity for innovative new approaches to rise to the surface.
@@
 
-The educational focus implies that it may well be the most committed teachers who end up coordinating and reviewing contributions to \gp from other instructors and developers.
-We would also argue for the relevance of grouping sets of Atoms together in basic libraries such as `foundations' (the basics of variables and data structures); `description' (describing data); etc.
-Within these sets, contributions would be managed by a coordinator to ensure that examples, style, and other features are consistent for use in a set of Sessions introducing Python programming, or Local Indicators of Spatial Autocorrelation.
+So although we also expect interested educators to begin almost immediately picking holes in the organisational structure proposed in Table \ref{table:2}, some kind of starting point is needed.
+Moreover, as we've mentioned elsewhere, in the event of serious disagreement other instructors are free `fork' the repository and begin changing material as they see fit.
+Indeed, there is nothing to prevent \gp Sessions and Modules drawing on content spread across multiple repositories following different organisational and developmental strategies: all that's need is a \textsc{url}!
 
-The key to \gp is that it seeks to balance the benefits of code-sharing with those of local expertise: the instructor is free to write their own exegesis, if you will, of the code and its relevance to a particular Session or Module, but the burden of developing a cogent, domain-specific demonstration can be shared with others and compelling examples more widely adopted without the effort of reinventing the wheel.
+The educational focus of \gp implies that it may well be the most committed teachers who end up as `section editors' who coordinate and review contributions.
+The editorial approach also aligns with the obvious benefits of grouping sets of Atoms together into basic sections such as \texttt{foundations} (the basics of variables and data structures); \texttt{describe} (describing data); etc.
+The section editor ensures that examples, style, and other features are consistent across individual Atoms to make it easy to generate a set of Sessions introducing the Unix file system or Local Indicators of Spatial Autocorrelation.
+
+\gp therefore seeks to balance the benefits of code-sharing with those of local expertise: the instructor is free to write their own exegesis, if you will, of the code and its relevance to a particular Session or Module, but the burden of developing a cogent, domain-specific demonstration can be shared with others and compelling examples more widely adopted without the effort of reinventing the wheel.
 Where irreconcilable differences arise between pedagogical approaches or models, then we might expect to see small groups of collaborators `fork' the codebase and offer their own models; this is, of course, a valid approach with open source code and one from which all instructors can ultimately benefit!
@@
 
-\gp also acknowledges the reality of the need for peer and professional recognition by incorporating attribution mechanisms through which faculty can be credited for their work.
-Contributors can be recognised as the author of content elements through a Digital Object Identifier (\textsc{doi}), and The \gp team would also include section editors responsible for the curation of atoms and for maintenance of official \gp \textsc{doi}s.
-Instructors who use \gp to compose sessions or courses can also contribute those modules back to the project and would also be recognised as the author of record in a similar fashion.
-Authorship of \gp educational components, be they atoms, sessions or modules, provides the contributors with peer-evaluated, impactful materials to add to their promotion applications.
+The use of Git/GitHub also gives us access to contributor information in the `commit' (\ie editing) logs that we can propagate into notebooks so that all contributors are recognised in the final output.
+Although the open nature of the project means that we cannot strictly enforce attribution, \gp seeks to make this the easier to do `by default' through the insertion of metadata into the compiled notebook which is then used to append a list of contributors to the end of each notebook, along with any other relevant acknowledgments or copyright notices.
 
-Although the open source nature of the project means that we cannot strictly
-enforce attribution, \gp seeks to make this the easier to do `by default'
-through the use of Git to track and insert contributory metadata into the
-compiled notebook. We can use this metadata to append a list of contributors to
-the end of each notebook, along with any other relevant acknowledgments or
-copyright notices. To facilitate re-use while protecting contributions from
-unacknowledged exploitation textual content in \gp is covered by a Creative
-Commons license; however, to deal with the fact that \gp relies extensively on
-open source contributions which are incompatible with some \textsc{cc} licenses
-\citep[see discussion in][]{osswatch2013}, code blocks are licensed under the
-\textsc{mit} license. The manner in which contributions from authors at
-different institutions can be combined also `pollutes' the materials in ways
-that inhibit institutional assertions of ownership over \gp content.
+To facilitate re-use while protecting contributions from unacknowledged exploitation textual content in \gp is covered by a Creative Commons license; however, to deal with the fact that \gp relies extensively on open source contributions which are incompatible with some \textsc{cc} licenses \citep[see discussion in][]{osswatch2013}, code blocks are licensed under the \textsc{mit} license.
+The manner in which contributions from authors at different institutions can be combined also `pollutes' the materials in ways that inhibit institutional assertions of ownership over \gp content.
+
+\gp also recognises the reality of the need for peer and professional recognition by incorporating attribution mechanisms directly into the compilation process.
+Digital Object Identifiers (\textsc{doi}s) can be created and curated by the `section editors', but authorship of \gp components---be they Atoms, Sessions or Modules---provides the contributors with peer-evaluated, impactful materials to add to promotion applications.
+Our intention is that both users and institutions come to better-understand the extent to which open educational resources can be a \emph{joint} project relying on the contributions of many teachers and developers.
 
@@
 
-Although we are profoundly excited both by the possibility of shifting to
-compositional approaches to content-development, and by the impact that shared
-materials could have on our teaching, it is highly unlikely that our first
-attempt to divide up the entire field into discrete units of instruction will be
-entirely successful. For this, it will be useful to reference documents such as
-the \textit{Body of Knowledge} \citep{bok2018} and \textit{Subject Benchmark
-Statement} \citep{QAA2014} for the `why' and `what' of instruction, leaving \gp
-to deal with the `how'. Moreover, the open source, contributory nature of the
-project positions us to build \gp on top of the shared understanding of many
-specialists with a range of ideas about how to break apart---and put back
-together---the constituent elements of our domain knowledge in ways that speak
-to different types of students.
+As we noted above in connection with Table \ref{table:2}, it is rather unlikely that our first attempt to divide up the entire field into discrete units of instruction will be entirely successful.
+We would also expect to draw on reference documents such as the \emph{Body of Knowledge} \citep{bok2018} and \emph{Subject Benchmark Statement} \citep{QAA2014} for the `why' and `what' of instruction, leaving \gp to deal with the `how'.
+Moreover, the open, contributory nature of the project positions us to build \gp on top of the shared understanding of many specialists with a range of ideas about how to break apart, and put back together, the constituent elements of our domain's  knowledge in ways that speak to different types of students.
 
-For the time being we have also deliberately hobbled \gp in one important way:
-an \texttt{include} command must be in a cell that does not contain any other
-text or code. This was done primarily for simplicity: it's a lot easier to look
-for whole (text) cells that match a target pattern than to have to try to parse
-long blocks of text or code on the off-chance that something might match; it
-also avoids any ambiguity as to whether the \texttt{include} is a \gp command or
-`real' code as might happen, for instance, if \gp were used with a different
-programming language. Not coincidentally, it is also a good deal easier to know
-that you are replacing an entire block (the one with an include) with one or
-more entire blocks that are a mix of text and code, than trying to work out
-whether the block in which an include was found needs to be `closed out' before
-the import starts.
+For the time being we have also deliberately hobbled \gp in one important way: an \texttt{include} command must be in a cell that does not contain any other text or code.
+This was done primarily for simplicity: it's a lot easier to look for whole (text) cells that match a target pattern than to have to try to parse long blocks of text or code on the off-chance that an \texttt{include} might be found; it also avoids any ambiguity as to whether the \texttt{include} is a \gp or `native' command.
+Not coincidentally, it is also a good deal easier to replace an entire cell (the one containing an \texttt{include}) with one or more entire cells, than to try to work out if a cell needs to be `closed out' first.
 
@@
 
-From practical experience, conference presentations, and code we tend to already
-know who is a good \emph{programmer} or \emph{theorist}, \gp provides a
-mechanism for discovering who is a good \emph{teacher}. Sometimes these
-abilities may reside in the same person, but more often we expect that they will
-not: the strongest developers tend to be people who have been practicing
-software development for many years and, consequently, may have difficulty
-communicating their ideas to beginner- or intermediate-level students (or
-teachers!). For this reason we see the broad-based community-of-practice aspect
-of \gp as integral at all stages of the project: system enhancement, content
-development, expanding coverage, and instructional design.
+From practical experience, conference presentations, and code we tend to already know who is a good \emph{programmer} or \emph{theorist}, \gp provides a mechanism for discovering who is a good \emph{teacher}.
+Sometimes these abilities may reside in the same person, but more often we expect that they will not: the strongest developers tend to be people who have been practicing software development for many years and, consequently, may have difficulty communicating their ideas to beginner- or intermediate-level students or teachers (see: \citealp{Chapman:2010aa})!
+For this reason we see the broad-based community-of-practice aspect of \gp as integral at all stages of the project: system enhancement, content development, expanding coverage, and instructional design.
 
-Consequently, \gp has a lot in common---both philosophically and
-practically---with the Software Carpentry movement \citep{SCF2016}, and although
-we seek to tackle a slightly narrower set of issues with a more re-usable set of
-resources, we can take both inspiration and warning from their experience. The
-benefit, we think, is that while it is possible to design \gp sessions and
-modules that follow the popular `bootcamp' approach to instruction (though see
-critique in \citealp{Feldon2017}), we want to enable the \textit{same} content
-to be employed in a flexible carpentry format as well as a `normal' classroom or
-\textsc{mooc} as required\ldots Or even to enable the instructor to mash all of
-those formats together such that they use a `bootcamp' format for the
-introduction to Unix and the command line, an online-formatted \gp resource for
-foundational concepts in computer science, and a traditional course format for
-the (geo)data analysis instruction. All pulled from the same source!
+Consequently, \gp has a lot in common---both philosophically and practically---with the Software Carpentry movement \citep{SCF2016}, and although we seek to tackle a slightly narrower set of issues with a more re-usable set of resources, we can take both inspiration and warning from their experience.
+The benefit, we think, is that while it is possible to design Sessions and Modules that follow the popular `bootcamp' approach to instruction (though see critique in \citealp{Feldon2017}), we want to enable the \emph{same} content to be employed in a carpentry format as well as a `normal' classroom or \textsc{mooc} as required\ldots Or even to enable the instructor to mash all of those formats together such that they use a `bootcamp' format for the introduction to Unix and the command line, an online-formatted resource for foundational concepts in computer science, and a traditional course format for the (geo)data analysis instruction.
+All pulled from the same set of source Atoms!
 
-Ultimately, although \gp was developed with teaching needs in mind there is, of
-course, no reason why it couldn't be put to other uses: in combination with with
-\texttt{nteract} (\url{https://github.com/nteract/nteract}) it would allow
-developers or researchers to build fully-fledged applications as scripts
-assembled from a collection of notebooks; or as an addition to Netflix's
-notebook ecology to allow for enhanced resource-sharing and standardisation
-during the development phase before features and interfaces are `fixed' as
-libraries \citep{Ufford2018}. Nonetheless, our focus for the time being remains
-the cohort of teachers at all levels tasked with introducing programming
-material to their students and wondering where to begin. We hope that \gp makes
-a valuable contribution to this application domain and look forward to working
-with others to roll out a rich, reusable teaching framework.
+Ultimately, although \gp was developed with teaching needs in mind there is, of course, no reason why it couldn't be put to other uses: in combination with with \texttt{nteract} (\url{https://github.com/nteract/nteract}) it would allow developers or researchers to build fully-fledged applications as scripts assembled from a collection of notebooks; or as an addition to Netflix's notebook ecology to allow for enhanced resource-sharing and standardisation during the development phase before features and interfaces are `fixed' as libraries \citep{Ufford2018}.
+Nonetheless, our focus for the time being remains the cohort of university teachers at all levels tasked with introducing programming material to their students and wondering where to begin.
+We hope that \gp makes a valuable contribution to this application domain and look forward to working with others to roll out a rich, reusable teaching framework.
 

### 2020-02-26 Geopyter.tex
Added useful OER links.

@@
 
-\comment{Work in OER material below}
-
 Going back to the drawing board, we looked again at our own experiences of successful programming and instructional efforts.
-In particular, we took note of two powerful streams of thought in software and education: the importance attached by the Free and Open Source Software (\textsc{foss}) movement to both peer-recognition \citep{raymond_cathedral_1999} and distributed version control; and the rise of Virtual Learning Environments (\textsc{vle}s) and of Massive Open Online Courses (\textsc{mooc}s) with their use of rich, online learning experiences (\eg \citealp{Trafford2011,Cabiria2012}).
+In particular, we took note of two powerful streams of thought in software and education with which we already familiar: the importance attached by the Free and Open Source Software (\textsc{foss}) movement to both peer-recognition \citep{raymond_cathedral_1999} and distributed control; and the rise of Virtual Learning Environments (\textsc{vle}s) and Massive Open Online Courses (\textsc{mooc}s) with their use of rich, online learning experiences (\eg \citealp{Trafford2011,Cabiria2012}).
 Reflecting on our initial failure, we decided against trying to supply a single set of polished course materials and instead looked for a way to provide teachers with building blocks from which they could assemble an \emph{individualised} course suited to their institutional and pedagogical needs.
@@
 As such, this positions our work within contextualised programming initiatives \citep{Guzdial2010,Lukkarinen2016}, of which the CS+X approach is perhaps best-known \citep[\eg][]{Mir2017}; however, \gp represents an important advance in this area since it not only builds in context at the level of individual modules, but at the level of teaching resources as well!
+
+Without realising it, we were plugging into a larger framework of thinking emerging from the literature on Open Educational Resources (\textsc{oer}), which began as largely `legal and economic concept' designed to support royalty and license free access to educational resources \citep{butcher2010open}.
+It should be clear from this definition that there is no \emph{necessary} link to \textsc{foss} since a free e-book can qualify as an \textsc{oer}.
+However, these two areas are conceptually aligned in many ways, not least through their shared interest in the `4 Rs': reuse, redistribution, revision, and remixing \citep{Hilton2010}.
+Within \citeauthor{du2017}'s \citeyearpar{du2017} \textsc{oer} typology \gp is clearly `open courseware', but it is conceptually distinct from \textsc{mit}'s eponymous OpenCourseWare or Standford's Coursera since it is designed to promote what \citeauthor{Ehlers2011} calls `Open Educational Practice' \citeyearpar{Ehlers2011} in a manner not altogether dissimilar to that the Geographic Science \& Technology Body of Knowledge \citep{bok2018}.
+
+In line with \citet{Mishra2017} and with \citeauthor{Knox2013}'s \citeyearpar{Knox2013} call for a role for pedagogy in \textsc{oer}, \gp envisions an essential role for `teachers as (co)creators': they not only have an active role in selecting components for a Session or Module, they can contribute source material as well.
+The educator and section editors, of which more later, may therefore also use this process to extend their own understanding in an applied workplace setting through collaboration (see \citealp{littlejohn2017} and \citealp{eraut2008} for relevant discussions).
+Of course, we must also recognise that under \citeauthor{Wiley2009}'s \citeyearpar{Wiley2009} \textsc{alms} typology, although \gp provides direct Access to editing tools, is Meaningfully editable, and offers Source-file access, the Level of expertise required is not insubstantial though we also expect instructors in this domain to be more comfortable than most with the process.
+We therefore believe that \gp is unique in incorporating the instructor's domain expertise and pedagogical strategies by design.
 
@@
 To illustrate this more clearly, an Atom on Python's approach to dealing with lists could be incorporated into a longer Session as follows:
-
-\comment{Convert into a Jupyter-like example.}
 

### 2020-03-04 response.tex
ENH: response done and small edits on main

% LaTeX rebuttal letter example. 
% 
% Copyright 2019 Friedemann Zenke, fzenke.net
%
% Based on examples by Dirk Eddelbuettel, Fran and others from 
% https://tex.stackexchange.com/questions/2317/latex-style-or-macro-for-detailed-response-to-referee-report
% 
% Licensed under cc by-sa 3.0 with attribution required.
% See https://creativecommons.org/licenses/by-sa/3.0/
% and https://stackoverflow.blog/2009/06/25/attribution-required/

\documentclass[11pt]{article}
\usepackage[utf8]{inputenc}
\usepackage{lipsum} % to generate some filler text
\usepackage{fullpage}
\usepackage[colorlinks=false, allcolors=black]{hyperref}

% import Eq and Section references from the main manuscript where needed
% \usepackage{xr}
% \externaldocument{manuscript}

% package needed for optional arguments
\usepackage{xifthen}
% define counters for reviewers and their points
\newcounter{reviewer}
\setcounter{reviewer}{0}
\newcounter{point}[reviewer]
\setcounter{point}{0}

% This refines the format of how the reviewer/point reference will appear.
\renewcommand{\thepoint}{P\,\thereviewer.\arabic{point}}

% command declarations for reviewer points and our responses
\newcommand{\reviewersection}{\stepcounter{reviewer} \bigskip \hrule
                  \section*{Reviewer \thereviewer}}

\newenvironment{point}
   {\refstepcounter{point} \bigskip \noindent {\textbf{Reviewer~Point~\thepoint} } ---\ }
   {\par }

\newcommand{\shortpoint}[1]{\refstepcounter{point}  \bigskip \noindent 
	{\textbf{Reviewer~Point~\thepoint} } ---~#1\par }

\newenvironment{reply}
   {\medskip \noindent \begin{sf}\textbf{Reply}:\  }
   {\medskip \end{sf}}

\newcommand{\shortreply}[2][]{\medskip \noindent \begin{sf}\textbf{Reply}:\  #2
	\ifthenelse{\equal{#1}{}}{}{ \hfill \footnotesize (#1)}%
	\medskip \end{sf}}

\begin{document}

\section*{Response to the editor}

Dear Serge and Jon,

Many thanks for your paper submitted in response to the invitation to contribute to the special issue on Open Software for Spatial Science. I have received three reviewer reports now. At this point, I am sending the paper back to you with an invitation to make revisions.

For your information, the reviewers' comments can be found at the end of this email or can be accessed by following the provided link.

Your revision is due by 09 Mar 2020. That said, two papers for the special issue came in later than the rest, so there is some flexibility with the due date.

\begin{reply}
  Thank you, and the reviewers, for the careful evaluation of our revised manuscript.

  We find the points raised are most helpful in strengthening the contribution of this work, and below we detail the changes we made in response to each point. Following these points, we have included a marked-up version that shows the
  detailed changes in the manuscript.

  We feel that the substantial revision has resulted in a significantly improved
  paper, and we are grateful for the opportunity to revise this work.

\end{reply}
\pagebreak
\section*{Response to the reviewers}
% General intro text goes here
We thank the reviewers for their critical assessment of our work. 
In the following we address their concerns point by point.

% Let's start point-by-point with Reviewer 1
\reviewersection
The paper presents some details on Geopyter, a hub for the distribution of contents in computational and spatial analytic instruction. It shows the design of the system and explains the types of innovative workflows that it enables for lecturers. I believe the effort is very promising and having the paper to both explain and advertise the initiative is worthy of publication by JGSY. However, I believe there are some points that can improve the contribution of the paper, as follows.

\begin{point}
On page 4, line 21, please write in full the acronym MOOC.
\end{point}
\begin{reply}
  Although MOOC had been defined earlier, we have changed to just 'online course'.
\end{reply}

\begin{point}
On page 5, the authors try to explain what they call Atoms, Sessions and Modules. The differentiation between these is however not very clear at this point. I suggest maybe adding a box to better explain what each of these elements is and how they differ. \end{point}
\begin{reply}
A small table has been created to outline the relationship between these elements.
\end{reply}

\begin{point}
On page 6, line 45, the paper explains how to select only parts of an Atom. The example with 'a.super' doesn't really deliver a good explanation, since it is not really clear what this class is. A more extensive example may be more helpful.

\end{point}

\begin{reply}
We have rewritten this section substantially so that it is more clear to the reader and added a figure to try to illustrate what we mean \textbf{before} we get to the more complicated representation further on.
\end{reply}

\begin{point}
Actually, to have a full Jupyter notebook as an example, with print-screens may be a good idea to achieve the objective of addressing the "the cohort of teachers at all levels" (p.11). On page 7, the authors present an example of how to include an Atom in a Session. However, for a teacher at a beginner level, it may not be clear where/how that code is supposed to be inserted. Hence, I believe having a full notebook as example would really be more helpful to achieve the beginner level of teachers. In my opinion, this is the greatest weakness of the paper: not showing a full example that can be used by beginners to build upon it and from there start creating their own Sessions with a mixing approach as proposed by the authors. I strongly suggest the addition of a full example to the paper.

\end{point}

\begin{reply}
A working repository has been created with useable content illustrating two of the potential applications for Geopyter. However, our feeling is that including full Python/Jupyter code here would be confusing.
\end{reply}

\begin{point}
The url used for the include examples takes the reader to a BlueHost advertising page. The authors should change it.

\end{point}

\begin{reply} The repository for the source code is on
\url{https://github.com/pysal/geopyter}. A live version of geopyter can be
accessed at \url{https://mybinder.org/v2/gh/pysal/geopyter/master},
where the user can explore working examples without having to install geopyter
locally.
\end{reply}

\begin{point}
The authors mention that as the user base grows, so will the contributor base and from there a cumulative circular causation would start operating to increase the platform. I remain, however, in doubt of how Geopyter will be able to organise all this varying mix of contents and keep all of it following standard patterns of coding, documentation, etc. It would be interesting to know how this is planned.

\end{point}

\begin{reply}
We have added a subsection regarding this issue: it's an important point and we have some experience navigating these issues thanks to one author's work with PySAL. Clearly, developing teaching resources is more challenging than developing executable code; however, we believe that by placing groups of related Atoms under the 'control' of a small number of dedicated contributors a degree of coherence of style and presentation will be developed such that students on an introductory or specialist module would not encounter wide variation.
\end{reply}

\begin{point}

In sum, I believe the initiative, and hence the paper, is very interesting. However, I am not confident the paper can, as it currently stands, be of much help to beginning lecturers as it proposes. Some work on it is required in order to achieve this.

\end{point}

\begin{reply}
Thank you for your constructive suggestions. We feel that by taking them on in
this revision, the manuscript has been substantially improved.
\end{reply}

\pagebreak
% Begin a new reviewer section
\reviewersection
 This manuscript introduces a novel and innovative instructional system that integrates code with instructional material in mix-and-matchable Jupyter Notebooks. The manuscript is well written, but does have a few sections that mix technical concepts with domain concepts, for example. This is a useful contribution to the literature and with some minor changes is publishable. I added more detailed comments below.

\begin{point}
How does Open Education Resources (OER) factor into this work? This is a growing
movement in the education space and intersects with OSS and MOOCs as well. The
authors should consider discussing this as related work.

\end{point}

\begin{reply}
We thank the reviewer for drawing our attention to this literature—it is not something we'd previously come across even though we have been regularly publishing our teaching material and source code!
\end{reply}

\begin{point}

Introduction
P1l14 - I would recommend removing the 'online' qualifier and reference simply teaching of these technologies.

\end{point}

\begin{reply}
  Done.
\end{reply}

\begin{point}
Pl15 - This is a jarring pivot that readers may not fully follow. Readers without a background in data science may not know that there is an 'open' ethos nor what an 'open' ethos entails. So a brief 1-sentence introduction for these topics may be useful.

\end{point}

\begin{reply}
First paragraph rewritten to remove reference to 'open' as this made it simpler while removing the jarring effect.
\end{reply}

\begin{point}

I would recommend that the introduction introduce and cite Jupyter Notebooks. 
\end{point}

\begin{reply}
  Done
\end{reply}

\begin{point}

P2l12 - This first paragraph mixes curriculum with software. It would be helpful to disentangle these two elements while also describing their interrelationships. 
\end{point}

\begin{reply}
Tweaked to try to make this distinction more clear.  
\end{reply}

\begin{point}

P2l48 - space missing after Rey 2019 citation, the citation is the incorrect style. 
\end{point}

\begin{reply}
Done.  
\end{reply}

\begin{point}

P2l50 - wrong citation style
\end{point}

\begin{reply}
  Corrected.

\end{reply}

\begin{point}

P3l16 - e.g., should be before the citation
\end{point}

\begin{reply}

  Corrected.
\end{reply}

\begin{point}

P3l29 - the authors may consider broadening the quote to say something like: 'computational social/environmental/geospatial scientist' to make it more inclusive to a broader readership.  
\end{point}

\begin{reply}
  Done.

\end{reply}

\begin{point}

P3l34 - speed can also be an issue in these cases.
\end{point}

\begin{reply}
Indeed!
\end{reply}

\begin{point}
P5l4 - An introduction of Git and Github would be helpful for readers that don't have a lot of experience with them.

\end{point}

\begin{reply}
We've tweaked this slightly to try to make their nature a bit more clear but think that a more detailed introduction to Git/GitHub is not appropriate in this article.  
\end{reply}

\begin{point}
Figure 1. This is a nice solution. The ability to test notebooks using the Markdown header hierarchy is very elegant. 
\end{point}

\begin{reply}
  Thank you!

\end{reply}

\begin{point}
Use Cases
P8l22 - Define HE

\end{point}

\begin{reply}
  Done.

\end{reply}

\begin{point}
The descriptions in the use cases are clear. However, it would be nice to see a more specific use case with actual materials in addition to these broader discussion use cases. 
\end{point}

\begin{reply}
This is was a request also made by Reviewer #1: it is difficult to know how we
can provide a full demonstration of the tool \emph{within} the article itself and we had always planned to provide a core set of 'Atoms' as part of the publication of the article, so we have opted to focus on the core concept(s) in the article and to leave a fuller demonstration for the 'Supplementary Online Materials'. However, we \emph{also} recognise the challenges of getting one's head around the idea in the first place and have attempted to deal with this both by tidying up our explication and by offering more illustrations of how the process takes a document's implicit structure and uses that to extract relevant sections from another notebook.
\end{reply}

\begin{point}
The rest of the manuscript summarizes the key points and extensions. I was happy to see the potential connection to other projects such as the Body of Knowledge.

\end{point}

\begin{reply}
Thank you for your constructive suggestions. We are glad you see the value in
linking this work to projects such as the Body of Knowledge.
\end{reply}

\pagebreak
\reviewersection

\begin{point}
Fascinating to see how the development of this resource came about, and how its design aims to fill in a niche discovered within the teaching of spatial analytics. I was intrigued by the idea that this could connect the streams from the FOSS and MOOC movements, as the abstract pointed towards, but found myself not completely understanding how this fit in that trajectory after all in the sections describing the process. I believe a better historical grounding with actual telling of timelines and developments in those powerful movements would have situated this GeoPyTeR story more strongly.

\end{point}

\begin{reply}

We hope that the substantial edits and incorporation of OER literature has helped to position GeoPyTer more clearly in the reviewer's mind. We would suggest that a timeline is not strictly relevant here as GeoPyTer is not a MOOC and does not depend on any specific FOSS technologies other than iPython/Jupyter (where we've slightly adjusted our initial presentation of the material) to make this more clear.

\end{reply}

\begin{point}

That said, this design narrative seems a bit smoothed over. It omits reflection
on some of the challenges, and glosses over what might be a good opportunity to
reveal what might be remaining or conquered flaws, false starts,
wheel-reinventing, or other elements that surely hold some insights for audience
of this journal.

\end{point}

\begin{reply}
We have tried to expand on some of the unresolved challenges
and in particular that of ensuring overall coherence in the examples (including
data!) provided. Space constraints precluded a more thorough examination of the
'false starts', but we hope that revisions to the initial submission have helped
in this regard.
\end{reply}

\begin{point}
Not being a developer, I had a hard time following some of the heavy jargon in
many spots that might only need some signposts to reach a broader community of
interest. It is important to identify the "instructors" audience more clearly: I
am assuming that this is for university faculty teaching to undergraduates /
graduates at mostly US / European institutions, but if that assumption is wrong,
that underscores the need for the author to make that explicit.

\end{point}

\begin{reply}
We have
specified a bit more clearly that the default assumption here is of technically
sophisticated HE staff, but we also wish to be highlight that this is not the
*only* group that could make use of such material. Hopefully, our edits and
additions (especially the illustrative material) have mitigated at least a
little bit some of the more forbidding jargon.

\end{reply}

\begin{point} The introduction should start with the problem and purpose. Two
other brief mentions seem missed opportunities: the reference to "best
practices" and reference to the Body of Knowledge, which seem could catapult
uptake of such a resource.
\end{point}

\begin{reply}
We have included an earlier reference to the BoK
where we think that this helps to set expectations about the managed nature of
contributions.

\end{reply}

\begin{point} Finally, it surprised me that the use case scenarios were only
"potential use" and not actual end-user cases. This led me to wonder who is
using the resource in what actual context? How many? where? for which content?
for which students? Does it rise to the usage potential with MOOCs and reach a
massive audience? would some of the pitfalls of MOOCs befall this resource as
well or not? How might some of the chalelnges of the open movement be resolved
or remain? This also reveals how the article is devoid of time and timelines,
which would orient the reader.

\end{point}

\begin{reply}
This does not replace or augment MOOCs, nor do
we expect it to have the same kind of impact as it sits alongside such tools
within domains where specific technical training is called for. We have used it
in fits and starts already, and have made a set of Atoms and Sessions available
via the Supplementary Online Material, but our own challenges of not having
enough time to either finish development and write up the tool for publication
have hampered wider adoption. We are hoping to secure funding to allow further
development but felt that we needed to 'get the word out'.
\end{reply}

\end{document}

### 2020-03-04 Geopyter.tex

@@
 
-Regardless of whether this is a revolution \citep{Wyly2014,Torrens2010} or an evolution \citep{Barnes2013,Barnes2014}, there is no question that teaching students and researchers how to code and how to \emph{think} computationally presents new challenges \citep{Etherington2016,Muller2014,rey_09jgs} and that there is little in the way of a tradition of doing so to a high level within Geography (see surveys in \citealp{Bowlick2017,Bowlick2018}).
+Regardless of whether this is a revolution \citep{Wyly2014,Torrens2010} or an
+evolution \citep{Barnes2013,Barnes2014}, there is no question that teaching
+students and researchers how to code and how to \emph{think} computationally \citep{Barba_2015_bids} presents new challenges \citep{Etherington2016,Muller2014,rey_09jgs} and that there is little in the way of a tradition of doing so to a high level within Geography (see surveys in \citealp{Bowlick2017,Bowlick2018}).
 Consequently, the majority of spatial data science and analytics resources used in the classroom are being developed from scratch at each institution, often by newly-appointed lecturers and assistant professors coming from the relatively small number of elite research facilities active in this domain (see discussion in \citealp{esrc2013}).
@@
 In short, the instructor writes whatever content they wish but, using syntax similar to the examples below, wherever they want to incorporate contributed content from \gp (or elsewhere) they have only to `include' it by specifying both a source and a selection.
-And this approach works recursively: a notebook can include content from a notebook that itself includes content from another notebook\ldots
-
+And this approach works recursively: a notebook can include content from a notebook that itself includes content from another notebook.
 \subsection{Syntax}\label{syntax}
@@
 The extensible nature of the \textsc{json} format also allows us to read and write both data and metadata not only to each notebook, but also to each and every cell in a notebook.
-Since metadata that is not understood by Jupyter is simply ignored, we can add our own fields to provide useful information related to instruction such as who should be given credit for contributing and any dependencies or requirements for installed libraries\ldots
+Since metadata that is not understood by Jupyter is simply ignored, we can add our own fields to provide useful information related to instruction such as who should be given credit for contributing and any dependencies or requirements for installed libraries.
 
@@
 
-Associate Professor X wishes to offer a \emph{Foundations of Spatial Data Science} module to second year undergraduates, but anticipates high demand exceeding her capacity to teach in a a computer cluster and high attrition since many students will decide that programming is `not for them'.
+Associate Professor X wishes to offer a \emph{Foundations of Spatial Data Science} module to second year undergraduates, but anticipates high demand exceeding her capacity to teach in  a computer cluster and high attrition since many students will decide that programming is `not for them'.
 Consequently, the decision is made to offer this module on a distance-learning basis with pre-recorded video content and other rich media to support student learning.
@@
 
-The instructor might therefore simply import the same group of Atoms as above, but use a Session and Module template that leaves out the rich media and more detailed explanations since these will be discussed in-class with the students.
-The Sessions then wrap up with an additional mini-project or mini-assessment that requires the students to translate the concepts into a new problem domain or investigate the process in more detail: ``we've seen how we can use \texttt{pandas} and \texttt{bokeh} to explore and compare the distribution of demographic groups in London, here's a link to equivalent open data for Phoenix, Arizona\ldots{}''
+The instructor might therefore simply import the same group of Atoms as above,
+but use a Session and Module template that leaves out the rich media and more
+detailed explanations since these will be discussed in-class with the students.
+The Sessions then wrap up with an additional mini-project or mini-assessment
+that requires the students to translate the concepts into a new problem domain
+or investigate the process in more detail: ``we've seen how we can use
+\texttt{pandas} and \texttt{bokeh} to explore and compare the distribution of
+demographic groups in London, here's a link to equivalent open data for Phoenix,
+Arizona\ldots{}''.\footnote{Interactive examples that demonstrate these use
+  cases are available on-line at \url{https://mybinder.org/v2/gh/pysal/geopyter/master}.}
 
@@
 \subsection{Coordination}
-
-\comment{Is this characterisation accurate?}
 

### 2021-01-04 Geopyter.tex
implemented final revisions requested. Not sure that link to GitHub page fulfills the replicability request but it moves us closer.

@@
 University educators in many disciplines are playing catch-up, but Geography---thanks to its long relationship with computation \citep{ArribasBel2018}---has been acutely aware of these developments: changes in the volume and extent of spatial data available (\eg \citealp{Graham2013,gonzalez2013big,reades2016}), and in the number and range of methods employed to identify patterns in those data (\eg \citealp{Fan2016,Naik2017,Santibanez2015,Stevens2015,ArribasBel2017}), are allowing us to tackle questions thought unanswerable just a few years ago.
-At the same time, there is evidence of a `hollowing out' of once domain-specific skills \citep{Singleton2014,Singleton2016}: most of what used to be done with specialist Geographic Information Systems (\textsc{gis}) software can now be done using free online resources (\eg Google's Fusion Tables, MapBox, Carto, ArcOnline).
-However, there remains unmet demand for graduates who can \emph{code} using the latest (geo)algorithms and techniques, so in this article we present one means of bringing the benefits of (spatial) data science-approaches to instruction using Jupyter `notebooks' for the teaching of the Python programming language and geospatial analysis.
+In academia, there has been a `turn' towards coding (\eg \citealp{Brunsdon:2020aa}), but this has not been matched by changes in our teaching practice and there is evidence of a `hollowing out' of once domain-specific skills \citep{Singleton2014,Singleton2016}: most of what used to be done with specialist Geographic Information Systems (\textsc{gis}) software can now be done using free online resources (\eg Google's Fusion Tables, MapBox, Carto, ArcOnline).
+However, most countries have actually seen an increase in `geospatial jobs' \citep{Solis:2020aa} indicating unmet demand for graduates who can \emph{code} using the latest (geo)algorithms and techniques.
+In this article we present one means of bringing the benefits of (spatial) data science-approaches to instruction using Jupyter `notebooks' for the teaching of the Python programming language and geospatial analysis.
 
@@
 Our approach therefore seeks to develop a rich community of developers and teachers who work \emph{together} on new teaching materials, who make their output entirely \emph{open} to the wider community, and who \emph{support} instructors in quickly adapting existing materials to new delivery formats.
+It should also be noted that our tool is fully compatible with notebooks written in other languages, such as R and its powerful geospatial analytics framework (see \citealp{Bivand:2020aa} for an overview).
 
@@
 Without realising it, we were plugging into a larger framework of thinking emerging from the literature on Open Educational Resources (\textsc{oer}), which began as largely `legal and economic concept' designed to support royalty and license free access to educational resources \citep{butcher2010open}.
-It should be clear from this definition that there is no \emph{necessary} link to \textsc{foss} since a free e-book can qualify as an \textsc{oer}.
+It should be clear from this definition that there is no \emph{necessary} link to \textsc{foss}, or to open and reproducible research \citep{Brunsdon:2020aa}, since a free e-book can qualify as an \textsc{oer}.
 However, these two areas are conceptually aligned in many ways, not least through their shared interest in the `4 Rs': reuse, redistribution, revision, and remixing \citep{Hilton2010}.
@@
 
+\section{Supplemental}
+
+Demonstration notebooks can be found in the `sessions' directory of the \gp project on GitHub: \href{https://github.com/pysal/geopyter/tree/master/sessions}{github.com/pysal/geopyter/tree/master/sessions}.
+
 \bibliographystyle{apacite}
